---
title: "Briefing: Freshwater Recreational Effort by Mode & Location"
subtitle: "Preparing for PST Economic Study Consultant Meeting"
author: "WDFW Freshwater Creel"
date: today
execute:
  warning: false
  message: false
---

## Setup

```{r}
#| label: setup

library(creelutils)
library(tidyverse)
library(readxl)
library(hms)
library(glue)
library(gt)
library(here)
library(DT)

# Create output directory for CSV exports
# output_dir, persist_dir (below), and assembly_dir (later in this document)
# all resolve to analysis/pst/outputs/ — kept as three separate names so the
# code at each write site still reads the same and stays reviewable.
output_dir <- here::here("analysis", "pst", "outputs")
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

# Cache directory for raw data pulls
cache_dir <- here::here(".cache")
if (!dir.exists(cache_dir)) dir.create(cache_dir, recursive = TRUE)
```

This document produces Track B of the freshwater effort deliverable: guided/
unguided × bank/boat proportions derived from creel interview data, at the
fishery, monthly, and weekly grain, for application against CRC-based total
trip estimates. The analytical framework these proportions plug into — the
P1/P2/P3 tier hierarchy, the Track A vs. Track B split, assumptions A1–A4,
and the rule that ratios do not travel between blocks — is documented in the
parent document, `analysis/pst/pst_fw_angler_trips.qmd`. This file is a
standalone producer: it is run manually to refresh Track B and is not
included by the parent document.

## Pull All Fishery Interview Data

```{r}
#| label: pull-all-interviews

cache_file <- file.path(cache_dir, "all_interviews.rds")

if (file.exists(cache_file)) {
  cat("Loading cached interview data from", cache_file, "\n")
  cat("Cache date:", format(file.mtime(cache_file), "%Y-%m-%d %H:%M"), "\n")
  all_interviews <- readRDS(cache_file)
} else {
  # Connect and get fishery names
  conn <- connect_creel_db()
  fishery_names <- fishery_lut(conn = conn) |>
    pull(fishery_name) |>
    unique()

  cat(glue("{length(fishery_names)} fisheries found in fishery_lut\n\n"))

  # Iterate over all fisheries and bind interview data
  all_interviews <- purrr::map(
    fishery_names,
    \(fn) {
      tryCatch(
        {
          dat <- fetch_data(
            conn = conn,
            fishery_name = fn,
            tables = "interview",
            data_source = "internal"
          )
          # Coerce all columns to character before binding to prevent
          # type conflicts (logical from all-NA, double vs character, etc.)
          dat$interview |>
            mutate(across(everything(), as.character))
        },
        error = function(e) {
          cli::cli_alert_warning("Failed for {fn}: {e$message}")
          NULL
        }
      )
    }
  ) |>
    purrr::compact() |>
    bind_rows()

  DBI::dbDisconnect(conn)

  # Save to cache
  saveRDS(all_interviews, cache_file)
  cat(glue("Cached {nrow(all_interviews)} rows to {cache_file}\n"))
}

cat(glue("{nrow(all_interviews)} total interview records across {n_distinct(all_interviews$fishery_name)} fisheries\n"))

# Persist alongside the pipeline's other outputs, not just .cache/. The cache
# exists purely to skip a slow DB pull on re-render and is expected to be
# deleted/ignored; analysis/pst/outputs/ is where every other script in this
# repo (multi_fishery_creel_summary.R, the CRC parser, etc.) writes what it
# actually produces, and downstream users shouldn't have to know this
# document has its own separate cache directory to find the underlying
# interview data.
persist_dir <- here::here("analysis", "pst", "outputs")
if (!dir.exists(persist_dir)) dir.create(persist_dir, recursive = TRUE)

saveRDS(all_interviews, file.path(persist_dir, "all_interviews.rds"))
write_csv(all_interviews, file.path(persist_dir, "all_interviews.csv"))
cat(glue("Persisted {nrow(all_interviews)} rows to ",
         "{file.path(persist_dir, 'all_interviews.{rds,csv}')}\n"))
```

::: {.callout-tip}
To force a fresh pull from the database, delete `.cache/all_interviews.rds`
and re-render.
:::

::: {.callout-note}
**Hanford Reach / Yakima / McNary are not pulled through this document.**
Tried folding them into the standard pull above via `fishery_lut()` +
`fetch_data()`; a diagnostic sweep across `fishery_name`, `project_name`, and
`water_body` found zero matching rows anywhere, including a direct
`fishery_lut()` re-query, so these three rivers aren't reachable through the
per-fishery loop under any field this document checked. Not worth chasing
further, because neither Track A nor Track B in the assembly script actually
depends on it:

- **Location (bank/boat)** for these rivers already comes from `angler_final`
  in R3's own trip data (`mid_columbia_yakima_creel_summary.csv`, source_id
  `R3_external`) — `pst_fw_angler_trips_assembly.R` reads it directly as
  `location_basis = "design_stratum"`. This document's interview pull was
  never the source for that.
- **Mode (guided/unguided)** can't be resolved from interviews regardless of
  whether the pull works — R3's programs don't collect a `trip_guided` field
  at all, a structural gap independent of how the data gets pulled. It's
  already coded `mode_basis = "not_collected"` in the assembly script, not an
  unstated zero.

So this document's proportions apply only to Puget Sound and WA Coast, and
`R3_external` rivers deliver with `mode = "unknown"` by design. See
`pst_input_manifest.csv` for the record of what was attempted.
:::

## Derive Classification Fields

The interview view contains:

- `trip_guided`: "Guided" / "Non-guided" / "Unknown" (newer fisheries)
- `angler_type`: "Bank" / "Boat" / "Unk" (older fisheries, direct classification)
- `boat_used`: "Yes" / "No" (newer fisheries)
- `fish_from_boat`: "Boat" / "Bank" (supplement for parties with boats)
- `boat_type`: vessel detail (Pontoon/Kayak/Drift Boat/Motorized)

Bank/boat derivation follows the logic in
[`prep_dwg_interview_angler_types()`](https://github.com/wdfw-fp/CreelEstimates/blob/main/R_functions/prep_dwg_interview_angler_types.R):
older datasets use `angler_type` directly; newer datasets derive from `boat_used`.

Fishing hours follow
[`prep_dwg_interview_fishing_time()`](https://github.com/wdfw-fp/CreelEstimates/blob/main/R_functions/prep_dwg_interview_fishing_time.R):
for incomplete trips or missing end times, `interview_time` substitutes as
the end time. Total fishing hours = per-party fishing time × angler count.

```{r}
#| label: derive-fields

# Helper: safely parse HH:MM:SS strings to hms; returns NA on failure
safe_hms <- function(x) {
  out <- suppressWarnings(hms::as_hms(x))
  out[is.nan(as.numeric(out))] <- NA
  out
}

interviews <- all_interviews |>
  # Normalize sentinel strings ("NA", "", whitespace-only) to proper NA
  mutate(across(everything(), ~ if_else(.x %in% c("NA", "", " "), NA_character_, .x))) |>
  mutate(
    event_date = as.Date(event_date),
    year = year(event_date),
    month = floor_date(event_date, "month"),
    week = floor_date(event_date, "week"),
    # Guided status: standardize trip_guided to binary
    guided_status = case_when(
      trip_guided == "Guided"     ~ "Guided",
      trip_guided == "Non-guided" ~ "Unguided",
      TRUE                        ~ NA_character_
    ),
    # Bank/Boat: use angler_type if available, otherwise derive from boat_used
    angler_final = case_when(
      # Older datasets with direct angler_type classification
      !is.na(angler_type) & angler_type == "Bank" ~ "Bank",
      !is.na(angler_type) & angler_type == "Boat" ~ "Boat",
      # Newer datasets: derive from boat_used
      boat_used == "No"  ~ "Bank",
      boat_used == "Yes" & fish_from_boat == "Bank" ~ "Bank",
      boat_used == "Yes" ~ "Boat",
      TRUE ~ NA_character_
    ),
    # Fishing time (hours) per CreelEstimates methodology:
    # Use fishing_end_time for complete trips; interview_time for incomplete/missing
    trip_status = tidyr::replace_na(trip_status, "Unknown"),
    end_time_final = if_else(
      trip_status == "Incomplete" | is.na(fishing_end_time),
      interview_time,
      fishing_end_time
    ),
    # Safe time parsing — returns NA for malformed/empty strings
    fishing_time = as.numeric(
      safe_hms(end_time_final) - safe_hms(fishing_start_time)
    ) / 3600,
    # Angler count: treat <= 0 as missing (e.g., -1 means "not recorded")
    angler_count_num = as.numeric(angler_count),
    angler_count_num = if_else(angler_count_num <= 0, NA_real_, angler_count_num),
    # Total fishing hours = fishing_time * angler_count
    fishing_time_total = fishing_time * angler_count_num
  ) |>
  # Drop rows with non-positive or missing fishing time
  mutate(
    fishing_time_total = if_else(
      is.na(fishing_time_total) | fishing_time_total <= 0,
      NA_real_,
      fishing_time_total
    )
  )

# Summary of field coverage
cat("Guided status coverage:\n")
interviews |> count(guided_status) |> print()
cat("\nBank/Boat coverage:\n")
interviews |> count(angler_final) |> print()
cat("\nFishing time total (hours) summary:\n")
summary(interviews$fishing_time_total) |> print()
```

## Fisheries Included

```{r}
#| label: fishery-listing

fishery_summary <- interviews |>
  group_by(fishery_name) |>
  summarise(
    n_interviews = n(),
    date_range = paste(min(event_date, na.rm = TRUE), "to", max(event_date, na.rm = TRUE)),
    pct_guided_recorded = round(100 * mean(!is.na(guided_status)), 1),
    pct_bank_boat_recorded = round(100 * mean(!is.na(angler_final)), 1),
    pct_hours_available = round(100 * mean(!is.na(fishing_time_total)), 1),
    .groups = "drop"
  ) |>
  arrange(fishery_name)

cat(glue("{nrow(fishery_summary)} fisheries included in analysis\n"))
```

```{r}
#| label: tbl-fishery-list

fishery_summary |>
  datatable(
    caption = "Fisheries included in analysis (searchable)",
    filter = "top",
    rownames = FALSE,
    options = list(pageLength = 15, autoWidth = TRUE),
    colnames = c(
      "Fishery" = "fishery_name",
      "Interviews" = "n_interviews",
      "Date Range" = "date_range",
      "% Guided/Unguided Recorded" = "pct_guided_recorded",
      "% Bank/Boat Recorded" = "pct_bank_boat_recorded",
      "% Hours Available" = "pct_hours_available"
    )
  )
```

## Calculate Proportions (Interview Counts)

### Fishery-Level Proportions (Counts)

```{r}
#| label: fishery-level-proportions

# Guided vs Unguided proportions by fishery
guide_fishery <- interviews |>
  filter(!is.na(guided_status)) |>
  count(fishery_name, guided_status) |>
  group_by(fishery_name) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

# Bank vs Boat proportions by fishery
type_fishery <- interviews |>
  filter(!is.na(angler_final)) |>
  count(fishery_name, angler_final) |>
  group_by(fishery_name) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

# Export
write_csv(guide_fishery, file.path(output_dir, "guide_proportions_fishery.csv"))
write_csv(type_fishery, file.path(output_dir, "type_proportions_fishery.csv"))
```

### Monthly Proportions (Counts)

```{r}
#| label: monthly-proportions

guide_monthly <- interviews |>
  filter(!is.na(guided_status)) |>
  count(fishery_name, month, guided_status) |>
  group_by(fishery_name, month) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

type_monthly <- interviews |>
  filter(!is.na(angler_final)) |>
  count(fishery_name, month, angler_final) |>
  group_by(fishery_name, month) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

# Export
write_csv(guide_monthly, file.path(output_dir, "guide_proportions_monthly.csv"))
write_csv(type_monthly, file.path(output_dir, "type_proportions_monthly.csv"))
```

### Weekly Proportions (Counts)

```{r}
#| label: weekly-proportions

guide_weekly <- interviews |>
  filter(!is.na(guided_status)) |>
  count(fishery_name, week, guided_status) |>
  group_by(fishery_name, week) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

type_weekly <- interviews |>
  filter(!is.na(angler_final)) |>
  count(fishery_name, week, angler_final) |>
  group_by(fishery_name, week) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

# Export
write_csv(guide_weekly, file.path(output_dir, "guide_proportions_weekly.csv"))
write_csv(type_weekly, file.path(output_dir, "type_proportions_weekly.csv"))
```

## Calculate Proportions (Combined: Guide × Bank/Boat)

Cross-classification of guided status and angler type into four categories
most applicable to expanding estimated angler trips.

```{r}
#| label: combined-derive

# Create combined category
interviews <- interviews |>
  mutate(
    combined_type = case_when(
      !is.na(guided_status) & !is.na(angler_final) ~
        paste(guided_status, angler_final, sep = "-"),

      TRUE ~ NA_character_
    )
  )

cat("Combined type coverage:\n")
interviews |> count(combined_type) |> print()
```

### Fishery-Level (Combined, Counts)

```{r}
#| label: combined-fishery-counts

combined_fishery <- interviews |>
  filter(!is.na(combined_type)) |>
  count(fishery_name, combined_type) |>
  complete(fishery_name, combined_type, fill = list(n = 0L)) |>
  group_by(fishery_name) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

write_csv(combined_fishery, file.path(output_dir, "combined_proportions_fishery.csv"))
```

### Monthly (Combined, Counts)

```{r}
#| label: combined-monthly-counts

combined_monthly <- interviews |>
  filter(!is.na(combined_type)) |>
  count(fishery_name, month, combined_type) |>
  complete(nesting(fishery_name, month), combined_type, fill = list(n = 0L)) |>
  group_by(fishery_name, month) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

write_csv(combined_monthly, file.path(output_dir, "combined_proportions_monthly.csv"))
```

### Weekly (Combined, Counts)

```{r}
#| label: combined-weekly-counts

combined_weekly <- interviews |>
  filter(!is.na(combined_type)) |>
  count(fishery_name, week, combined_type) |>
  complete(nesting(fishery_name, week), combined_type, fill = list(n = 0L)) |>
  group_by(fishery_name, week) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

write_csv(combined_weekly, file.path(output_dir, "combined_proportions_weekly.csv"))
```

### Fishery-Level (Combined, Hours)

```{r}
#| label: combined-fishery-hrs

combined_fishery_hrs <- interviews |>
  filter(!is.na(combined_type), !is.na(fishing_time_total)) |>
  group_by(fishery_name, combined_type) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  complete(fishery_name, combined_type, fill = list(hours = 0)) |>
  group_by(fishery_name) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

write_csv(combined_fishery_hrs, file.path(output_dir, "combined_proportions_fishery_hrs.csv"))
```

### Monthly (Combined, Hours)

```{r}
#| label: combined-monthly-hrs

combined_monthly_hrs <- interviews |>
  filter(!is.na(combined_type), !is.na(fishing_time_total)) |>
  group_by(fishery_name, month, combined_type) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  complete(nesting(fishery_name, month), combined_type, fill = list(hours = 0)) |>
  group_by(fishery_name, month) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

write_csv(combined_monthly_hrs, file.path(output_dir, "combined_proportions_monthly_hrs.csv"))
```

### Weekly (Combined, Hours)

```{r}
#| label: combined-weekly-hrs

combined_weekly_hrs <- interviews |>
  filter(!is.na(combined_type), !is.na(fishing_time_total)) |>
  group_by(fishery_name, week, combined_type) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  complete(nesting(fishery_name, week), combined_type, fill = list(hours = 0)) |>
  group_by(fishery_name, week) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

write_csv(combined_weekly_hrs, file.path(output_dir, "combined_proportions_weekly_hrs.csv"))
```

### Searchable Fishery Proportion Table (Combined)

```{r}
#| label: tbl-combined-searchable

# Wide-format summary: one row per fishery with proportions for each combined type
combined_wide <- combined_fishery |>
  select(fishery_name, combined_type, n, total, prop) |>
  pivot_wider(
    names_from = combined_type,
    values_from = c(n, prop),
    names_glue = "{combined_type}_{.value}"
  ) |>
  rename(total_interviews = total) |>
  distinct()

# Also join hours-based proportions
combined_wide_hrs <- combined_fishery_hrs |>
  select(fishery_name, combined_type, prop) |>
  pivot_wider(
    names_from = combined_type,
    values_from = prop,
    names_prefix = "hrs_"
  )

combined_summary <- combined_wide |>
  left_join(combined_wide_hrs, by = "fishery_name")

write_csv(combined_summary, file.path(output_dir, "combined_proportions_summary.csv"))

combined_summary |>
  datatable(
    caption = "Combined guided × bank/boat proportions by fishery (searchable)",
    filter = "top",
    rownames = FALSE,
    options = list(pageLength = 15, scrollX = TRUE, autoWidth = TRUE)
  ) |>
  formatPercentage(columns = grep("prop|hrs_", names(combined_summary), value = TRUE), digits = 1)
```

## Calculate Proportions (Fishing Hours)

Proportions based on total angler fishing hours (fishing_time × angler_count)
per [`prep_dwg_interview_fishing_time()`](https://github.com/wdfw-fp/CreelEstimates/blob/main/R_functions/prep_dwg_interview_fishing_time.R).

### Fishery-Level Proportions (Hours)

```{r}
#| label: fishery-level-proportions-hrs

# Guided vs Unguided by total fishing hours
guide_fishery_hrs <- interviews |>
  filter(!is.na(guided_status), !is.na(fishing_time_total)) |>
  group_by(fishery_name, guided_status) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  group_by(fishery_name) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

# Bank vs Boat by total fishing hours
type_fishery_hrs <- interviews |>
  filter(!is.na(angler_final), !is.na(fishing_time_total)) |>
  group_by(fishery_name, angler_final) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  group_by(fishery_name) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

# Export
write_csv(guide_fishery_hrs, file.path(output_dir, "guide_proportions_fishery_hrs.csv"))
write_csv(type_fishery_hrs, file.path(output_dir, "type_proportions_fishery_hrs.csv"))
```

### Monthly Proportions (Hours)

```{r}
#| label: monthly-proportions-hrs

guide_monthly_hrs <- interviews |>
  filter(!is.na(guided_status), !is.na(fishing_time_total)) |>
  group_by(fishery_name, month, guided_status) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  group_by(fishery_name, month) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

type_monthly_hrs <- interviews |>
  filter(!is.na(angler_final), !is.na(fishing_time_total)) |>
  group_by(fishery_name, month, angler_final) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  group_by(fishery_name, month) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

# Export
write_csv(guide_monthly_hrs, file.path(output_dir, "guide_proportions_monthly_hrs.csv"))
write_csv(type_monthly_hrs, file.path(output_dir, "type_proportions_monthly_hrs.csv"))
```

### Weekly Proportions (Hours)

```{r}
#| label: weekly-proportions-hrs

guide_weekly_hrs <- interviews |>
  filter(!is.na(guided_status), !is.na(fishing_time_total)) |>
  group_by(fishery_name, week, guided_status) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  group_by(fishery_name, week) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

type_weekly_hrs <- interviews |>
  filter(!is.na(angler_final), !is.na(fishing_time_total)) |>
  group_by(fishery_name, week, angler_final) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  group_by(fishery_name, week) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

# Export
write_csv(guide_weekly_hrs, file.path(output_dir, "guide_proportions_weekly_hrs.csv"))
write_csv(type_weekly_hrs, file.path(output_dir, "type_proportions_weekly_hrs.csv"))
```

## Variability Analysis

Compare the spread of proportions when calculated at different temporal
aggregations (entire fishery vs monthly vs weekly).

```{r}
#| label: variability-summary

# Helper to safely summarise proportions (returns NA row if no data)
summarise_variability <- function(df, scale_label) {
  if (nrow(df) == 0) {
    return(tibble(
      scale = scale_label, mean_prop = NA_real_, sd_prop = NA_real_,
      cv = NA_real_, min_prop = NA_real_, max_prop = NA_real_, n_units = 0L
    ))
  }
  df |> summarise(
    scale = scale_label,
    mean_prop = mean(prop, na.rm = TRUE),
    sd_prop = sd(prop, na.rm = TRUE),
    cv = sd_prop / mean_prop,
    min_prop = min(prop, na.rm = TRUE),
    max_prop = max(prop, na.rm = TRUE),
    n_units = n()
  )
}

# Guided proportion variability across scales
guide_variability <- bind_rows(
  guide_fishery |> filter(guided_status == "Guided") |> summarise_variability("Fishery (overall)"),
  guide_monthly |> filter(guided_status == "Guided") |> summarise_variability("Monthly"),
  guide_weekly |> filter(guided_status == "Guided") |> summarise_variability("Weekly")
)

# Boat proportion variability across scales
type_variability <- bind_rows(
  type_fishery |> filter(angler_final == "Boat") |> summarise_variability("Fishery (overall)"),
  type_monthly |> filter(angler_final == "Boat") |> summarise_variability("Monthly"),
  type_weekly |> filter(angler_final == "Boat") |> summarise_variability("Weekly")
)
```

### Variability Tables

```{r}
#| label: tbl-guide-variability
#| tbl-cap: "Variability in guided angler proportion across temporal scales"

guide_variability |>
  gt() |>
  fmt_number(columns = c(mean_prop, sd_prop, cv, min_prop, max_prop), decimals = 3) |>
  cols_label(
    scale = "Scale",
    mean_prop = "Mean Proportion",
    sd_prop = "SD",
    cv = "CV",
    min_prop = "Min",
    max_prop = "Max",
    n_units = "N (units)"
  )
```

```{r}
#| label: tbl-type-variability
#| tbl-cap: "Variability in boat angler proportion across temporal scales"

type_variability |>
  gt() |>
  fmt_number(columns = c(mean_prop, sd_prop, cv, min_prop, max_prop), decimals = 3) |>
  cols_label(
    scale = "Scale",
    mean_prop = "Mean Proportion",
    sd_prop = "SD",
    cv = "CV",
    min_prop = "Min",
    max_prop = "Max",
    n_units = "N (units)"
  )

# Export variability summaries
write_csv(guide_variability, file.path(output_dir, "guide_variability_summary.csv"))
write_csv(type_variability, file.path(output_dir, "type_variability_summary.csv"))
```

### Variability Tables (Fishing Hours)

```{r}
#| label: variability-summary-hrs

# Guided proportion variability by fishing hours
guide_variability_hrs <- bind_rows(
  guide_fishery_hrs |> filter(guided_status == "Guided") |> summarise_variability("Fishery (overall)"),
  guide_monthly_hrs |> filter(guided_status == "Guided") |> summarise_variability("Monthly"),
  guide_weekly_hrs |> filter(guided_status == "Guided") |> summarise_variability("Weekly")
)

# Boat proportion variability by fishing hours
type_variability_hrs <- bind_rows(
  type_fishery_hrs |> filter(angler_final == "Boat") |> summarise_variability("Fishery (overall)"),
  type_monthly_hrs |> filter(angler_final == "Boat") |> summarise_variability("Monthly"),
  type_weekly_hrs |> filter(angler_final == "Boat") |> summarise_variability("Weekly")
)
```

```{r}
#| label: tbl-guide-variability-hrs
#| tbl-cap: "Variability in guided angler proportion (fishing hours) across temporal scales"

guide_variability_hrs |>
  gt() |>
  fmt_number(columns = c(mean_prop, sd_prop, cv, min_prop, max_prop), decimals = 3) |>
  cols_label(
    scale = "Scale",
    mean_prop = "Mean Proportion",
    sd_prop = "SD",
    cv = "CV",
    min_prop = "Min",
    max_prop = "Max",
    n_units = "N (units)"
  )
```

```{r}
#| label: tbl-type-variability-hrs
#| tbl-cap: "Variability in boat angler proportion (fishing hours) across temporal scales"

type_variability_hrs |>
  gt() |>
  fmt_number(columns = c(mean_prop, sd_prop, cv, min_prop, max_prop), decimals = 3) |>
  cols_label(
    scale = "Scale",
    mean_prop = "Mean Proportion",
    sd_prop = "SD",
    cv = "CV",
    min_prop = "Min",
    max_prop = "Max",
    n_units = "N (units)"
  )

# Export
write_csv(guide_variability_hrs, file.path(output_dir, "guide_variability_summary_hrs.csv"))
write_csv(type_variability_hrs, file.path(output_dir, "type_variability_summary_hrs.csv"))
```

### Counts vs Hours Comparison

```{r}
#| label: tbl-counts-vs-hours
#| tbl-cap: "Comparing CV of proportions: interview counts vs fishing hours"

bind_rows(
  guide_variability |> mutate(metric = "Guided (counts)", .before = 1),
  guide_variability_hrs |> mutate(metric = "Guided (hours)", .before = 1),
  type_variability |> mutate(metric = "Boat (counts)", .before = 1),
  type_variability_hrs |> mutate(metric = "Boat (hours)", .before = 1)
) |>
  select(metric, scale, mean_prop, cv, n_units) |>
  gt() |>
  fmt_number(columns = c(mean_prop, cv), decimals = 3) |>
  cols_label(
    metric = "Metric",
    scale = "Scale",
    mean_prop = "Mean Proportion",
    cv = "CV",
    n_units = "N"
  )
```

### Combined (Guide × Bank/Boat) Variability

```{r}
#| label: variability-combined

# Variability for each combined category across scales
combined_categories <- c("Guided-Bank", "Guided-Boat", "Unguided-Bank", "Unguided-Boat")

combined_variability <- purrr::map_dfr(combined_categories, \(cat) {
  bind_rows(
    combined_fishery |> filter(combined_type == cat) |> summarise_variability("Fishery (overall)"),
    combined_monthly |> filter(combined_type == cat) |> summarise_variability("Monthly"),
    combined_weekly |> filter(combined_type == cat) |> summarise_variability("Weekly")
  ) |> mutate(category = cat, .before = 1)
})

combined_variability_hrs <- purrr::map_dfr(combined_categories, \(cat) {
  bind_rows(
    combined_fishery_hrs |> filter(combined_type == cat) |> summarise_variability("Fishery (overall)"),
    combined_monthly_hrs |> filter(combined_type == cat) |> summarise_variability("Monthly"),
    combined_weekly_hrs |> filter(combined_type == cat) |> summarise_variability("Weekly")
  ) |> mutate(category = cat, .before = 1)
})

write_csv(combined_variability, file.path(output_dir, "combined_variability_summary.csv"))
write_csv(combined_variability_hrs, file.path(output_dir, "combined_variability_summary_hrs.csv"))
```

```{r}
#| label: tbl-combined-variability
#| tbl-cap: "Variability in combined guide × bank/boat proportions (counts)"

combined_variability |>
  gt(groupname_col = "category") |>
  fmt_number(columns = c(mean_prop, sd_prop, cv, min_prop, max_prop), decimals = 3) |>
  cols_label(
    scale = "Scale",
    mean_prop = "Mean Proportion",
    sd_prop = "SD",
    cv = "CV",
    min_prop = "Min",
    max_prop = "Max",
    n_units = "N (units)"
  )
```

```{r}
#| label: tbl-combined-variability-hrs
#| tbl-cap: "Variability in combined guide × bank/boat proportions (fishing hours)"

combined_variability_hrs |>
  gt(groupname_col = "category") |>
  fmt_number(columns = c(mean_prop, sd_prop, cv, min_prop, max_prop), decimals = 3) |>
  cols_label(
    scale = "Scale",
    mean_prop = "Mean Proportion",
    sd_prop = "SD",
    cv = "CV",
    min_prop = "Min",
    max_prop = "Max",
    n_units = "N (units)"
  )
```

## Fishery Sample Size Breakdown

A per-fishery diagnostic table showing interview counts by category to help
validate proportions and identify fisheries with small or imbalanced samples.

::: {.callout-warning}
**Data collection procedures vary by fishery and need further review.**
Some fisheries show patterns suggesting field-level recording conventions
that inflate the "Unknown" category. For example, Snohomish Fall Salmon
(2023) has nearly all unguided interviews classified as unknown guided
status — it is possible samplers left the guided field blank to imply
unguided rather than explicitly recording "Non-guided." These
fishery-specific data entry issues need to be resolved with project leads
before finalizing proportions for delivery.
:::

```{r}
#| label: fishery-sample-breakdown

fishery_breakdown <- interviews |>
  group_by(fishery_name) |>
  summarise(
    n_total = n(),
    n_guided = sum(guided_status == "Guided", na.rm = TRUE),
    n_unguided = sum(guided_status == "Unguided", na.rm = TRUE),
    n_guide_unknown = sum(is.na(guided_status)),
    n_bank = sum(angler_final == "Bank", na.rm = TRUE),
    n_boat = sum(angler_final == "Boat", na.rm = TRUE),
    n_type_unknown = sum(is.na(angler_final)),
    n_guided_bank = sum(combined_type == "Guided-Bank", na.rm = TRUE),
    n_guided_boat = sum(combined_type == "Guided-Boat", na.rm = TRUE),
    n_unguided_bank = sum(combined_type == "Unguided-Bank", na.rm = TRUE),
    n_unguided_boat = sum(combined_type == "Unguided-Boat", na.rm = TRUE),
    n_combined_unknown = sum(is.na(combined_type)),
    pct_classifiable = round(100 * (1 - n_combined_unknown / n_total), 1),
    mean_hours = round(mean(fishing_time_total, na.rm = TRUE), 2),
    median_hours = round(median(fishing_time_total, na.rm = TRUE), 2),
    total_hours = round(sum(fishing_time_total, na.rm = TRUE), 1),
    year_min = suppressWarnings(min(year(event_date), na.rm = TRUE)),
    year_max = suppressWarnings(max(year(event_date), na.rm = TRUE)),
    n_years = n_distinct(year(event_date), na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    year_min = if_else(is.infinite(year_min), NA_real_, year_min),
    year_max = if_else(is.infinite(year_max), NA_real_, year_max),
    years_span = if_else(
      is.na(year_min),
      "No dates",
      paste0(year_min, "-", year_max, " (", n_years, " yrs)")
    ),
    guide_ratio = paste0(n_guided, ":", n_unguided),
    type_ratio = paste0(n_bank, ":", n_boat)
  ) |>
  arrange(desc(n_total))

write_csv(fishery_breakdown, file.path(output_dir, "fishery_sample_breakdown.csv"))
```

```{r}
#| label: tbl-fishery-breakdown

fishery_breakdown |>
  select(
    fishery_name, n_total, years_span,
    n_guided, n_unguided, n_guide_unknown, guide_ratio,
    n_bank, n_boat, n_type_unknown, type_ratio,
    n_guided_bank, n_guided_boat, n_unguided_bank, n_unguided_boat,
    pct_classifiable, mean_hours, median_hours, total_hours
  ) |>
  datatable(
    caption = "Fishery sample size breakdown by category (searchable, sortable)",
    filter = "top",
    rownames = FALSE,
    options = list(
      pageLength = 20,
      scrollX = TRUE,
      autoWidth = FALSE,
      columnDefs = list(
        list(className = "dt-right", targets = 1:18)
      )
    ),
    colnames = c(
      "Fishery" = "fishery_name",
      "Total N" = "n_total",
      "Years" = "years_span",
      "Guided" = "n_guided",
      "Unguided" = "n_unguided",
      "Unknown (Guide)" = "n_guide_unknown",
      "Guided:Unguided" = "guide_ratio",
      "Bank" = "n_bank",
      "Boat" = "n_boat",
      "Unknown (Type)" = "n_type_unknown",
      "Bank:Boat" = "type_ratio",
      "G-Bank" = "n_guided_bank",
      "G-Boat" = "n_guided_boat",
      "UG-Bank" = "n_unguided_bank",
      "UG-Boat" = "n_unguided_boat",
      "% Classifiable" = "pct_classifiable",
      "Mean Hrs" = "mean_hours",
      "Median Hrs" = "median_hours",
      "Total Hrs" = "total_hours"
    )
  ) |>
  formatStyle("% Classifiable",
    backgroundColor = styleInterval(c(50, 80), c("#f8d7da", "#fff3cd", "#d4edda"))
  )
```

::: {.callout-note}
**Reading this table:**

- **G:UG** = Guided:Unguided count ratio. A fishery showing `0:500` is entirely unguided.
- **B:Bt** = Bank:Boat count ratio.
- **% Classifiable** = percent of interviews where both guided status AND bank/boat are known (usable for the combined 4-way split). Color-coded: red < 50%, yellow 50-80%, green > 80%.
- **Mean/Median Hrs** = average interview fishing hours. Large discrepancies suggest outliers.
- Fisheries with very small N or low % Classifiable should be interpreted with caution.
:::

```{r}
#| label: tbl-flags

# Flag fisheries with potential data quality concerns
flags <- fishery_breakdown |>
  filter(n_total < 30 | pct_classifiable < 50) |>
  select(fishery_name, n_total, pct_classifiable, years_span, guide_ratio, type_ratio) |>
  arrange(pct_classifiable)

if (nrow(flags) > 0) {
  cat(glue("{nrow(flags)} fisheries flagged for small sample size (N<30) or low classifiability (<50%):\n\n"))
  flags |>
    datatable(
      caption = "Flagged fisheries: small N or low classifiability",
      rownames = FALSE,
      options = list(pageLength = 50, dom = "t"),
      colnames = c(
        "Fishery" = "fishery_name",
        "Total N" = "n_total",
        "% Classifiable" = "pct_classifiable",
        "Years" = "years_span",
        "Guided:Unguided" = "guide_ratio",
        "Bank:Boat" = "type_ratio"
      )
    )
} else {
  cat("No fisheries flagged — all have N >= 30 and >= 50% classifiable.\n")
}
```

## Visualizations

### Guided Proportion by Time Scale

```{r}
#| label: fig-guide-distribution
#| fig-cap: "Distribution of guided angler proportions across temporal scales"
#| fig-height: 8
#| fig-width: 10

bind_rows(
  guide_fishery |>
    filter(guided_status == "Guided") |>
    select(fishery_name, prop) |>
    mutate(scale = "Fishery"),
  guide_monthly |>
    filter(guided_status == "Guided") |>
    select(fishery_name, prop) |>
    mutate(scale = "Monthly"),
  guide_weekly |>
    filter(guided_status == "Guided") |>
    select(fishery_name, prop) |>
    mutate(scale = "Weekly")
) |>
  mutate(scale = factor(scale, levels = c("Fishery", "Monthly", "Weekly"))) |>
  ggplot(aes(x = prop, fill = scale)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~scale, ncol = 1, scales = "free_y") +
  scale_x_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    x = "Proportion Guided",
    y = "Density",
    title = "Distribution of Guided Angler Proportions",
    subtitle = "Comparing fishery-level, monthly, and weekly calculations"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

### Bank/Boat Proportion by Time Scale

```{r}
#| label: fig-type-distribution
#| fig-cap: "Distribution of boat angler proportions across temporal scales"
#| fig-height: 8
#| fig-width: 10

bind_rows(
  type_fishery |>
    filter(angler_final == "Boat") |>
    select(fishery_name, prop) |>
    mutate(scale = "Fishery"),
  type_monthly |>
    filter(angler_final == "Boat") |>
    select(fishery_name, prop) |>
    mutate(scale = "Monthly"),
  type_weekly |>
    filter(angler_final == "Boat") |>
    select(fishery_name, prop) |>
    mutate(scale = "Weekly")
) |>
  mutate(scale = factor(scale, levels = c("Fishery", "Monthly", "Weekly"))) |>
  ggplot(aes(x = prop, fill = scale)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~scale, ncol = 1, scales = "free_y") +
  scale_x_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    x = "Proportion Boat Anglers",
    y = "Density",
    title = "Distribution of Boat Angler Proportions",
    subtitle = "Comparing fishery-level, monthly, and weekly calculations"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

### Per-Fishery Time Series

```{r}
#| label: fig-guide-timeseries
#| fig-cap: "Guided proportion over time by fishery (monthly)"
#| fig-height: 10
#| fig-width: 12

# Show top 12 fisheries by interview count for readability
top_fisheries <- interviews |>
  filter(!is.na(guided_status)) |>
  count(fishery_name, sort = TRUE) |>
  slice_head(n = 12) |>
  pull(fishery_name)

guide_monthly |>
  filter(fishery_name %in% top_fisheries, guided_status == "Guided") |>
  ggplot(aes(x = month, y = prop)) +
  geom_line() +
  geom_point(aes(size = total), alpha = 0.5) +
  facet_wrap(~fishery_name, scales = "free_x", ncol = 3) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    x = NULL, y = "Proportion Guided",
    size = "Sample Size",
    title = "Monthly Guided Proportion (Top 12 Fisheries by Interview Count)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| label: fig-type-timeseries
#| fig-cap: "Boat angler proportion over time by fishery (monthly)"
#| fig-height: 10
#| fig-width: 12

top_fisheries_type <- interviews |>
  filter(!is.na(angler_final)) |>
  count(fishery_name, sort = TRUE) |>
  slice_head(n = 12) |>
  pull(fishery_name)

type_monthly |>
  filter(fishery_name %in% top_fisheries_type, angler_final == "Boat") |>
  ggplot(aes(x = month, y = prop)) +
  geom_line() +
  geom_point(aes(size = total), alpha = 0.5) +
  facet_wrap(~fishery_name, scales = "free_x", ncol = 3) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    x = NULL, y = "Proportion Boat Anglers",
    size = "Sample Size",
    title = "Monthly Boat Proportion (Top 12 Fisheries by Interview Count)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Counts vs Hours: Per-Fishery Comparison

```{r}
#| label: fig-counts-vs-hours-guided
#| fig-cap: "Guided proportion: interview counts vs fishing hours (fishery level)"
#| fig-height: 7
#| fig-width: 9

# Compare fishery-level guided proportions: counts vs hours
compare_guide <- guide_fishery |>
  filter(guided_status == "Guided") |>
  select(fishery_name, prop_counts = prop) |>
  inner_join(
    guide_fishery_hrs |>
      filter(guided_status == "Guided") |>
      select(fishery_name, prop_hours = prop),
    by = "fishery_name"
  )

ggplot(compare_guide, aes(x = prop_counts, y = prop_hours)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.6) +
  scale_x_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    x = "Proportion Guided (interview counts)",
    y = "Proportion Guided (fishing hours)",
    title = "Guided Proportion: Counts vs Hours",
    subtitle = "Each point = one fishery; dashed line = 1:1"
  ) +
  theme_minimal()
```

```{r}
#| label: fig-counts-vs-hours-boat
#| fig-cap: "Boat proportion: interview counts vs fishing hours (fishery level)"
#| fig-height: 7
#| fig-width: 9

compare_type <- type_fishery |>
  filter(angler_final == "Boat") |>
  select(fishery_name, prop_counts = prop) |>
  inner_join(
    type_fishery_hrs |>
      filter(angler_final == "Boat") |>
      select(fishery_name, prop_hours = prop),
    by = "fishery_name"
  )

ggplot(compare_type, aes(x = prop_counts, y = prop_hours)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.6) +
  scale_x_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    x = "Proportion Boat (interview counts)",
    y = "Proportion Boat (fishing hours)",
    title = "Boat Proportion: Counts vs Hours",
    subtitle = "Each point = one fishery; dashed line = 1:1"
  ) +
  theme_minimal()
```

## Export: Mode × Location Proportions for Assembly

This is the hand-off to `analysis/pst/03_analysis/pst_fw_angler_trips_assembly.R`. Everything above
describes and diagnoses the proportions; this section writes the one table the
assembly script consumes.

### Grain and normalization

The export is at **`fishery_name × year × location × mode`**, and `prop` sums to
1.0 *within each `fishery_name × year × location`* — not across all four
combined categories. That is deliberate and it matters:

- **Location (bank/boat) is a design stratum.** It arrives from
  `est_pe_effort()` with its own effort estimate. The proportions must not
  re-derive it from interviews, because interview counts are not effort-weighted
  and would overwrite a design-based quantity with a convenience sample.
- **Mode (guided/unguided) is post-stratification *within* the location
  stratum.** So the quantity the assembly needs is
  $\hat{p}(\text{guided} \mid \text{boat}, \text{fishery}, \text{year})$,
  not $\hat{p}(\text{guided-boat})$.

Multiplying a design-based bank/boat effort split by an unconditional
four-category proportion would double-count the location dimension. The
conditional form is the only one that composes correctly.

Case is normalized to lowercase on both fields, because `angler_final` is
`"Bank"`/`"Boat"` in the interview data and `"bank"`/`"boat"` in
`multi_fishery_creel_trips.csv`. An unnormalized join silently drops every row.

```{r}
#| label: export-assembly-props

assembly_dir <- here::here("analysis", "pst", "outputs")
if (!dir.exists(assembly_dir)) dir.create(assembly_dir, recursive = TRUE)

mode_location_props <- interviews |>
  filter(!is.na(guided_status), !is.na(angler_final), !is.na(year)) |>
  mutate(
    location = str_to_lower(angler_final),
    mode     = str_to_lower(guided_status)
  ) |>
  count(fishery_name, year, location, mode, name = "n_interviews") |>
  # Both modes must exist in every fishery-year-location cell so that a
  # structural zero (a fishery with no guided effort) is an explicit 0 rather
  # than an absent row the assembly would read as "no proportion available".
  complete(
    nesting(fishery_name, year, location), mode = c("guided", "unguided"),
    fill = list(n_interviews = 0L)
  ) |>
  group_by(fishery_name, year, location) |>
  mutate(
    n_location = sum(n_interviews),
    prop       = if_else(n_location > 0, n_interviews / n_location, NA_real_)
  ) |>
  ungroup()

# Hours-weighted companion, carried as a diagnostic rather than the primary.
# Counts are the primary because the assembly multiplies a *trip* total; an
# hours-weighted proportion answers a different question (share of angler-hours,
# not share of trips) and would bias toward whichever mode fishes longer.
mode_location_props_hrs <- interviews |>
  filter(!is.na(guided_status), !is.na(angler_final), !is.na(year),
         !is.na(fishing_time_total)) |>
  mutate(location = str_to_lower(angler_final),
         mode     = str_to_lower(guided_status)) |>
  group_by(fishery_name, year, location, mode) |>
  summarise(hrs = sum(fishing_time_total), .groups = "drop_last") |>
  group_by(fishery_name, year, location) |>
  mutate(prop_hrs = hrs / sum(hrs)) |>
  ungroup() |>
  select(fishery_name, year, location, mode, prop_hrs)

mode_location_props <- mode_location_props |>
  left_join(mode_location_props_hrs, by = c("fishery_name", "year", "location", "mode")) |>
  mutate(prop_delta = prop - prop_hrs) |>
  arrange(fishery_name, year, location, mode)

write_csv(mode_location_props,
          file.path(assembly_dir, "interview_mode_location_props.csv"))

cat(glue("Wrote {nrow(mode_location_props)} rows covering ",
         "{n_distinct(mode_location_props$fishery_name)} fisheries × ",
         "{n_distinct(mode_location_props$year)} years\n\n"))

# Sanity: every cell with support must sum to 1
chk <- mode_location_props |>
  filter(n_location > 0) |>
  group_by(fishery_name, year, location) |>
  summarise(s = sum(prop), .groups = "drop") |>
  filter(abs(s - 1) > 1e-8)
if (nrow(chk) > 0) {
  warning(glue("{nrow(chk)} fishery-year-location cells do not sum to 1"))
} else {
  cat("All supported cells sum to 1.0\n")
}
```

### Cells resting on thin support

`pst_fw_angler_trips_assembly.R` falls back to block-pooled proportions below
`MIN_INTERVIEWS = 30`. This table is where that threshold actually bites, so it
should be read before accepting the deliverable — a fishery appearing here is
one whose mode split is borrowed, not measured.

```{r}
#| label: tbl-thin-support

thin_support <- mode_location_props |>
  distinct(fishery_name, year, location, n_location) |>
  mutate(
    support = case_when(
      n_location == 0  ~ "none — no proportion available",
      n_location < 30  ~ "thin — assembly falls back to block pooled",
      n_location < 100 ~ "moderate",
      TRUE             ~ "adequate"
    )
  )

thin_support |> count(support) |> print()

thin_support |>
  filter(n_location < 30) |>
  arrange(n_location) |>
  datatable(
    caption = "Fishery-year-location cells below the fallback threshold",
    filter = "top", rownames = FALSE,
    options = list(pageLength = 15, scrollX = TRUE)
  )
```

## Cross-Check: Interviews vs. Batch Creel Estimates

The interview pull and `multi_fishery_creel_summary.R` draw on the same creel
database but travel different paths — the batch script runs interviews through
`prep_dwg_interview_fishing_time()` and the PE estimators, while this document
reads them directly. Where the two disagree, one of them is wrong, and it is
worth knowing which before either feeds the deliverable.

Three checks, in increasing order of what they'd cost us if they failed.

```{r}
#| label: crosscheck-load

# multi_fishery_trip_summary.csv was superseded by multi_fishery_creel_summary.R,
# whose trips output is multi_fishery_creel_trips.csv; the old file now lives
# at analysis/outputs/archive/multi_fishery_trip_summary_superseded_20260817.csv.
# No repo-root fallback here: a silent path fallback changes behavior when the
# primary path breaks, which is worse than failing loudly.
batch_path <- here::here("analysis", "pst", "outputs", "multi_fishery_creel_trips.csv")

batch <- if (file.exists(batch_path)) {
  read_csv(batch_path, show_col_types = FALSE) |>
    mutate(location = str_to_lower(angler_final))
} else {
  cat("Batch creel output not found — cross-check skipped.\n")
  NULL
}

int_monthly <- interviews |>
  filter(!is.na(angler_final), !is.na(year)) |>
  mutate(location  = str_to_lower(angler_final),
         month_num = month(event_date)) |>
  group_by(fishery_name, year, month = month_num, location) |>
  summarise(
    n_int          = n(),
    n_int_complete = sum(trip_status == "Complete", na.rm = TRUE),
    mean_hrs_int   = mean(fishing_time, na.rm = TRUE),
    .groups = "drop"
  )
```

### Check 1 — Coverage: which side is missing strata

A fishery-year-location with creel effort but no interviews gets a borrowed mode
split. One with interviews but no effort estimate is either a batch failure or a
fishery outside the trip-summary scope. Both are worth naming explicitly.

```{r}
#| label: crosscheck-coverage

if (!is.null(batch)) {
  batch_cells <- batch |>
    filter(!is.na(total_trips_est)) |>
    distinct(fishery_name, year, location)
  int_cells <- int_monthly |> distinct(fishery_name, year, location)

  coverage_cmp <- full_join(
    batch_cells |> mutate(in_batch = TRUE),
    int_cells   |> mutate(in_interviews = TRUE),
    by = c("fishery_name", "year", "location")
  ) |>
    replace_na(list(in_batch = FALSE, in_interviews = FALSE)) |>
    mutate(status = case_when(
      in_batch & in_interviews  ~ "both",
      in_batch & !in_interviews ~ "effort, no interviews — mode will be borrowed",
      TRUE                      ~ "interviews, no effort — batch gap or out of scope"
    ))

  coverage_cmp |> count(status) |> print()

  coverage_cmp |>
    filter(status != "both") |>
    arrange(status, fishery_name, year) |>
    datatable(caption = "Strata present on only one side",
              filter = "top", rownames = FALSE,
              options = list(pageLength = 15, scrollX = TRUE))
}
```

### Check 2 — Interview counts

`n_completed_angler_trips` in the batch output is the interview support behind
each month's `mean_trip_length`. It should track the completed-trip count here.
A large gap means the two paths are filtering interviews differently — most
likely on `trip_status` or on a fishing-time validity rule — and the batch
filter is the one that governs the effort estimate.

```{r}
#| label: crosscheck-counts

if (!is.null(batch)) {
  count_cmp <- batch |>
    select(fishery_name, year, month, location, n_batch = n_completed_angler_trips) |>
    inner_join(int_monthly, by = c("fishery_name", "year", "month", "location")) |>
    mutate(
      diff     = n_batch - n_int_complete,
      pct_diff = if_else(n_int_complete > 0, 100 * diff / n_int_complete, NA_real_)
    )

  cat(glue("{nrow(count_cmp)} matched fishery-year-month-location cells\n"))
  cat(glue("Median absolute % difference: ",
           "{round(median(abs(count_cmp$pct_diff), na.rm = TRUE), 1)}%\n"))
  cat(glue("Cells differing by more than 20%: ",
           "{sum(abs(count_cmp$pct_diff) > 20, na.rm = TRUE)}\n\n"))

  count_cmp |>
    filter(abs(pct_diff) > 20) |>
    arrange(desc(abs(pct_diff))) |>
    datatable(caption = "Interview count disagreements over 20%",
              filter = "top", rownames = FALSE,
              options = list(pageLength = 10, scrollX = TRUE))
}
```

### Check 3 — Mean trip length, and the units question

This is the one that matters most. `total_trips_est = total_effort_hrs ÷
mean_trip_length` presumes effort is in **angler-hours** and trip length is
**elapsed hours per trip, independent of party size**. If effort were in
boat-hours or party-hours, the division would return boats rather than anglers
and every downstream number would be wrong by roughly the mean group size.

`mean_trip_length` from the batch and `mean(fishing_time)` here are computed
from the same interviews by different code, and neither multiplies by
`angler_count`. So they should agree closely. If instead the batch value runs
near a *multiple* of the interview value — and that multiple tracks
`mean_group_size` — the elapsed-vs-angler-hour distinction has collapsed
somewhere in the pipeline.

```{r}
#| label: crosscheck-trip-length

if (!is.null(batch)) {
  tl_cmp <- batch |>
    select(fishery_name, year, month, location,
           mean_trip_length, mean_group_size, total_effort_hrs) |>
    inner_join(int_monthly, by = c("fishery_name", "year", "month", "location")) |>
    filter(!is.na(mean_trip_length), !is.na(mean_hrs_int), mean_hrs_int > 0) |>
    mutate(
      ratio          = mean_trip_length / mean_hrs_int,
      ratio_vs_group = ratio / mean_group_size
    )

  cat(glue("Median batch/interview trip-length ratio: ",
           "{round(median(tl_cmp$ratio, na.rm = TRUE), 3)}  (expect ~1.00)\n"))
  cat(glue("Median ratio ÷ mean_group_size: ",
           "{round(median(tl_cmp$ratio_vs_group, na.rm = TRUE), 3)}  ",
           "(a value near 1.00 here instead would indicate party-hour contamination)\n\n"))

  ggplot(tl_cmp, aes(mean_hrs_int, mean_trip_length)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
    geom_point(aes(colour = location), alpha = 0.6) +
    labs(
      title    = "Mean trip length: batch PE vs. direct interview calculation",
      subtitle = "Dashed line is 1:1. Systematic departure indicates a units or filtering difference.",
      x = "Interview mean fishing_time (hrs)",
      y = "Batch mean_trip_length (hrs)",
      colour = "Location"
    ) +
    theme_minimal()
}
```

```{r}
#| label: crosscheck-export

if (!is.null(batch)) {
  crosscheck <- coverage_cmp |>
    left_join(
      count_cmp |> select(fishery_name, year, month, location, n_batch,
                          n_int_complete, pct_diff),
      by = c("fishery_name", "year", "location"),
      relationship = "many-to-many"
    ) |>
    left_join(
      tl_cmp |> select(fishery_name, year, month, location, mean_trip_length,
                       mean_hrs_int, ratio),
      by = c("fishery_name", "year", "month", "location")
    )

  write_csv(crosscheck,
            file.path(assembly_dir, "interview_batch_crosscheck.csv"))
  cat(glue("Wrote crosscheck: {nrow(crosscheck)} rows\n"))
}
```

::: {.callout-note}
**Deferred to the parent document.** The analytical framework — the P1/P2/P3
tier hierarchy, Track A vs. Track B, assumptions A1–A4, and the rule that ratios
do not travel between blocks — is described here only as far as it explains what
this document produces. That parent document, covering methods across all
four regional blocks, now exists at `analysis/pst/pst_fw_angler_trips.qmd`,
with this file reduced to the Track B derivation it actually performs.
:::
