<!-- _20_track_a_assembly.qmd — Track A tiering as executed: P2 block-ratio extrapolation mechanics, its leave-one-out validation, and the CRC-vs-creel bias handle that justifies the ratio's denominator choice. -->

## P2 extrapolation in practice

P2 covers **CRC areas** with no creel survey. Everything operates at
`catch_area_code`, not river: rivers span several areas (Quillayute is
398|400|402|404|406), so a river-keyed ratio cannot be applied to a CRC row at
all. Roll-up to river happens at the end.

The estimator is a **ratio of sums within a block** — P1 creel trips over CRC
harvest for the areas that have both — applied to the CRC harvest of areas
that don't:

$$
\hat T_{a,y} \;=\; H^{CRC}_{a,y} \times \frac{\sum_{d \in D_{b,y}} T^{creel}_{d,y}}{\sum_{d \in D_{b,y}} H^{CRC}_{d,y}}
$$

where $D_{b,y}$ is the donor set: areas in block $b$ with **both** a
design-based trip estimate and a CRC harvest figure in year $y$.

**The denominator is CRC, deliberately.** The assembly script also computes a
creel-denominated trips-per-salmon ratio (2.06–4.39 by block-year), and that
figure is the right one for describing fishery behaviour. It is the wrong one
for this expansion: applying it to a CRC numerator would assume CRC and creel
harvest agree, when the run's own bias check puts the median at 0.77. Matching
the donor denominator to the target measure keeps that bias inside the ratio
instead of stacking it on top.

Ratio of sums rather than mean of area ratios — the sum form weights donors by
harvest, so a small area with a thin denominator can't swing the block.
Per-area ratios are still computed, but only as a coherence diagnostic.

Guardrails, each of which sends an area-year to the gap register rather than
producing a number: minimum two donor areas, a donor-harvest floor of 100
salmon, a plausibility band on the ratio, and a cap on the within-block CV of
donor ratios. Year-specific ratios are preferred; a block-pooled ratio is the
only fallback, labeled in `ratio_basis` per R5.

The CRC-stream-to-river crosswalk that previously blocked this step already
exists: `crc_areas` in the registry crosswalk maps area → river → block, and
the bias check confirms CRC `stream_code` and creel `catch_area_code` share a
coding scheme. What genuinely remains is narrower — CRC stream_codes present in
the harvest file but in no crosswalk row. Those are unassignable and are logged,
not guessed.

`build_block_ratios()` derives **observed trips-per-salmon from the paired
merged creel outputs** — both numerator and denominator from the same
`est_pe_effort()` / `est_pe_catch()` run on the same strata, so the ratio is
internally consistent rather than two independent estimates divided. Most
recently observed block-year ratios:

| Block | 2022 | 2023 | 2024 | 2025 |
|---|---|---|---|---|
| Puget Sound | 4.24 | 1.45 | 3.15 | 1.35 |
| WA Coast | 1.98 | 2.53 | 1.93 | 2.19 |
| Columbia tribs | 2.76 | 1.89 | 2.11 | 1.71 |

Puget Sound odd years sit far below even years — the pink-salmon signal the
NEPA workbook encodes as its 3.44/8.65 odd/even split. Direction matches
independently, which is a useful check that the pairing is right. Magnitudes
differ from NEPA's because these are creel-observed totals rather than
CRC-expanded.

**What still needs wiring:** applying those ratios to uncovered river-years.
Given 127 CRC areas carry harvest with no creel behind them, this remains a
large share of the eventual deliverable — logged as a `blocker`, not guessed.

```{r}
#| label: p2-ratios

p2_ratios <- read_if(file.path(OUT_DIR, "pst_fw_p2_area_ratios.csv"))

if (is.null(p2_ratios)) {
  cat("P2 ratio table not found. Run analysis/pst/03_analysis/pst_fw_effort_assembly.R with the P2 module wired in.\n")
} else {
  p2_ratios |>
    filter(ratio_basis == "block_year") |>
    select(block, year, n_donor_rivers, donor_harvest, ratio, donor_ratio_cv, usable, fail_reason) |>
    arrange(block, year) |>
    gt() |>
    fmt_number(columns = c(ratio, donor_ratio_cv), decimals = 3) |>
    fmt_number(columns = donor_harvest, decimals = 0) |>
    tab_header(
      title = "Block-year trips-per-salmon ratios",
      subtitle = "Estimated from creel-covered donor rivers; applied to uncovered CRC rivers in the same block"
    )
}
```

### How wrong is it likely to be?

Rather than asking anyone to accept the extrapolation on structural grounds,
the module runs a **leave-one-out check**: each donor river is held out in
turn, the block ratio is rebuilt without it, and its trips are predicted as
if it were uncovered. Comparing that prediction to its actual design-based
estimate gives an empirical error band per block.

Median absolute percent error is the number to quote when asked how much
confidence P2 deserves. The signed median is worth reading alongside it — a
block with modest scatter but consistent one-directional bias is a different
problem from one with large symmetric error.

```{r}
#| label: p2-loo

p2_loo <- read_if(file.path(OUT_DIR, "pst_fw_p2_loo_summary.csv"))

if (is.null(p2_loo)) {
  cat("P2 leave-one-out summary not found.\n")
} else {
  p2_loo |>
    gt() |>
    fmt_number(columns = c(median_ape, p90_ape, bias_pct), decimals = 1) |>
    cols_label(
      n_tests = "Held-out tests",
      median_ape = "Median |error| %",
      p90_ape = "90th pct |error| %",
      bias_pct = "Median signed error %"
    ) |>
    tab_header(
      title = "P2 leave-one-out validation by block",
      subtitle = "Each donor river predicted from the remaining donors in its block"
    )
}
```

::: {.callout-warning}
**Two biases to keep visible.** The numerator is creel trips for a *surveyed
season*; the denominator is CRC harvest for the *calendar year*. Where the
creel season is materially shorter than the CRC window, the ratio is biased
low and P2 understates trips. Separately, a block ratio built on
coho-dominated donors and applied to a Chinook-dominated river transports the
wrong catch rate — holding species mix roughly constant is much of what the
block definition is doing, and why R4 forbids crossing them.

P2 produces a **total only**. Mode and location are `"unknown"` until Track B
supplies a block-level conditional proportion; where none exists they stay
unknown in the deliverable, which is a correct result rather than an
incomplete one.
:::

## CRC vs. creel: the reporting-bias handle

Both sources estimate the same quantity — total salmon kept in a CRC area-year —
by unrelated means. Creel is design-based and observed on the ground; CRC is an
angler-reported card expansion. `catch_area_code` is carried through the whole
pipeline rather than collapsed at the river roll-up, because it is the only key
that joins creel effort back to CRC harvest — rivers can span several CRC areas
(Quillayute is 398|400|402|404|406), so rolling to river grain destroys that key
one-way. Two roll-ups are written: river grain for the consultant's requested
shape, area grain for this comparison. Their ratio is the empirical handle on
reporting bias, and it is computable only because `catch_area_code` survives to
the end of the pipeline instead of being collapsed to river early.

Current run: **median CRC/creel = 0.77 across 62 comparable area-years**, IQR
0.52–1.93. CRC reports less harvest than the design-based estimate at the
median, but the spread is wide enough that the direction is not stable per
area — another instance of R4: the bias doesn't travel between systems any
better than the ratios do.

This is what forces the P2 estimator to be **CRC-denominated**. A ratio built as
creel-trips / creel-harvest and then applied to CRC harvest would inherit that
bias as a systematic understatement. Building it as creel-trips / CRC-harvest
puts the bias inside the ratio, where it cancels against the CRC numerator it
is applied to.

::: {.callout-note}
Basis differs — CRC is license-year (Apr 1 – Mar 31), creel is calendar-month.
The comparison is indicative at annual grain and is not a month-for-month
reconciliation. 2025 pairs return zero CRC harvest because CRC publishes
through license year 2024 only; those rows join but are flagged
`comparable = FALSE` and excluded from the median, so a coverage artifact
isn't misread as a reporting collapse.
:::

::: {.callout-warning}
**Figure discrepancy, not yet reconciled.** An earlier documentation layer
(`PST_ANALYSIS.md` §8.6, being retired into this document) reported **median
CRC/creel ≈ 0.60 across 74 paired area-years**, and illustrated the
non-uniformity with two examples: area 618 (Drano) running *above* 1.0 in
2023–2024 while area 358 (Humptulips) sat near 0.35. The figure carried
forward above — **0.77 across 62 comparable area-years** — comes from the
current parent document and from `pst_p2_block_ratio.R`, and is treated here
as the newer of the two, but the two runs disagree on both the ratio (0.60 vs
0.77) and the pair count (74 vs 62). That gap is large enough to matter for
anyone using this number to correct P2, and it has not been reconciled. Do
not treat either figure as settled until it is checked against a fresh
assembly run.
:::

This matters for P2 directly: if CRC is the harvest denominator for
uncovered river-years but runs meaningfully below the design-based truth,
then expanding CRC harvest by a creel-derived trips-per-salmon ratio
inherits that bias. Whether to correct for it is a decision for Jim, not
something to bury in the pipeline.

```{r}
#| label: crc-bias

if (!is.null(crc_bias)) {
  crc_bias |>
    filter(comparable) |>
    summarise(
      n_area_years = n(),
      median_ratio = median(crc_over_creel, na.rm = TRUE),
      q25          = quantile(crc_over_creel, 0.25, na.rm = TRUE),
      q75          = quantile(crc_over_creel, 0.75, na.rm = TRUE)
    ) |>
    gt() |>
    fmt_number(columns = c(median_ratio, q25, q75), decimals = 2) |>
    tab_header(title = "CRC harvest / creel harvest, same area-year")
}
```
