<!-- _20_track_a_assembly.qmd — Track A tiering as executed: P2 block-ratio extrapolation mechanics, its leave-one-out validation, and the CRC-vs-creel bias handle that justifies the ratio's denominator choice. -->

## P2 extrapolation in practice

P2 covers **CRC areas** with no creel survey. Everything operates at
`catch_area_code`, not river: rivers span several areas (Quillayute is
398|400|402|404|406), so a river-keyed ratio cannot be applied to a CRC row at
all. Roll-up to river happens at the end.

The estimator is a **ratio of sums within a block** — P1 creel trips over CRC
harvest for the areas that have both — applied to the CRC harvest of areas
that don't:

$$
\hat T_{a,y} \;=\; H^{CRC}_{a,y} \times \frac{\sum_{d \in D_{b,y}} T^{creel}_{d,y}}{\sum_{d \in D_{b,y}} H^{CRC}_{d,y}}
$$

where $D_{b,y}$ is the donor set: areas in block $b$ with **both** a
design-based trip estimate and a CRC harvest figure in year $y$.

**The denominator is CRC, deliberately.** The assembly script also computes a
creel-denominated trips-per-salmon ratio (2.06–4.39 by block-year), and that
figure is the right one for describing fishery behaviour. It is the wrong one
for this expansion: applying it to a CRC numerator would assume CRC and creel
harvest agree, when the run's own bias check puts the median at 0.77. Matching
the donor denominator to the target measure keeps that bias inside the ratio
instead of stacking it on top.

Ratio of sums rather than mean of area ratios — the sum form weights donors by
harvest, so a small area with a thin denominator can't swing the block.
Per-area ratios are still computed, but only as a coherence diagnostic.

Guardrails, each of which sends an area-year to the gap register rather than
producing a number: minimum two donor areas, a donor-harvest floor of 100
salmon, a plausibility band on the ratio, and a cap on the within-block CV of
donor ratios. Year-specific ratios are preferred; a block-pooled ratio is the
only fallback, labeled in `ratio_basis` per R5.

The CRC-stream-to-river crosswalk that previously blocked this step already
exists: `crc_areas` in the registry crosswalk maps area → river → block, and
the bias check confirms CRC `stream_code` and creel `catch_area_code` share a
coding scheme. What genuinely remains is narrower — CRC stream_codes present in
the harvest file but in no crosswalk row. Those are unassignable and are logged,
not guessed.

`build_block_ratios()` derives **observed trips-per-salmon from the paired
merged creel outputs** — both numerator and denominator from the same
`est_pe_effort()` / `est_pe_catch()` run on the same strata, so the ratio is
internally consistent rather than two independent estimates divided. Most
recently observed block-year ratios:

| Block | 2022 | 2023 | 2024 | 2025 |
|---|---|---|---|---|
| Puget Sound | 4.24 | 1.45 | 3.15 | 1.35 |
| WA Coast | 1.98 | 2.53 | 1.93 | 2.19 |
| Columbia tribs | 2.76 | 1.89 | 2.11 | 1.71 |

Puget Sound odd years sit far below even years — the pink-salmon signal the
NEPA workbook encodes as its 3.44/8.65 odd/even split. Direction matches
independently, which is a useful check that the pairing is right. Magnitudes
differ from NEPA's because these are creel-observed totals rather than
CRC-expanded.

**What still needs wiring:** applying those ratios to uncovered river-years.
Given 127 CRC areas carry harvest with no creel behind them, this remains a
large share of the eventual deliverable — logged as a `blocker`, not guessed.

```{r}
#| label: p2-ratios

p2_ratios <- read_if(file.path(OUT_DIR, "pst_fw_p2_area_ratios.csv"))

if (is.null(p2_ratios)) {
  cat("P2 ratio table not found. Run analysis/pst/03_analysis/pst_fw_angler_trips_assembly.R with the P2 module wired in.\n")
} else {
  p2_ratios |>
    filter(ratio_basis == "block_year") |>
    select(block, year, n_donor_areas, donor_crc, ratio, donor_ratio_cv, usable, fail_reason) |>
    arrange(block, year) |>
    gt() |>
    fmt_number(columns = c(ratio, donor_ratio_cv), decimals = 3) |>
    fmt_number(columns = donor_crc, decimals = 0) |>
    tab_header(
      title = "Block-year trips-per-salmon ratios",
      subtitle = "Estimated from creel-covered donor rivers; applied to uncovered CRC rivers in the same block"
    )
}
```

### How wrong is it likely to be?

Rather than asking anyone to accept the extrapolation on structural grounds,
the module runs a **leave-one-out check**: each donor river is held out in
turn, the block ratio is rebuilt without it, and its trips are predicted as
if it were uncovered. Comparing that prediction to its actual design-based
estimate gives an empirical error band per block.

Median absolute percent error is the number to quote when asked how much
confidence P2 deserves. The signed median is worth reading alongside it — a
block with modest scatter but consistent one-directional bias is a different
problem from one with large symmetric error.

```{r}
#| label: p2-loo

p2_loo <- read_if(file.path(OUT_DIR, "pst_fw_p2_loo_summary.csv"))

if (is.null(p2_loo)) {
  cat("P2 leave-one-out summary not found.\n")
} else {
  p2_loo |>
    gt() |>
    fmt_number(columns = c(median_ape, p90_ape, bias_pct), decimals = 1) |>
    cols_label(
      n_tests = "Held-out tests",
      median_ape = "Median |error| %",
      p90_ape = "90th pct |error| %",
      bias_pct = "Median signed error %"
    ) |>
    tab_header(
      title = "P2 leave-one-out validation by block",
      subtitle = "Each donor river predicted from the remaining donors in its block"
    )
}
```

::: {.callout-note}
**The ratio's own construction is already season-matched.** An earlier version
of this document warned that the numerator (creel trips for a surveyed
season) and denominator (CRC harvest for the calendar year) didn't line up,
biasing the ratio low. That's no longer how the donor side works:
`build_p2_donors()` restricts CRC harvest to the SAME calendar months the
creel survey actually covered, per donor area-year — the same construction
now used for the CRC-vs-creel bias comparison above. Both sides of the ratio
are measured over the same window.
:::

::: {.callout-warning}
**What's still imperfect, and deliberately so: applying that ratio to a
target's full-year CRC harvest.** The ratio is calibrated on the donor's own
season; the target (by definition, uncovered by any creel program) gets that
ratio applied against its ENTIRE calendar year of CRC harvest, not just the
donor's matched months. Restricting the target side to the donor's months
was considered and rejected: CRC harvest can only occur when a river's
fishery is actually open, so a target's "full-year" total is already
concentrated in whatever months that river's own season runs — which we have
no independent way to know matches the donor's timing. Discarding target
harvest outside the donor's window wouldn't correct for a timing mismatch,
it would just arbitrarily drop real harvest the ratio was never given a
chance to explain, making the estimate more incomplete without making it
more accurate. The risk that remains is narrower and unavoidable: the
donor-calibrated trips-per-fish RATE may not transfer perfectly to however
the target's own season is timed, regulated, or fished — the same kind of
risk as the species-mix caveat below, not a fixable bookkeeping error.
LOO (below) cannot catch this either way, since every donor shares whatever
timing they have in common.

Separately: a block ratio built on coho-dominated donors and applied to a
Chinook-dominated river transports the wrong catch rate — holding species
mix roughly constant is much of what the block definition is doing, and why
R4 forbids crossing them.

P2 produces a **total only**. Mode and location are `"unknown"` until Track B
supplies a block-level conditional proportion; where none exists they stay
unknown in the deliverable, which is a correct result rather than an
incomplete one.
:::

## CRC vs. creel: the reporting-bias handle

Both sources estimate the same quantity — total salmon kept in a CRC area-year —
by unrelated means. Creel is design-based and observed on the ground; CRC is an
angler-reported card expansion. `catch_area_code` is carried through the whole
pipeline rather than collapsed at the river roll-up, because it is the only key
that joins creel effort back to CRC harvest — rivers can span several CRC areas
(Quillayute is 398|400|402|404|406), so rolling to river grain destroys that key
one-way. Two roll-ups are written: river grain for the consultant's requested
shape, area grain for this comparison. Their ratio is the empirical handle on
reporting bias, and it is computable only because `catch_area_code` survives to
the end of the pipeline instead of being collapsed to river early.

**Season-matched as of 2026-08-21.** CRC harvest in this comparison is now
restricted to the same calendar months the creel survey actually covered for
that area-year — the identical construction `build_p2_donors()` uses for the
P2 ratio's own denominator (see the P2 section above). Previously this
compared creel harvest for a surveyed season against CRC's FULL calendar
year for the same area-year, which conflated two different things: genuine
CRC-vs-creel reporting bias, and the mechanical fact that CRC counts months
nobody was creeling at all. The full-year CRC total is retained in the
output as `crc_harvest_full_year` so the size of that season-coverage gap
stays visible, but it is no longer part of the ratio itself.

The **0.77 across 62 comparable area-years** figure quoted in earlier
versions of this document was computed under the old, full-year methodology
and is stale — it will change once the pipeline is re-run under the fix
above, and the new number is not yet in hand. Until a fresh run produces it,
treat any specific CRC/creel ratio figure in this document as provisional.

This is what forces the P2 estimator to be **CRC-denominated**. A ratio built as
creel-trips / creel-harvest and then applied to CRC harvest would inherit that
bias as a systematic understatement. Building it as creel-trips / CRC-harvest
puts the bias inside the ratio, where it cancels against the CRC numerator it
is applied to.

::: {.callout-note}
License-year vs. calendar-year is NOT a live concern here — resolved already,
and at the finest possible grain: `parse_crc_freshwater_harvest.R` converts
every monthly cell from the raw license-year layout (Apr–Dec of
`license_year`, Jan–Mar of `license_year + 1`) to a genuine `calendar_year`
and `calendar_month`, so both sides of this comparison already share real
calendar time. 2025 pairs return zero matched-month CRC harvest because CRC
publishes through license year 2024 only (calendar 2025 has just Jan–Mar
coverage); those rows join but are flagged `comparable = FALSE` and excluded
from the median, so a coverage artifact isn't misread as a reporting
collapse.
:::

::: {.callout-warning}
**Figure discrepancy, superseded by the methodology change above, not yet
re-measured.** An earlier documentation layer (`PST_ANALYSIS.md` §8.6, being
retired into this document) reported median CRC/creel ≈ 0.60 across 74
paired area-years; a later run of the full-year methodology reported 0.77
across 62. Neither figure used the season-matched construction now in code,
so reconciling the two against each other is moot — both need to be
superseded by a fresh run under the current methodology, not compared to one
another.
:::

This matters for P2 directly: if CRC is the harvest denominator for
uncovered river-years but runs meaningfully below the design-based truth,
then expanding CRC harvest by a creel-derived trips-per-salmon ratio
inherits that bias. Whether to correct for it is a decision for Jim, not
something to bury in the pipeline.

```{r}
#| label: crc-bias

if (!is.null(crc_bias)) {
  crc_bias |>
    filter(comparable) |>
    summarise(
      n_area_years         = n(),
      avg_pct_year_covered = mean(pct_year_covered_by_creel, na.rm = TRUE),
      median_ratio         = median(crc_over_creel, na.rm = TRUE),
      q25                  = quantile(crc_over_creel, 0.25, na.rm = TRUE),
      q75                  = quantile(crc_over_creel, 0.75, na.rm = TRUE)
    ) |>
    gt() |>
    fmt_number(columns = c(median_ratio, q25, q75), decimals = 2) |>
    fmt_number(columns = avg_pct_year_covered, decimals = 1) |>
    cols_label(avg_pct_year_covered = "Avg % of year covered by creel") |>
    tab_header(
      title = "CRC harvest / creel harvest, same area-year AND same matched creel months",
      subtitle = "crc_harvest_full_year (in the CSV) carries the unrestricted total for comparison, not used in the ratio"
    )
}
```
