<!-- _21_track_b_proportions.qmd — Track B results as committed to the repo: the mode x location proportion table and its cross-check against batch-level records. Derivation logic lives in the producer, not here. -->

## Track B results: mode × location proportions

The framework for this track — bank/boat as a design stratum, guided/unguided
as a post-stratified conditional proportion within the boat stratum, and why
the two are exported as a conditional $P(\text{guided}\mid\text{boat})$ rather
than a joint four-way split — is described in Track B — mode × location
above. This section presents the results that framework produces, not the
framework itself.

**The derivation does not run here.** It happens in the standalone producer
`analysis/pst/02_ingest/interview_proportions.qmd`, which queries the
interview database directly and is **not** included by this parent — it needs
live DB access that a rendered deliverable should not depend on. What follows
reads only the committed artifacts that producer writes to
`INTERVIEW_PROPS_DIR` (`analysis/pst/outputs/04_interview_proportions/`).

As of this writing, `INTERVIEW_PROPS_DIR` holds two of the producer's outputs:
`interview_mode_location_props.csv` and `interview_batch_crosscheck.csv`. The
producer also writes roughly two dozen additional per-fishery proportion and
variability tables — guided/type proportions and their variability summaries
at fishery, monthly, and weekly grain, both interview-count-weighted and
hours-weighted (`guide_proportions_fishery.csv`,
`combined_proportions_weekly_hrs.csv`, `guide_variability_summary.csv`, and
similar). Those were previously written to a repo-root `output/` directory
that was never committed; `output_dir` in the producer now resolves to this
same `INTERVIEW_PROPS_DIR`, but the producer has not been re-run since that
path fix, so those files are not yet present here. Every chunk below
therefore degrades through `read_if()` exactly like the rest of this
document.

```{r}
#| label: trackb-mode-location-props

props <- read_if(file.path(INTERVIEW_PROPS_DIR, "interview_mode_location_props.csv"))

if (is.null(props)) {
  cat("Track B proportions table not found. Run analysis/pst/02_ingest/interview_proportions.qmd.\n")
} else {
  show_cols <- intersect(
    c("fishery_name", "year", "location", "mode", "n_interviews", "n_location",
      "n_anglers", "n_anglers_location", "prop", "prop_basis", "prop_parties",
      "prop_hrs", "prop_delta", "prop_party_delta"),
    names(props)
  )
  num_cols <- intersect(c("prop", "prop_parties", "prop_hrs", "prop_delta",
                          "prop_party_delta"), show_cols)

  props |>
    arrange(fishery_name, year, location, mode) |>
    select(all_of(show_cols)) |>
    gt() |>
    fmt_number(columns = all_of(num_cols), decimals = 3) |>
    tab_header(
      title = "Interview-derived mode proportions",
      subtitle = "Grain: fishery_name x year x location x mode. prop is the angler-weighted share; prop_parties is the same split by interview (party) counts; prop_hrs weights by interview-reported hours; prop_delta is prop - prop_hrs and prop_party_delta is prop - prop_parties"
    )
}
```

An interview row is a fishing **party**, not a single angler, and the total
this proportion multiplies in the assembly is an angler-trip total. So `prop`
is the share of **anglers** at a location classified guided, not the share of
interviews — the two differ by the ratio of mean party sizes, and because
guided parties run larger than unguided ones, the party-count version
understates the guided share. `prop_parties` carries that older count-based
figure and `prop_party_delta` the size of the correction; `prop_basis` flags
the cells where `angler_count` coverage was too thin to weight by anglers at
all and the party share stands in.

`prop_hrs` re-weights the same split by hours reported per interview, since a
guided trip and an unguided trip are not sampled at equal effort-intensity.
`prop_delta` is provided so a location-year where it and `prop` diverge
sharply is visible without recomputing it — a large delta says the trip-based
and effort-based stories disagree, which is worth a second look before the
proportion is applied to a Track A total.

Note what this table does **not** contain: no row here carries
`mode = "unknown"`. That value is real in the assembled deliverable (see
Status & gaps), but it is a Track A/assembly-time artifact — rivers where the
guided field was never collected (all mid-Columbia R3 programs) or where the
interview sample fell below `MIN_INTERVIEWS` never reach this table at all,
because there is no interview signal to compute a proportion from. An
"unknown" share that will not go to zero on its own is the single number
that gates delivery, and treating it as answered by this table would be a
mistake — it is unresolved, not resolved, and needs either a documented
donor decision from Jim or an explicit unclassified category agreed with
Melissa.

### Cross-check against batch-level records

`interview_batch_crosscheck.csv` joins the interview records behind the
table above against the independently produced batch effort file, at
`fishery_name x year x location x month` grain, to catch two failure modes:
interview coverage claimed where no batch effort record exists, and a
trip-length figure computed from interviews (`mean_hrs_int`) diverging sharply
from the one computed from the batch file (`mean_trip_length`). The raw file
carries exact duplicate rows for some cells (an artifact of the upstream
join, not a QA finding), so the chunk below de-duplicates before summarizing.

```{r}
#| label: trackb-batch-crosscheck

crosscheck <- read_if(file.path(INTERVIEW_PROPS_DIR, "interview_batch_crosscheck.csv"))

if (is.null(crosscheck)) {
  cat("Interview/batch crosscheck table not found.\n")
} else {
  crosscheck |>
    distinct() |>
    group_by(status) |>
    summarise(
      cells           = n(),
      median_pct_diff = median(pct_diff, na.rm = TRUE),
      median_ratio    = median(ratio, na.rm = TRUE),
      .groups = "drop"
    ) |>
    gt() |>
    fmt_number(columns = c(median_pct_diff, median_ratio), decimals = 2) |>
    tab_header(
      title = "Batch-record vs. interview-record coverage, by month x location cell",
      subtitle = "ratio = mean_trip_length (batch) / mean_hrs_int (interview); pct_diff = 100 x (n_int_complete - n_batch) / n_batch"
    )
}
```

A `status` of `"interviews, no effort — batch gap or out of scope"` marks a
cell where interviews exist but the batch effort file has nothing to compare
against — either a genuine batch gap or a month/location combination outside
that fishery's surveyed scope. It is not, by itself, evidence of a defect;
it is a worklist for whoever reconciles the two files next.
