<!-- _22_status_and_gaps.qmd — where the work actually stands: rows-by-block status, known blockers/defects, the coverage map placeholder, the open-gap ledger, documented deviations, and the delivery plan. Folds in PST_ANALYSIS.md §6-8 (minus §8.2/8.4/8.6, owned by the Track A/B children). -->

# Status

```{r}
#| label: status-summary

if (is.null(effort)) {
  cat("No assembly output found. Run analysis/pst_fw_angler_trips_assembly.R.\n")
} else {
  effort |>
    count(block, year, name = "rows") |>
    pivot_wider(names_from = year, values_from = rows, values_fill = 0) |>
    gt() |>
    tab_header(title = "Rows by block and year",
               subtitle = "Deliverable blocks only; ColumbiaMainstem is out of scope")
}
```

```{r}
#| label: status-mode-unknown

# The single number that gates delivery: share of trips still carrying
# mode = "unknown". Track B cannot resolve R3_external (field not collected)
# or P2 (CRC has no mode field), so this will not reach zero by itself.
if (!is.null(effort)) {
  effort |>
    group_by(block) |>
    summarise(
      angler_trips     = sum(angler_trips, na.rm = TRUE),
      pct_mode_unknown = round(100 * sum(angler_trips[mode == "unknown"]) /
                                 sum(angler_trips), 1),
      pct_tier_p2      = round(100 * sum(angler_trips[grepl("P2", tier)]) /
                                 sum(angler_trips), 1),
      .groups = "drop"
    ) |>
    gt() |>
    fmt_number(columns = angler_trips, decimals = 0) |>
    tab_header(title = "Unresolved mode share by block",
               subtitle = "Must be understood, not necessarily zero, before delivery")
}
```

## Known blockers

| # | Item | Owner | Effect if unresolved |
|---|---|---|---|
| 1 | Guided/unguided absent for all mid-Columbia (ColumbiaUpper: Hanford, Yakima, McNary) | Todd Miller | Those rivers deliver `mode = "unknown"` |
| 2 | 2025 coastal freshwater has no fallback identified | — | 2025 coastal is empty; PS covered by NEPA projection |
| 3 | Upper Columbia district (Wenatchee/Entiat/Methow/Okanogan, ColumbiaUpper) | Chad Jackson, pending since 8/11 | No P1 at all; CRC-only P2 candidates until his data lands |

Resolved by scope cut: the LCR undocumented scalar, the Bonneville–McNary
totals-only split, and the Buoy 10 Charter Boat mode question are no longer
delivery blockers — that whole block (Columbia mainstem) is out of scope; see
Purpose & deliverable contract. They'd only resurface if Jim/Melissa ever want
WDFW to independently reproduce or reconcile against the consultant's own
ODFW numbers.

## Known defects

- **`district_creel_ingestion.R`** — `Hanford Reach fall Chinook
  2025` rows carry `year = 2024`. The assembly script works around it by
  trusting the year token in `fishery_name` and logging every corrected row,
  but the ingestion script should be patched.
- **Snake River (R1) crosswalk coverage — RESOLVED (2026-08-19).** Added six
  `source_id = "R1_external"` rows (`block = "ColumbiaSnake"`, one per
  fishery-year, matching `district_creel_summary.csv`'s exact `fishery_name`
  values), `area_coverage = "covered_unpartitioned"` so P2 doesn't also try to
  ratio-expand the six Snake R. System CRC areas independently. Real trips now
  flow into the deliverable instead of resolving `block = "unknown"`.
- **`multi_fishery_creel_run_ledger.csv`** — 98 rows, all `ok` as of the
  2026-08-14 run. The previous 3 errors / 1 skip are resolved. Lower Cowlitz
  remains absent from the output entirely: it lacks the data needed for
  effort estimation, a genuine gap rather than a pipeline bug.
- **Superseded files on disk — RESOLVED (2026-08-17).** `multi_fishery_trip_summary.csv`
  was stale but still present, and had already caused one analysis to be run
  against wrong numbers; `results.csv` is no longer anywhere in the repo. The
  reorganization moved the former to
  `analysis/outputs/archive/multi_fishery_trip_summary_superseded_20260817.csv`,
  and archived the two scripts that wrote the live pipeline's filenames
  (`multi_fishery_trip_summary.R`, `multi_fishery_harvest_summary.R`) so they
  can no longer clobber `multi_fishery_creel_summary.R`'s outputs. The
  assembly script still logs a `defect` if it finds a superseded file.
- **CRC compile** — region labels lost for 2023–2024; subtotal blocks
  retained as leaf rows causing ~2× double-count. Confirm the current
  `crc_freshwater_harvest_2010_2024_tidy.csv` postdates the fix.
- **NEPA workbook** — fill-right errors in the 2022/2024 even-year columns.

## Why the assembly output is not yet the deliverable

`analysis/pst/03_analysis/pst_fw_angler_trips_assembly.R` currently answers "what
can we support right now," not "what do we owe Northern Economics." The P2
ratio/expansion gap and the CRC-vs-creel bias handle are covered in Track A —
assembly in practice above; the unresolved `mode = "unknown"` share is
covered in Track B results above. Three more stand between the two, covered
here.

### Trip expansion — resolved

An earlier version of this document reported that 29% of measured effort was
being silently dropped (Nisqually, Puyallup/Carbon). **That finding was an
artifact of reading a superseded file** (`multi_fishery_trip_summary.csv`).
Against the current `multi_fishery_creel_trips.csv`:

- All 49 fisheries expand cleanly — **zero** with `NA` trips.
- The run ledger is 98 rows, **all `ok`** — no errors, no skips.
- Nisqually 2022 *and* 2023, all four Puyallup/Carbon years, and all four
  Drano years now carry trip estimates. The `[S4]` donor hierarchy is doing
  the work: 778 rows use own-month trip length, 166 `area_season`, 66
  `type_season`, 8 `fishery_season`.
- Lower Cowlitz is absent entirely — the one genuine, confirmed data gap.

The assembly script still logs expansion failures if they reappear, but
there are none today.

::: {.callout-warning}
**The real hazard in the new file is the opposite of a drop — it's a
double-count.** `multi_fishery_creel_trips.csv` stacks both the week- and
month-stratified runs, so every stratum appears twice. Naively summing gives
**1,881,197** trips against a correct month-basis figure of **954,177**. The
assembly script filters to `PE_PERIOD = "month"` and asserts uniqueness of
the design grain afterward — but any ad hoc query against the raw file that
skips that filter will silently double every number in it.
:::

### P3 fallbacks not yet ingested

`ingest_published_fallback()` only logs notes today. The NEPA projection for
Puget Sound 2025 and any coastal 2025 fallback are not yet read.

Separately, the year-specific **historical** NEPA estimates for 2022–2025 have
not been located. The workbook in hand projects 2026/2027 and carries
historical base-year data, but not final per-season estimates for the four
years the consultant asked for. Either locate the previously-reported figures
or re-run the projection logic for those years. (Carried over from the Track B
producer's open-issues list, which is retired into this document.)

### Trips vs. angler-days — still open

See Glossary & unit conventions for the trips-vs-days distinction itself. No
conversion is applied anywhere in this pipeline, and that decision has not
been made — it belongs to Jim and Northern Economics, not to a pipeline
default.

## Coverage map {#sec-map}

::: {.callout-note}
**Placeholder.** Intent: a WA state map with river systems symbolized by
assembly status — resolved (P1), resolved by ratio (P2), published-figure
fallback (P3), and open gap — with a year facet or toggle. This is the
progress tracker for the remaining outreach, and the single artifact most
worth showing Jim in a status conversation.

Requires: a river-system spatial layer (`data/spatial/`) joined to the
registry crosswalk on its key, and the coverage ledger for the status fill.
This layer does not exist in the repo as of this writing — the map is
blocked on the data, not on the code below.
:::

```{r}
#| label: coverage-map
#| eval: false

# Sketch — not yet wired. The spatial layer this references does not exist
# in the repo yet; the path below is aspirational, not a stale reference to
# a file that moved.
# library(sf); library(leaflet)
#
# rivers_sf <- st_read(here::here("data", "spatial", "wa_river_systems.gpkg")) |>
#   left_join(crosswalk, by = c("river_name" = "crosswalk_river")) |>
#   left_join(coverage,  by = c("river", "year"))
#
# pal <- colorFactor(
#   palette = c(P1 = "#1b7837", P2 = "#7fbf7b", P3 = "#fdb863", gap = "#b2182b"),
#   domain  = rivers_sf$tier_or_gap
# )
#
# leaflet(rivers_sf) |>
#   addProviderTiles("CartoDB.Positron") |>
#   addPolylines(color = ~pal(tier_or_gap), weight = 3, label = ~river) |>
#   addLegend(pal = pal, values = ~tier_or_gap, title = "Assembly status")
```

## Open gaps

```{r}
#| label: gap-table

if (is.null(gaps)) {
  cat("Gap ledger not found.\n")
} else {
  gaps |>
    arrange(block, severity, source_id) |>
    gt() |>
    tab_header(title = "Logged gaps (R2)")
}
```

Gaps tracked outside the ledger, because they are outreach rather than code:

| Gap | Owner | Status |
|---|---|---|
| Yakima & McNary workbooks, 2023–2025 | Todd Miller (R3) | Received commitment 2026-08-12; not yet ingested |
| Upper Columbia | Chad Jackson (R2) | Outreach 2026-08-11; pending |
| Snake River | Jeremy Trump | Outreach 2026-08-11; pending |
| Green River (Green-Duwamish) creel data | Nathanael Overman | Outreach 2026-08-19; pending. Until it lands, Green-Duwamish (CRC area 746) is a P2 candidate via the crosswalk row added the same day — real P1 trips from Overman's data would supersede that ratio estimate |
| Columbia Basin empirical effort/catch 2022–2025 (to replace FEIS rates) | Mark S. / CRD | Team check pending |
| Coastal freshwater 2025 | — | No fallback identified |
| McNary 2022 | Todd Miller | Confirmed missing; no source |
| 2025 CRC harvest | Heidi / Harvest Data Team | Publishes October — binding constraint on 2025 |
| NEPA odd/even ratio misapplication | Christina Iverson | Call pending; affects 2022/2024 columns |
| Lower Cowlitz | — | Genuine data gap, not a bug; `KNOWN_FAILED` |
| Yakima 2023–2025 trip expansion | Evan / Todd | **Blocker.** 4,393 / 5,529 / 3,533 effort-hrs present, `total_trips_est` NA for every row (OQ4 — 2023+ workbooks dropped the expanded trips column). ColumbiaUpper totals are undercounts until resolved |
| Lower Chehalis 2023 NA-area shadow rows | Evan | Defect. Strata emitted twice, once coded and once `catch_area_code = NA`; dropped downstream, but fix the section→area join in `analysis/pst/02_ingest/multi_fishery_creel_summary.R` |
| Drano `angler_final = "fail"` | Evan | Defect. Failed bank/boat classification recoded `"unknown"`; fix in the Drano PE branch |
| Additivity check, 122 strata | Evan | Relative differences ~1e-4 — floating-point, not a missing species. Consider raising the tolerance so the check stops crying wolf |
| week vs. month `pe_period` | Jim / Evan | 17 fisheries differ >15% between stratifications. `PE_PERIOD` is currently `"month"`; the choice needs to be made explicitly, not defaulted |
| River-to-creel crosswalk | Evan | Raised as open in the Track B producer before the current design landed. Appears superseded: matching now keys on `catch_area_code` via `pst_river_block_crosswalk.csv`, not river name. Confirm and close, or reopen with specifics |
| Pooled proportions for uncreeled rivers | Evan | Same vintage. Appears superseded by P2 within-block ratios with leave-one-out validation — which is the regional fallback that request asked for. Confirm and close |

## Documented methodology deviations

| # | Deviation | Rationale | Scope |
|---|---|---|---|
| MD1 | Mid-Columbia and Snake River Track A totals accepted as **pre-computed** inputs (`district_creel` source_ids: `R3_external`, `R1_external`), bypassing the P1 hierarchy | Todd Miller's (R3) and Jeremy Trump's (R1) spreadsheet totals are the only expansion that exists for these systems; unvetted external calculations, not design-based PE runs | Hanford Reach, Yakima, McNary (ColumbiaUpper), Snake River (ColumbiaSnake) — all delivering as of 2026-08-19. **Not generalizable**; R2 will need the same call when it arrives. |
| MD2 | Guided/unguided coded `"unknown"` for all mid-Columbia | Field not collected by R3 programs | Per R3 |

For reference, the 2022 Hanford Reach figures accepted under MD1:
26,342 boat trips and 2,103 bank trips (92.6 / 7.4). Track B location is
therefore satisfied empirically for that river-year even though Track A is
not design-based.

# Delivery plan

| Step | Owner | Blocking |
|---|---|---|
| 1. Land manual runs (Puyallup/Carbon, Nisqually, Drano) | Evan | No |
| 2. Ingest Todd's 2023–2025 workbooks | Evan | Todd |
| 3. Resolve odd/even ratio with Christina | Evan | Christina call |
| 4. Confirm D5 (upriver spring Chinook) | Jim | Jim |
| 5. Replace FEIS rates with CRD empirical where available | Evan | Mark / CRD |
| 6. Freeze 2022–2024; flag 2025 as provisional | Evan | CRC October |
| 7. ~~Reconcile A1–A4 wording between this doc and `PST_ANALYSIS.md`~~ — done 2026-08-17; `PST_ANALYSIS.md` folded into the intro/methods children and retired | — | — |
| 8. Wire the coverage map | Evan | Spatial layer |
| 9. Recover Yakima 2023–2025 trip expansion | Evan | Todd (OQ4) |
| 10. Decide `PE_PERIOD` (week vs. month) on the record | Jim | No |

::: {.callout-warning}
**2025 cannot be complete before October.** The 2025 CRC harvest file is the
input for every P2 and P3 row in that year. Options are to deliver 2022–2024
final with 2025 following, or to deliver 2025 as projected and labeled per R5.
This should be surfaced to the consultant now rather than at delivery.
:::
