<!-- _22_status_and_gaps.qmd — where the work actually stands: rows-by-block status, known blockers/defects, the coverage map placeholder, the open-gap ledger, documented deviations, and the delivery plan. Folds in PST_ANALYSIS.md §6-8 (minus §8.2/8.4/8.6, owned by the Track A/B children). -->

# Status

```{r}
#| label: status-summary

if (is.null(effort)) {
  cat("No assembly output found. Run analysis/pst_fw_angler_trips_assembly.R.\n")
} else {
  effort |>
    count(block, year, name = "rows") |>
    pivot_wider(names_from = year, values_from = rows, values_fill = 0) |>
    gt() |>
    tab_header(title = "Rows by block and year",
               subtitle = "Deliverable blocks only; ColumbiaMainstem is out of scope")
}
```

```{r}
#| label: status-mode-unknown

# The single number that gates delivery: share of trips still carrying
# mode = "unknown". Track B cannot resolve R3_external (field not collected)
# or P2 (CRC has no mode field), so this will not reach zero by itself.
if (!is.null(effort)) {
  effort |>
    group_by(block) |>
    summarise(
      angler_trips     = sum(angler_trips, na.rm = TRUE),
      pct_mode_unknown = round(100 * sum(angler_trips[mode == "unknown"]) /
                                 sum(angler_trips), 1),
      pct_tier_p2      = round(100 * sum(angler_trips[grepl("P2", tier)]) /
                                 sum(angler_trips), 1),
      .groups = "drop"
    ) |>
    gt() |>
    fmt_number(columns = angler_trips, decimals = 0) |>
    tab_header(title = "Unresolved mode share by block",
               subtitle = "Must be understood, not necessarily zero, before delivery")
}
```

## Known blockers

| # | Item | Owner | Effect if unresolved |
|---|---|---|---|
| 1 | Guided/unguided absent for all mid-Columbia (ColumbiaUpper: Hanford, Yakima, McNary) | Todd Miller | Those rivers deliver `mode = "unknown"` |
| 2 | 2025 coastal freshwater has no fallback identified | — | WACoast 2025 rows come only from P3's 6-yr trailing mean where history allows; areas failing `min_history_years` have nothing at all. **Note this row changed meaning 2026-08-21:** it used to also say "PS covered by NEPA projection," which is no longer true — Puget Sound 2025 is now covered by our own same-parity P3 (`CRC_PROJECTION_CONTROL_PS`), not by any NEPA-derived figure |
| 3 | ~~Upper Columbia district (Wenatchee/Entiat/Methow/Okanogan, ColumbiaUpper)~~ | ~~Chad Jackson, pending since 8/11~~ | **Resolved 2026-08-28.** Chad Jackson replied with a single combined district-wide angler-trip total (2022-2025), ingested via the new `R2_external` source — see the "Upper Columbia" row in Open gaps and the MD6 deviation below. Not a design-based P1 run and not split by river/mode/location; genuine scope-mismatch open questions remain (Methow/Chelan/mainstem boundary, species scope) — see the three R2 scope-mismatch rows in Open gaps |
| 4 | **Nothing in this document's numbers has been re-verified against a post-2026-08-21 pipeline run** | Evan | A large batch of correctness fixes landed 2026-08-21/22 (donor hierarchy restored, NA-shadow strata, Green-Duwamish P2 double-count, Yakima 2023–25 expansion, season-matched CRC/creel comparison, PS P3). Every count, ratio, and total quoted below predates them. Treat specific figures here as provisional until a clean end-to-end run is done — see delivery plan step 1 |

Resolved by scope cut: the LCR undocumented scalar, the Bonneville–McNary
totals-only split, and the Buoy 10 Charter Boat mode question are no longer
delivery blockers — that whole block (Columbia mainstem) is out of scope; see
Purpose & deliverable contract. They'd only resurface if Jim/Melissa ever want
WDFW to independently reproduce or reconcile against the consultant's own
ODFW numbers.

## Known defects

- **CRC 508 and 665 river-label corrections — NOT APPLIED, confirmed correct
  as-is 2026-08-29 against the official pamphlet.** Evan initially flagged
  these as wrong (508 should be "White Salmon River" not "Big White Salmon
  River"; 665 should be "Wenaha River" not "Washougal River"), contradicting
  `crc_area_lut.csv`. Evan then photographed the "Catch Record Card Codes"
  page from the official 2025-26 Washington Sport Fishing Rules pamphlet
  (rules in effect 7/1/2025-6/30/2026) as a primary-source cross-check. That
  photo confirms the crosswalk was already right on both: it reads "508 Big
  White Salmon River" and "665 Washougal River, North and West Forks"
  verbatim, with no "Wenaha" entry anywhere on the page. No change made.
- **Photo cross-check of the full 2025-26 catch record card codes page —
  found four more real crosswalk defects, RESOLVED 2026-08-29.** Comparing
  every code visible in Evan's photo against `pst_river_block_crosswalk.csv`
  (not just the two Evan asked about) found: (1) CRC 812 was labeled generic
  "Quilcene" — the pamphlet's actual name is "Big Quilcene River" (paired
  with the existing "Little Quilcene River" row, 766, same Big/Little
  pattern as 508/618) — relabeled. (2) The 563 row (split from the combined
  Cowlitz row the same day) had been given the label "Cowlitz R. between
  Mayfield & Cowlitz Falls Dams" — Claude's own extrapolation from Evan's
  shorthand note, never actually confirmed against the pamphlet. The photo
  shows the real official name is "Mayfield Lake (Reservoir)" — corrected.
  (3) CRC 376 (Naselle, Upper) was missing from the crosswalk entirely —
  only 375 (Lower) was present; added, and 375 relabeled "Naselle (Lower)"
  for clarity. (4) CRC 748 (Big Soos Creek) was missing entirely; added. (5)
  CRC 679 (Wind River, Hwy. 14 to Shipherd Falls) was missing from the
  existing combined Wind River row (677\|680); added to it (not split into
  its own row — no evidence yet its season status differs from its
  siblings). Also noted but not acted on: `crc_area_lut.csv` itself has
  several plain typos (e.g. "Tonuge"/"Rockey"/"Handford"/"Klickiat"/
  "Grantie"/"Thurson"/"Oxbox" for Tongue/Rocky/Hanford/Klickitat/Granite/
  Thurston/Oxbow) and one internal duplicate-code conflict (566 mapped to
  both "Green River (Cowlitz Co.) Toutle Hatchery upstream", which the
  crosswalk correctly uses, and unrelated "Delemeter Creek", which is
  actually 568) — cosmetic issues in that reference file, not in the
  crosswalk or any deliverable output, so left alone. A photo of an official
  regs page is a genuinely useful periodic cross-check beyond what
  `crc_area_lut.csv` alone can catch — worth repeating when the next
  pamphlet cycle (2026-27) is available.
- **`pst_river_block_crosswalk.csv` — RESOLVED 2026-08-29.** Real
  data-quality defects, found by Evan while doing the 2025 season-status
  verification (see the gap row above), not by any code check: CRC 507 was
  mislabeled "Lower Cowlitz" (folded into that river's `crc_areas` alongside
  561) — it's actually Blue Creek (Lewis Co.), an unrelated tributary, per
  `crc_area_lut.csv`. CRC 559 and 563 were combined under one "Upper Cowlitz
  River" row despite being distinct, non-adjacent, dam-bounded reaches
  ("Cowlitz R. above Cowlitz Falls Dam" and "...between Mayfield & Cowlitz
  Falls Dams" respectively) with different season status, which the
  combined label was masking. CRC 583 (Elochoman River, mouth to Foster Rd.
  Bridge) was missing from the crosswalk entirely — only its neighbor reach
  584 was present. The combined "Kalama River" row (600|602|603|604) carried
  600, which is not a valid CRC area code in `crc_area_lut.csv` at all, and
  combined three genuinely different reaches with different salmon seasons
  (602 always closed, 604 open year-round per Evan) under one label. All
  four fixed: 507 removed from the Lower Cowlitz `crc_areas` and given its
  own row; 559/563 split into two rows; 583 added as a new row; the Kalama
  row split into three (602/603/604), 600 dropped. None of this changes
  which CRC areas' harvest gets projected or by how much — `project_crc_
  harvest()` already computes history/means per individual `stream_code`,
  not per combined crosswalk row — this only fixes what river_label a
  reader sees attached to each area's number.
- **`pst_p2_block_ratio.R`'s `apply_p2()` — partial-year ratio expansion —
  RESOLVED 2026-08-29.** Real, not theoretical: found by Evan reviewing the
  2025 output directly. Any CRC area with SOME 2025 CRC harvest (however
  thin) got a real P2 ratio expansion on it, even though CRC only publishes
  through license year 2024 — meaning every area's 2025 harvest figure is
  Jan-Mar only, a fraction of a real year. `apply_crc_projection()`'s (P3)
  "already covered" check only kicks in for an area with ZERO 2025 CRC data,
  so an area with a partial sliver never got the full-year projection
  treatment a fully-uncovered area already received — it just sat there
  looking like ordinary P2 output while being a severe undercount. Worst at
  the river-rollup level when a combined river_label's dominant CRC area
  happened to be the one with a partial sliver: Lewis River (611\|615)
  showed a combined 2025 total of ~1% of its 2022-2024 range, and the
  rollup's `tier` column read `P2\|P3` (611 correctly P3-projected, 615
  incorrectly P2-expanded on Jan-Mar-only harvest) — a legitimate-looking
  blended total actually dominated by the broken component. Also visible in
  Klickitat River (607\|608), the Kalama reaches (602/603/604), and Lower
  Cowlitz (561). Fixed universally, not per-river or per-block: `apply_p2()`
  now excludes any area-year whose CRC harvest comes from a partial-compiled
  year from ratio expansion entirely (computed dataset-wide in
  `run_p2_extrapolation()` from `crc_month`'s actual month coverage, not a
  hardcoded "2025" — self-adjusting for whichever year is partial next).
  Excluded rows fall through to `apply_crc_projection()`, which gives them
  the same full-year projection every other uncovered area gets. Effect:
  P3's 2025 candidate pool grew from 33 to 40 areas (5 genuinely new, plus 2
  already-verified Kalama reaches whose season status carried over
  automatically); overall 2025 P2 total dropped ~1.35M → 1,014,878 trips, P3
  grew accordingly to 507,975. The 5 new areas were verified the same day —
  see the gap
  row below.
- **`pst_p2_block_ratio.R`'s `expand_crosswalk_areas()` — RESOLVED
  2026-08-28.** Real double-counting, found while wiring in R2_external
  (below): when a CRC area maps to more than one crosswalk row — e.g. a new
  district-wide covered_unpartitioned row (R2_external's Upper Columbia
  total) colliding with that same area's pre-existing standalone `CRC_only`
  row (Entiat, Okanogan, Similkameen, Wenatchee River, Icicle Creek, and the
  Priest Rapids-Chief Joseph mainstem areas) — the area-collapsing step
  picked `area_coverage = "standard"` unless every colliding row agreed,
  silently discarding the `covered_unpartitioned` flag. Effect: P2
  independently re-expanded all six named tributaries plus five mainstem
  areas on top of R2's own combined total for the same years, inflating
  ColumbiaUpper by roughly 412K trips (727,352 vs. the corrected 315,603).
  Fixed by making `covered_unpartitioned` win any collision, unanimous or
  not, since that flag specifically means "some P1 source already reports
  this area's trips in aggregate" — collapsing a disagreement toward
  `"standard"` was always the double-counting direction. This is the general
  shape of the Green-Duwamish bug below (a coarse, unpartitioned P1 total
  going unrecognized by a check keyed on per-area attribution), just in the
  crosswalk-merge step rather than `apply_p2()`'s coverage check — worth
  remembering if a future district-wide source collides with existing
  per-river `CRC_only` rows again.
- **`pst_p2_block_ratio.R`'s `p2_loo_check()` — RESOLVED 2026-08-28.**
  Latent crash, also surfaced while wiring in R2_external: if no block/year
  has more than `control$min_donor_areas` distinct donor areas, the
  `group_modify()` step runs zero times and never creates the
  `predicted`/`observed`/`loo_ratio`/`n_donors` columns, so the following
  `mutate(abs_error = predicted - observed)` fails with "object 'predicted'
  not found" instead of returning the empty result callers already handle
  (`p2_loo_summary()` is only called when `nrow(loo) > 0`). Triggered here
  because marking six ColumbiaUpper tributaries `covered_unpartitioned`
  shrank that block's donor pool to fewer than `min_donor_areas` per year.
  Fixed by returning a correctly-typed zero-row tibble up front when no
  block/year is eligible, instead of letting the group-and-mutate pipeline
  run past its non-existent output.
- **`pst_p2_block_ratio.R`'s `apply_p2()` — RESOLVED 2026-08-22.** Real
  double-counting, not theoretical: Green-Duwamish (R4_external, area 746)
  has real P1 trips/harvest for 2022-2025 in the deliverable, but its
  `month` column is NA (`ingest_green_duwamish()` pools trip length across
  the full year, not by month) and `total_salmon_harvest` is NA (the R4
  effort workbook has no harvest column) — both mean it can never pass
  `build_p2_donors()`'s month-matched-CRC-harvest join, so it never appears
  in `donors` regardless of how much real trip data exists. `apply_p2()`'s
  "already covered" check was built from `donors`, so it incorrectly
  treated area 746 as uncovered and expanded it via P2 on top of the real
  R4_external rows already present — visible in the output as Green-Duwamish
  2022/2023 carrying both a P1/`R4_external` row and a P2/`p2_block_ratio`
  row for the same year. Fixed by checking `effort_long`'s real P1 rows
  directly (`apply_p2()` now takes `effort_long` as a parameter), mirroring
  the pattern `apply_crc_projection()` (P3) already used correctly. Worth
  checking the fresh output for any other P1 source that similarly lacks a
  real `month` or `total_salmon_harvest` value — the same failure mode
  would apply.
- **`pst_fw_angler_trips_assembly.R`'s `catch_area_codes` — RESOLVED
  2026-08-22.** Hanford Reach and Snake River (both `covered_unpartitioned`,
  no per-row `catch_area_code` since their district sources give one
  combined total) came out with a blank `catch_area_codes` in
  `pst_fw_trips_by_mode_location.csv`, unlike Quillayute
  (`398|400|402|404|406`) which shows its composite area list because its
  rows carry a real per-stratum `catch_area_code`. The crosswalk already
  knows the composite set for both (534|535|536 for Hanford,
  640|642|644|646|648|650 for Snake) — it just wasn't being surfaced.
  `effort_by_mode_location` now falls back to the crosswalk's own
  `crc_areas` value when the row-level aggregation is empty.
- **`district_creel_ingestion.R`** — `Hanford Reach fall Chinook
  2025` rows carry `year = 2024`. The assembly script works around it by
  trusting the year token in `fishery_name` and logging every corrected row,
  but the ingestion script should be patched.
- **Snake River (R1) crosswalk coverage — RESOLVED (2026-08-19).** Added six
  `source_id = "R1_external"` rows (`block = "ColumbiaSnake"`, one per
  fishery-year, matching `district_creel_summary.csv`'s exact `fishery_name`
  values), `area_coverage = "covered_unpartitioned"` so P2 doesn't also try to
  ratio-expand the six Snake R. System CRC areas independently. Real trips now
  flow into the deliverable instead of resolving `block = "unknown"`.
  Asked Jeremy (2026-08-20) whether CRC harvest at 640/644 (spring Chinook's
  open zones) and 644/648 (fall Chinook's) would align with the creel, given
  the creel only covers specific open zones within each CRC area, not the
  whole area. Confirmed clean, not a mismatch: harvest can only occur where
  fishing is legally open, so CRC harvest for those area-pairs reflects the
  same water the creel already covers completely - closed portions of a CRC
  area produce no harvest to miss. 642/646/650 remain genuinely uncovered by
  any creel, in any season. The real limitation is narrower than a spatial
  gap: Jeremy's total is one combined number per season across whichever
  areas were open (640+644 spring, 644+648 fall), not split area-by-area, so
  a season-specific trips/CRC-harvest ratio using the matching combined area
  harvest is a clean quantity in principle - not yet built as a real
  ColumbiaSnake donor pair (still `covered_unpartitioned`, relying on the
  `columbia_pooled` tier in `pst_p2_block_ratio.R` for now). Worth revisiting
  as a real, Snake-calibrated ratio rather than a permanent limitation.
- **`multi_fishery_creel_run_ledger.csv`** — 98 rows, all `ok` as of the
  2026-08-14 run. The previous 3 errors / 1 skip are resolved. Lower Cowlitz
  remains absent from the output entirely: it lacks the data needed for
  effort estimation, a genuine gap rather than a pipeline bug.
  **Stale as of 2026-08-21:** the row count will change on the next run —
  winter/summer gamefish fisheries are now filtered out of the ledger along
  with the estimates, and the restored `[S4]` donor hierarchy changes which
  fisheries produce usable trips. Re-read the ledger after the next run
  rather than carrying 98 forward.
- **Superseded files on disk — RESOLVED (2026-08-17).** `multi_fishery_trip_summary.csv`
  was stale but still present, and had already caused one analysis to be run
  against wrong numbers; `results.csv` is no longer anywhere in the repo. The
  reorganization moved the former to
  `analysis/outputs/archive/multi_fishery_trip_summary_superseded_20260817.csv`,
  and archived the two scripts that wrote the live pipeline's filenames
  (`multi_fishery_trip_summary.R`, `multi_fishery_harvest_summary.R`) so they
  can no longer clobber `multi_fishery_creel_summary.R`'s outputs. The
  assembly script still logs a `defect` if it finds a superseded file.
- **CRC compile** — region labels lost for 2023–2024; subtotal blocks
  retained as leaf rows causing ~2× double-count. Confirm the current
  `crc_freshwater_harvest_2010_2024_tidy.csv` postdates the fix. Still
  unverified as of 2026-08-22 — worth closing out, since this file is now
  the sole harvest basis for P2, P3 (both variants), and the CRC-vs-creel
  bias comparison.
- **NEPA workbook** — fill-right errors in the 2022/2024 even-year columns.
  **Downgraded 2026-08-21 from a delivery concern to a diagnostic-only
  one:** no NEPA-derived figure feeds any deliverable row. The workbook is
  read only by `run_nepa_pure_crc_comparison()` (PART 2 of
  `pst_crc_harvest_projection.R`), whose output
  (`pst_fw_nepa_vs_pure_crc_comparison.csv`) is explicitly diagnostic. These
  errors matter for interpreting that comparison, not for our numbers.

## Why the assembly output is not yet the deliverable

`analysis/pst/03_analysis/pst_fw_angler_trips_assembly.R` currently answers "what
can we support right now," not "what do we owe Northern Economics." The P2
ratio/expansion gap and the CRC-vs-creel bias handle are covered in Track A —
assembly in practice above; the unresolved `mode = "unknown"` share is
covered in Track B results above. Three more stand between the two, covered
here.

### Trip expansion — resolved, then silently regressed, then restored

An earlier version of this document reported that 29% of measured effort was
being silently dropped (Nisqually, Puyallup/Carbon). **That finding was an
artifact of reading a superseded file** (`multi_fishery_trip_summary.csv`).
Against `multi_fishery_creel_trips.csv` at that time:

- All 49 fisheries expanded cleanly — **zero** with `NA` trips.
- The run ledger was 98 rows, **all `ok`** — no errors, no skips.
- Nisqually 2022 *and* 2023, all four Puyallup/Carbon years, and all four
  Drano years carried trip estimates. The `[S4]` donor hierarchy was doing
  the work: 778 rows on own-month trip length, 166 `area_season`, 66
  `type_season`, 8 `fishery_season`.
- Lower Cowlitz is absent entirely — the one genuine, confirmed data gap.

::: {.callout-important}
**Those figures describe the PREDECESSOR script, and the fix they describe
was lost for a while.** When `multi_fishery_trip_summary.R` and
`multi_fishery_harvest_summary.R` were merged into
`multi_fishery_creel_summary.R`, the entire `[S4]` donor hierarchy was
dropped in the process — the merged script computed only a plain monthly
mean and went back to dropping any month × area × angler-type cell with real
effort but no completed interview. That regression surfaced 2026-08-21 not
as bad numbers but as a hard error: `pst_fw_angler_trips_assembly.R` expects
a `trip_length_source` column on every trips row, and the merged script no
longer produced one.

The hierarchy (month → `area_season` → `type_season` → `fishery_season` →
none), the `trip_length_source` / `n_donor_obs` labelling, and the
`trip_length_stability` diagnostic were restored to
`multi_fishery_creel_summary.R` on 2026-08-21, ported from
`analysis/archive/multi_fishery_trip_summary_superseded_20260817.R`.

**The tier counts above have not been re-measured since.** They are the
predecessor's numbers, carried here for shape only — expect the restored
run to differ, because the merged script also fixed the Lower Chehalis
NA-shadow duplication and now excludes gamefish fisheries. Replace these
four numbers from `multi_fishery_creel_trips.csv`'s own
`trip_length_source` breakdown after the next clean run.
:::

::: {.callout-warning}
**The real hazard in the new file is the opposite of a drop — it's a
double-count.** `multi_fishery_creel_trips.csv` stacks both the week- and
month-stratified runs, so every stratum appears twice. Naively summing gives
**1,881,197** trips against a correct month-basis figure of **954,177**. The
assembly script filters to `PE_PERIOD = "month"` and asserts uniqueness of
the design grain afterward — but any ad hoc query against the raw file that
skips that filter will silently double every number in it.
:::

### P3 fallbacks — Puget Sound now covered internally; coastal still open

**Superseded in part, 2026-08-21.** This section used to say Puget Sound
2025 was waiting on an external NEPA projection. It no longer is: PS 2025 is
now produced inside this pipeline by `pst_crc_harvest_projection.R`'s second
control variant (`CRC_PROJECTION_CONTROL_PS`) — a same-parity (odd-year)
5-year mean over 2015/2017/2019/2021/2023 pure CRC harvest, expanded through
PS's own real 2023 block-year creel/CRC ratio. Chosen over the 6-year
trailing mean used elsewhere because PS's odd-year Pink dominance makes a
trailing mean average an odd year's inflated harvest against even years'
near-absence of it, understating an odd target year. Provisional pending
Jim's sign-off.

What that leaves genuinely open:

- **Coastal 2025** — no fallback identified. WACoast gets P3's 6-year
  trailing mean where an area has enough history; areas failing
  `min_history_years` have nothing.
- **`ingest_published_fallback()` is still a stub** — it logs notes and
  returns `NULL`. That is now a *deliberate* state rather than a gap: no
  external published figure feeds any deliverable row, and the ratio
  question that motivated it is closed (see the struck-through odd/even row
  in Open gaps). Keep the stub only as the hook for a genuinely new
  published source; it is not a missing PS input any more.

The year-specific **historical** NEPA estimates for 2022–2025 were never
located. That is no longer a delivery dependency for the same reason — but
it does still matter for the `pst_fw_nepa_vs_pure_crc_comparison.csv`
diagnostic, which can only compare what the workbook actually contains.

### Trips vs. angler-days — still open

See Glossary & unit conventions for the trips-vs-days distinction itself. No
conversion is applied anywhere in this pipeline, and that decision has not
been made — it belongs to Jim and Northern Economics, not to a pipeline
default.

## Coverage map {#sec-map}

::: {.callout-note}
**Placeholder.** Intent: a WA state map with river systems symbolized by
assembly status — resolved (P1), resolved by ratio (P2), published-figure
fallback (P3), and open gap — with a year facet or toggle. This is the
progress tracker for the remaining outreach, and the single artifact most
worth showing Jim in a status conversation.

Requires: a river-system spatial layer (`data/spatial/`) joined to the
registry crosswalk on its key, and the coverage ledger for the status fill.
This layer does not exist in the repo as of this writing — the map is
blocked on the data, not on the code below.
:::

```{r}
#| label: coverage-map
#| eval: false

# Sketch — not yet wired. The spatial layer this references does not exist
# in the repo yet; the path below is aspirational, not a stale reference to
# a file that moved.
# library(sf); library(leaflet)
#
# rivers_sf <- st_read(here::here("data", "spatial", "wa_river_systems.gpkg")) |>
#   left_join(crosswalk, by = c("river_name" = "crosswalk_river")) |>
#   left_join(coverage,  by = c("river", "year"))
#
# pal <- colorFactor(
#   palette = c(P1 = "#1b7837", P2 = "#7fbf7b", P3 = "#fdb863", gap = "#b2182b"),
#   domain  = rivers_sf$tier_or_gap
# )
#
# leaflet(rivers_sf) |>
#   addProviderTiles("CartoDB.Positron") |>
#   addPolylines(color = ~pal(tier_or_gap), weight = 3, label = ~river) |>
#   addLegend(pal = pal, values = ~tier_or_gap, title = "Assembly status")
```

## Open gaps

Both tables below are long by design — the gap register is the audit trail,
not a summary — so each is scrollable rather than pushing the rest of the
document down the page.

```{=html}
<style>
.gap-scroll {
  max-height: 30rem;
  overflow: auto;               /* vertical AND horizontal - the gt gap
                                   register is wide as well as long */
  border: 1px solid var(--bs-border-color, #dee2e6);
  border-radius: 0.375rem;
  padding: 0.5rem 1rem;
  margin-bottom: 1.5rem;
  background: var(--bs-body-bg, #fff);
}
.gap-scroll table { margin-top: 0.25rem; }
</style>
```

::: {.gap-scroll}
```{r}
#| label: gap-table

if (is.null(gaps)) {
  cat("Gap ledger not found.\n")
} else {
  gaps |>
    arrange(block, severity, source_id) |>
    gt() |>
    tab_header(title = "Logged gaps (R2)")
}
```
:::

Gaps tracked outside the ledger, because they are outreach rather than code:

::: {.gap-scroll}
| Gap | Owner | Status |
|---|---|---|
| Yakima workbooks, 2023–2025 | Todd Miller (R3) | **Resolved 2026-08-21.** Files received; now ingested via `ingest_yakima_site_combined()` (see Known defects/Open gaps entry below) |
| McNary workbooks, 2023–2025 | Todd Miller (R3) | This row appears stale as of 2026-08-21: `ingest_mcnary()` already exists, is wired into `ingest_r3_mid_columbia()`, and all three files (2023-2025) are already present in `input_files/pst/R3_creel/`. Not verified this session (out of scope of the Yakima fix) — confirm output is actually landing in `district_creel_summary.csv` before treating this as still open |
| Upper Columbia | Chad Jackson (R2) | **Received 2026-08-28.** Chad replied (email, PDF captured at `input_files/pst/R2_creel/R2_Angler_Trips_2022-2025.pdf`) with a single combined district-wide angler-trip total per year — no river, mode, or location split, the coarsest source in the pipeline. Ingested via `ingest_r2_upper_columbia()`/`R2_external`. See the three new gap rows below for what this doesn't resolve |
| Lake Wenatchee (CRC area 670) | Chad Jackson (R2), part of the Upper Columbia request above | **Received 2026-08-28**, as part of Chad's combined total (his footnote explicitly names Lake Wenatchee). No longer excluded by the P2 harvest-scale guardrail (`max_target_harvest_multiple`) — that guardrail only ever applied to a *ratio-based* P2 extrapolation, and this area's trips now arrive via R2's district total instead, which sidesteps the guardrail rather than resolving whatever made it fire in the first place |
| R2 (Chad Jackson) scope mismatch: Methow River (621) | Evan / Chad | **Open.** Methow is grouped under this district in `pst_river_block_crosswalk.csv` but is NOT named in Chad's footnote (which lists Wenatchee, Entiat, Chelan, Okanogan, Similkameen, plus the Priest Rapids-Chief Joseph mainstem and Icicle River/Lake Wenatchee). Left `CRC_only`/active for independent P2 expansion rather than guessed either way — if Chad's total secretly includes Methow, this double-counts; if it doesn't, leaving it active is correct. Needs a direct question back to Chad |
| R2 (Chad Jackson) scope mismatch: Chelan (CRC 552) | Evan / Chad | **Open.** Chelan is named in Chad's footnote but has no `pst_river_block_crosswalk.csv` row at all (confirmed via `crc_area_lut.csv` — "Chelan River" exists as CRC 552, never added to the crosswalk). Invisible to P2 either way until a row is added; unclear whether Chad's total already includes it |
| R2 (Chad Jackson) scope mismatch: mainstem boundary | Evan / Chad | **Open, mostly resolved by inference.** Chad's footnote bounds the mainstem at Chief Joseph Dam; the crosswalk's pre-existing mainstem row spanned Priest Rapids all the way to Grand Coulee (7 areas). Per `crc_area_lut.csv`'s own dam-to-dam descriptions, areas 537/539/541/543/545 fall exactly within Priest Rapids-to-Chief-Joseph (marked covered by R2_external); 547 and 549 (Roosevelt Lake) are above Chief Joseph and were split into their own still-open `CRC_only` row. Confident in the split itself; not confident Chad's total is silent on 547/549 rather than simply forgetting to mention them |
| R2 (Chad Jackson) species scope | Evan / Chad | **Open, assumed not confirmed.** Per Evan's instruction (2026-08-28), R2's total is assumed to be PST-scope salmon per the original request (Jim Scott's email explicitly excludes steelhead) — Chad's reply doesn't itself restate the species scope, so this is an assumption carried from the original ask, not an explicit confirmation |
| Snake River | Jeremy Trump / Mike Gembala | Resolved 2026-08-20 — crosswalk added 2026-08-19; zone-coverage question (does creel cover all open water so a Snake-internal trips/CRC ratio would be clean?) answered 2026-08-20: yes, clean, see the Known defects entry above. Data refresh 2026-08-20 also filled in the previously-blank Fall CH 2025 row (5023 boat / 87 shore); added the matching crosswalk row the same day. A real Snake-calibrated donor pair (season-specific area harvest vs. Jeremy's trips) is now feasible but not yet built |
| Green River (Green-Duwamish) creel data | Nathanael Overman | **Received and ingested 2026-08-20/21.** Raw interviews + effort-estimate workbooks landed; `ingest_green_duwamish()` in `district_creel_ingestion.R` derives `total_trips_est` as R4's own angler-hours effort ÷ mean trip length computed from the raw interviews (completed trips only, pooled across the full year — see MD3 below). Four `R4_external` crosswalk rows added 2026-08-21 (2022-2025), superseding the `CRC_only` placeholder for those years. Cross-check against the workbook's own ANGLERS block: derived trips land within ~2-9% of it across all 8 year × mode combinations, boat consistently a little low. Asked Nathanael (2026-08-21) whether he could summarize the effort totals by month instead of by the current two-period split — would let this go from annual to monthly `pe_period` grain and give period-specific rather than pooled trip lengths; a potential future refinement, not blocking |
| Columbia Basin empirical effort/catch 2022–2025 (to replace FEIS rates) | Mark S. / CRD | Team check pending |
| Coastal freshwater 2025 | — | No fallback identified |
| McNary 2022 | Todd Miller | Confirmed missing; no source |
| 2025 CRC harvest | Heidi / Harvest Data Team | Publishes October — binding constraint on 2025 |
| ~~NEPA odd/even ratio misapplication~~ | ~~Christina Iverson~~ | **Closed 2026-08-22, no call needed.** This pipeline never consumed NEPA's 3.44/8.65 ratio: `ingest_published_fallback()` only ever logged notes, and every trips-per-salmon figure here is derived empirically from our own paired creel/CRC data (P1 design-based, P2 CRC-denominated block ratios). The workbook's "applies 3.44 to every year regardless of parity" defect is real but belongs to that external deliverable, not to ours. Retained in the intro/methods children as reference context for why R4 forbids cross-block ratio transfer |
| Lower Cowlitz | — | Genuine data gap, not a bug; `KNOWN_FAILED` |
| Yakima 2023–2025 trip expansion | Evan | **Resolved 2026-08-21.** OQ4's premise was wrong — the expansion was never dropped from the 2023+ workbooks, it was never read from the right sheet. The `Summary` sheet this parser used to read really is sampled-only, but the per-site `Prosser`/`Horn Rapids` tabs (previously ignored) each carry a `Combined` weekly table with the same expanded Angler Trips / Effort Pole Hrs columns 2022's `Sheet1` has. `ingest_yakima_site_combined()` now reads both site tabs (`Horn Rapids` is all-zero in every week of all three years — no coverage at that site in this data), sums per week, and prorates to months. Expanded totals: 3,854 / 3,634 / 3,675 trips and 7,781 / 8,609 / 6,428 effort-hrs for 2023/2024/2025 respectively, verified against each sheet's own season-total row before trusting the parser. Caught and fixed a real bug along the way: the `Combined`, `Bank Summary`, and `Boat Summary` blocks in each sheet reuse identical week-label text, so an unscoped label match would have summed all three blocks together (triple-counting); the parser now bounds rows strictly between the `Combined` and `Bank Summary` section labels |
| Lower Chehalis 2023 NA-area shadow rows | Evan | **Resolved 2026-08-21.** `section_crc_area` in `multi_fishery_creel_summary.R` now drops the NA variant for a section whenever a real `crc_area` is also present for that section, instead of only warning about the resulting fan-out |
| Drano `angler_final = "fail"` | Evan | **Follow-up task.** Failed bank/boat classification recoded `"unknown"` downstream; root cause is an unrecognised `count_type` value in `prep_dwg_effort_index()`'s Drano branch (4 rows, Drano Lake salmon and steelhead 2024, 1,745 effort-hrs). `multi_fishery_creel_summary.R` now writes `multi_fishery_creel_effort_index_fail_counts.csv`/`.rds`, recording every unrecognised `count_type` by fishery — next DB-connected run of the pipeline will surface the exact value(s) so the Drano branch's `case_when()` can be patched precisely instead of guessed at |
| Additivity check, 122 strata | Evan | Relative differences ~1e-4 — floating-point, not a missing species. Consider raising the tolerance so the check stops crying wolf |
| week vs. month `pe_period` | Jim / Evan | 17 fisheries differ >15% between stratifications. `PE_PERIOD` is currently `"month"`; the choice needs to be made explicitly, not defaulted |
| River-to-creel crosswalk | Evan | Raised as open in the Track B producer before the current design landed. Appears superseded: matching now keys on `catch_area_code` via `pst_river_block_crosswalk.csv`, not river name. Confirm and close, or reopen with specifics |
| Pooled proportions for uncreeled rivers | Evan | Same vintage. Appears superseded by P2 within-block ratios with leave-one-out validation — which is the regional fallback that request asked for. Confirm and close |
| Creel season truncation vs. legal harvest period | Evan | **Open, needs regulatory verification per fishery.** A month at the edge of a creel program's operating window can be incomplete for two different reasons, and only one is currently detectable: (1) a sampling gap within a fully-covered season — visible as low `prop_strata_estimated` (now threaded through into `effort_long` and surfaced as `min_prop_strata_estimated` in the P2 donor/ratio tables and the CRC-vs-creel comparison); or (2) the creel program itself stopped running before the fishery's legal harvest period — and therefore CRC's own reporting window — actually closed that year. (2) is invisible in the data: a creel program that ran flawlessly for the days it operated can show 100% stratum coverage and still only measure PART of the month's true activity, because harvest legally continued after the survey ended. This can only be resolved by checking each fishery's actual season regulations (open/close dates) against its creel program's own operating window, not by anything computable from the pipeline's own tables. `pst_fw_creel_season_edges.csv` (new, written by `pst_fw_angler_trips_assembly.R`) lists every fishery-year's first and last surveyed month as the candidate set to check — that is the where-to-look list, not a resolution. Once a specific fishery/year/month is confirmed truncated, add it to `SEASON_TRUNCATED_MONTHS` in `pst_p2_block_ratio.R`; that table (currently empty) is the entire mechanism for excluding it from every creel-trips/CRC-harvest ratio construction (both the P2 donor ratio and the CRC-vs-creel bias comparison) — no other code change needed once a row is added |
| 2025 CRC harvest projection vs. actual season closures | Evan / Jim | **RESOLVED 2026-08-29 — verified for all 40 candidates AND wired directly into the deliverable, running automatically on every pipeline run.** 2025 is not a future year being forecast -- it already happened; "was this fishery actually open in 2025?" is a checkable FACT in WDFW's own regulations. The correction (`compute_monthly_harvest_share()`/`scaffold_season_status_lookup()`/`apply_season_status_correction()`, now PART 3 of `pst_crc_harvest_projection.R`) runs as part of the assembly script's own P3 step -- no separate manual pass anymore (the standalone `pst_crc_projection_season_check.R` is archived, superseded). Evan checked the original 33 CRC areas against the 2024-25/2025-26 pamphlets, then 7 more (Lewis River 611/615, Klickitat River 607/608, Lower Cowlitz 561, plus 2 Kalama reaches whose status carried over automatically) that became new candidates once `apply_p2()`'s partial-year fix (see Known defects) stopped letting a thin Jan-Mar-only 2025 CRC sliver bypass projection. **Final tally across all 40: 7 fully closed** (zeroed), **22 partially closed** (prorated by historical month-of-year harvest share), **11 confirmed open**. `pst_fw_crc_projection.csv`'s `total_salmon_harvest`/`angler_trips` columns now carry the corrected, deliverable-feeding values directly (54,442 → 53,060 harvest, -2.5%; 507,974 → 494,286 trips), with the pre-correction figures kept alongside as `_uncorrected` columns for audit, never silently overwritten. A future P3 candidate that isn't yet in `pst_season_status_lookup.csv` gets auto-scaffolded there as `UNVERIFIED`, is left with its original uncorrected value, and is logged as a `crc_projection_season` gap -- verified end-to-end (confirmed a genuinely-unverified test area is correctly left unchanged and flagged, not silently assumed open). The reverse case (an area excluded by `min_history_years` for thin history but CONFIRMED open in 2025) is still unhandled in code -- no gap row generalizes to that yet. **Real data-quality defects surfaced during verification, fixed in `pst_river_block_crosswalk.csv` (2026-08-29):** CRC 507 was mislabeled "Lower Cowlitz" (it's Blue Creek, unrelated to the Cowlitz, and was incorrectly folded into the Lower Cowlitz creel_pe rows' `crc_areas`); CRC 559 and 563 were combined under one "Upper Cowlitz River" label despite being distinct, non-adjacent dam-bounded reaches with different season status; CRC 583 (Elochoman, mouth to Foster Rd. Bridge) was missing from the crosswalk entirely; the combined "Kalama River" row carried an invalid code (600, not a real CRC area) and hid that its reaches have different seasons under one label. All four split/corrected — see that file's own notes on each new row. **Two more corrections Evan proposed (CRC 508, 665) were checked against the official 2025-26 pamphlet and found NOT needed — the crosswalk was already right** — see Known defects below, which also covers four further defects a full photo cross-check of the pamphlet's codes page turned up beyond those two |
| Illegal/erroneous harvest in CRC records: keep or exclude? | Evan / Jim | **Open, needs a policy decision plus a data-QA task -- two different questions, don't conflate them.** Every CRC card is currently taken at face value in `parse_crc_freshwater_harvest.R` -- there is no mechanism anywhere in this pipeline that identifies, flags, or distinguishes (a) harvest that was genuinely illegal (an angler really did keep a fish, but doing so violated regs -- closed season/area, over limit, wrong size or sex) from (b) harvest that's simply erroneous (a card mis-filled, mis-dated, or mis-coded -- not a real event at all). These need different handling and different owners: (a) is a POLICY call, not a data fix -- an illegal-but-real trip still generated real economic activity (travel, lodging, guide spend), so there's a legitimate argument for keeping it in an economic valuation; equally there's a legitimate argument that WDFW should not implicitly certify illegal harvest as legitimate effort in a deliverable going to a PST treaty partner/co-manager audience. That call belongs to Jim, not a pipeline default. (b) is a correctness issue, not a judgment call -- an erroneous record should be excluded once identified, the only question is how to identify it. Distinct from (but related to) the "2025 CRC harvest projection vs. actual season closures" gap above: that one is about verifying whether a PROJECTED year's fishery was open before trusting a statistical projection onto it; this one is about whether ACTUAL, already-reported historical CRC harvest (2022-2024, feeding P2 ratio construction, the CRC-vs-creel bias comparison, and P2's real target-area expansion, not just the 2025 projection) might already include harvest that shouldn't count at all. Needed before either can be acted on: does CRC (or WDFW enforcement records) carry any existing violation/flag field this pipeline isn't reading, or does identifying (a) and (b) require a manual cross-reference against regulations and enforcement data the same way the season-closure check does? |
| Some P3 2025 projections look unrealistic next to the most recent real years | Evan / Jim | **Open, needs a plausibility check beyond what's currently run.** Worked example: Whatcom Creek (CRC area 884, PugetSound) shows real (P2-expanded, ratio-based on real donor areas) harvest of 61 (2022), 135 (2023), and 13 (2024) -- then a projected 2025 harvest of **728.4**, 5-30x every one of the three most recent real years, via the same-parity 5-yr mean (`CRC_PROJECTION_CONTROL_PS`). The existing sanity check (`check_projection_vs_partial_actual()`, `pst_fw_crc_projection_sanity.csv`) only compares the projection's implied Jan-Mar share against 2025's own partial actual -- it does NOT compare against the stream's own recent (2022-2024) full-year real harvest, so a case like this can sail through unflagged if the tiny bit of 2025 Jan-Mar data happens not to contradict it. Two non-exclusive explanations, and they call for different responses: (1) Whatcom Creek's true harvest genuinely declined between the 2015-2021 window the mean is built from and 2022-2024 -- run decline, habitat change, regulation tightening -- in which case the 5-yr mean is averaging in a baseline that no longer reflects current reality, related to but distinct from the closure-verification gap above (this isn't necessarily a full closure, just a lower baseline); or (2) one or more of the 5 historical odd years (2015/2017/2019/2021/2023) has a data problem (miscoded stream/area, a genuine one-year spike) inflating the mean by itself, which `min_history_years` doesn't catch since it only checks HOW MANY years are present, not whether an individual year's value is itself an outlier. Needed: an additional sanity check comparing each area's projected mean against its own most recent 1-3 real years (not just partial 2025 actuals), flagging anything outside a defensible multiple (a `p90_ape`-style or ratio-band check, mirroring how P2's own guardrails already do this for the ratio -- the projected HARVEST itself currently has no equivalent check against recent real harvest) -- and manually reviewing every stream this flags before treating its 2025 figure as final |
| Gamefish fisheries excluded from scope | Evan | **Applied 2026-08-21, worth a confirmation.** Winter/summer gamefish fisheries are now filtered out of `multi_fishery_creel_summary.R` at both ingest and final output, on the same PST-salmon-scope basis as the existing steelhead-primary exclusion. Mixed-target surveys naming both (e.g. `Skykomish summer Chinook and gamefish`) are deliberately NOT excluded — same carve-out logic the steelhead filter uses for `Lower Cowlitz salmon and steelhead`. Confirm the retained mixed-target set is what's intended before delivery |
:::

## Documented methodology deviations

| # | Deviation | Rationale | Scope |
|---|---|---|---|
| MD1 | Mid-Columbia and Snake River Track A totals accepted as **pre-computed** inputs (`district_creel` source_ids: `R3_external`, `R1_external`), bypassing the P1 hierarchy | Todd Miller's (R3) and Jeremy Trump's (R1) spreadsheet totals are the only expansion that exists for these systems; unvetted external calculations, not design-based PE runs | Hanford Reach, Yakima, McNary (ColumbiaUpper), Snake River (ColumbiaSnake) — all delivering as of 2026-08-19. Upper Columbia (R2) got the same call 2026-08-28 — see MD6, which documents the extra caveats R2's data raises beyond what MD1 covers for R1/R3. |
| MD2 | Guided/unguided coded `"unknown"` for all mid-Columbia | Field not collected by R3 programs | Per R3 |
| MD3 | Green-Duwamish Track A totals are **derived, not sourced** — unlike MD1's R1/R3 rows, R4 (Nathanael Overman) does not hand us a finished trip count. `ingest_green_duwamish()` computes `total_trips_est` itself as R4's own angler-hours effort estimate ÷ mean trip length computed from R4's raw creel interviews | R4 supplies effort (angler-hours) and raw interview microdata but no trip total; the workbook's own ANGLERS block is present but of uncertain accounting basis, so it's kept as a cross-check only rather than used as the source of `total_trips_est` (Evan's call, 2026-08-21) | Green-Duwamish (PugetSound) — 2022-2025. **Not generalizable**; this is a different kind of unvetted calculation than MD1's "accept the district's own total," worth flagging separately in review |
| MD4 | Puget Sound 2025 harvest is projected on a **same-parity (odd-year) 5-year mean**, not the 6-year trailing mean every other block uses (`CRC_PROJECTION_CONTROL_PS` vs. `CRC_PROJECTION_CONTROL` in `pst_crc_harvest_projection.R`) | PS Pink salmon return almost exclusively in odd years. A trailing mean averages an odd year's Pink-inflated harvest together with even years' near-absence of it, systematically understating an odd target year — 2025 is odd. Same-parity averaging is the mechanic NEPA's own workbook uses and that this file's PART 2 diagnostic already reconstructs; MD4 promotes it from diagnostic to projection basis for PS only. Window is 2015/2017/2019/2021/2023 over **pure** (pre-creel-substitution) CRC | PugetSound 2025 only. Ratio comes from PS's own real 2023 block-year creel/CRC pair (`max(history_years)` resolves to 2023 automatically), **not** from NEPA's 3.44/8.65 — those external ratios are referenced in the intro/methods children for context but are consumed nowhere. Provisional pending Jim's sign-off (2026-08-21) |
| MD5 | Winter/summer **gamefish** fisheries excluded from ingest and output | Same PST-salmon-scope basis as the pre-existing steelhead-primary exclusion [S1]. Mixed-target surveys naming both salmon and gamefish (e.g. `Skykomish summer Chinook and gamefish`) are deliberately retained, matching the carve-out the steelhead filter already makes for `Lower Cowlitz salmon and steelhead` | All blocks, applied in `multi_fishery_creel_summary.R` at both the candidate-fishery stage and the final output filter (2026-08-21) |
| MD6 | Upper Columbia (R2) Track A total accepted as a **pre-computed, single combined district-wide** input (`R2_external`), following MD1's pattern but coarser than every other MD1 source — no river, mode, or location split at all (R1/R3 at least split boat vs. bank) | Chad Jackson's (WDFW Region 2) 2026-08-27 reply is the only data this district has ever produced for this request; unvetted, un-partitioned, but real. `area_coverage = "covered_unpartitioned"` marks the CRC areas it covers (Entiat 586, Okanogan 627, Similkameen 629, Lake Wenatchee 670, Icicle Creek 672, Wenatchee River 674, and mainstem Priest Rapids-Chief Joseph 537\|539\|541\|543\|545) so P2/P3 don't independently re-expand them | Upper Columbia (ColumbiaUpper) 2022-2025. **Not fully resolved** — Methow (621), Chelan (CRC 552), and the mainstem stretch above Chief Joseph (547\|549) are deliberately left OUT of the covered set pending clarification from Chad (see the three R2 scope-mismatch rows in Open gaps); species scope (salmon-only) is assumed, not confirmed |

For reference, the 2022 Hanford Reach figures accepted under MD1:
26,342 boat trips and 2,103 bank trips (92.6 / 7.4). Track B location is
therefore satisfied empirically for that river-year even though Track A is
not design-based.

# Delivery plan

Re-ordered 2026-08-22 to put the verification work first: a batch of
correctness fixes landed 2026-08-21/22 and **none of their output has been
seen yet**, so every downstream judgment is waiting on one clean run.

| Step | Owner | Blocking |
|---|---|---|
| 1. **Clean end-to-end re-run + verify.** `multi_fishery_creel_summary.R` (needs VPN/DB) → `district_creel_ingestion.R` → `pst_fw_angler_trips_assembly.R` → render. Then re-read: run-ledger row count, `trip_length_source` tier breakdown, the season-matched CRC/creel median, and whether any P1 row still draws a duplicate P2 row. **Partially done 2026-08-28**: the non-DB half (`district_creel_ingestion.R` → `pst_fw_angler_trips_assembly.R` → `pst_fw_build_jim_workbook.R`) ran clean while wiring in R2 — found and fixed two real bugs in the process (see Known defects: `expand_crosswalk_areas()` double-count, `p2_loo_check()` crash). The DB-dependent half (`multi_fishery_creel_summary.R`) still has not run since the 2026-08-21/22 fixes landed | Evan | DB-dependent half still gates the P1 tier-count figures below |
| 1b. Follow up with Chad Jackson on the three R2 scope-mismatch questions (Methow, Chelan, species scope) — see Open gaps | Evan | Chad |
| 2. Confirm no other P1 source hits the Green-Duwamish failure mode (NA `month` or NA `total_salmon_harvest` → invisible to `build_p2_donors()` → wrongly treated as uncovered) | Evan | Step 1 |
| 3. Review P3 2025 projections against each area's own recent real years; Whatcom Creek (884: 61/135/13 real → 728 projected) is the worked example | Evan / Jim | Step 1 |
| 4. ~~Verify all P3 season candidates and wire the correction into the deliverable~~ — **done 2026-08-29**: all 40 candidates verified (7 fully closed, 22 partially closed, 11 open), and the correction now runs automatically inside `pst_fw_angler_trips_assembly.R`'s own P3 step on every pipeline run — `pst_fw_crc_projection.csv` carries the corrected values directly. ~~Confirm the CRC 508/665 relabels~~ — done, both confirmed correct as-is against the official pamphlet. Remaining: the thin-history-but-confirmed-open case still has no code | Evan / Jim | Design the thin-history-but-open case, when there's time |
| 5. Verify creel season truncation vs. legal harvest period; populate `SEASON_TRUNCATED_MONTHS` where confirmed | Evan | Regulations review |
| 6. Decide illegal/erroneous CRC harvest handling — policy call (keep/exclude) plus whether any violation flag exists to key on | Jim / Evan | Jim |
| 7. Sign off on the PS 2025 projection basis (MD4: same-parity 5-yr mean, our own 2023 ratio) | Jim | Jim |
| 8. Confirm McNary 2023–2025 output actually lands in `district_creel_summary.csv` — `ingest_mcnary()` exists and is wired in, so the "not yet ingested" gap row looks stale | Evan | Step 1 |
| 9. Confirm D5 (upriver spring Chinook) | Jim | Jim |
| 10. Replace FEIS rates with CRD empirical where available | Evan | Mark / CRD |
| 11. Decide `PE_PERIOD` (week vs. month) on the record | Jim | No |
| 12. Freeze 2022–2024; flag 2025 as provisional | Evan | CRC October |
| 13. Wire the coverage map | Evan | Spatial layer |
| 14. ~~Land manual runs (Puyallup/Carbon, Nisqually, Drano)~~ — superseded; all three carry trip estimates through the restored `[S4]` hierarchy, no manual runs outstanding | — | — |
| 15. ~~Ingest Todd's 2023–2025 Yakima workbooks~~ — done 2026-08-21; `ingest_yakima_site_combined()` reads the Prosser/Horn Rapids `Combined` tables directly, no correspondence with Todd needed (OQ4 resolved) | — | — |
| 16. ~~Resolve odd/even ratio with Christina~~ — dropped 2026-08-22; this pipeline derives every trips-per-salmon figure from its own paired creel/CRC data and never consumed NEPA's 3.44/8.65 | — | — |
| 17. ~~Reconcile A1–A4 wording between this doc and `PST_ANALYSIS.md`~~ — done 2026-08-17; `PST_ANALYSIS.md` folded into the intro/methods children and retired | — | — |

::: {.callout-warning}
**2025 cannot be complete before October.** The 2025 CRC harvest file is the
input for every P2 and P3 row in that year. Options are to deliver 2022–2024
final with 2025 following, or to deliver 2025 as projected and labeled per R5.
This should be surfaced to the consultant now rather than at delivery.
:::
