---
title: "Cowlitz: which guided trips count as salmon trips?"
subtitle: "Guide logbook vs. creel interviews, Cowlitz below Mayfield Dam (CRC 561)"
format:
  html:
    embed-resources: true
    toc: true
    fig-width: 10
    fig-height: 6
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| include: false
QUIET <- TRUE
source(here::here("analysis", "pst", "03_analysis", "cowlitz_salmon_trip_review.R"))
tab <- function(pattern) {
  hit <- names(tables)[grepl(pattern, names(tables), fixed = TRUE)]
  if (length(hit) == 0) return(cat("*Not available - source data missing.*"))
  knitr::kable(tables[[hit[1]]], format.args = list(big.mark = ","))
}
pl <- function(name) if (!is.null(plots[[name]])) print(plots[[name]]) else
  cat("*Not available - source data missing.*")
```

The guide logbook records what each guided trip **caught**, not what it was
**targeting**. The Cowlitz creel records both, for the parties it interviews.
This puts the two side by side to decide which logbook trips should count as
salmon trips.

## When salmon are kept (CRC harvest)

```{r}
pl("crc")
```

## Guide logbook: what guided trips caught

```{r}
pl("lb")
```

```{r}
pl("lb_pct")
```

```{r}
tab("LOGBOOK: guided angler-trips by month x catch")
```

By year:

```{r}
tab("LOGBOOK: guided angler-trips by year x catch")
```

## Creel: what guided anglers said they were targeting

```{r}
#| fig-height: 8
pl("cr_target")
```

```{r}
tab("CREEL: row % stated target by month")
```

## Creel: what guided parties caught

Same catch classes as the logbook, for comparison.

```{r}
pl("cr_caught")
```

## The key question: given what was caught, what was targeted?

Each row is what a guided party caught; the cells are the share that named
each target. The **steelhead only** and **nothing** rows are the logbook trips
we can't classify from catch alone.

```{r}
pl("calib")
```

```{r}
tab("CREEL (guided): row % stated target, given what was caught")
```

By month, for the two ambiguous groups:

```{r}
tab("CREEL (guided): steelhead-only and no-catch parties")
```

## What each counting rule would give

Guided salmon angler-trips on the Cowlitz under alternative rules. The current
pipeline uses **C**; **D** adds steelhead-only trips in September–November.

```{r}
tab("RULE OPTIONS")
```
