---
title: "Briefing: Freshwater Recreational Effort by Mode & Location"
subtitle: "Preparing for PST Economic Study Consultant Meeting"
author: "WDFW Freshwater Creel"
date: today
execute:
  warning: false
  message: false
---

## Setup

```{r}
#| label: setup

library(creelutils)
library(tidyverse)
library(readxl)
library(hms)
library(glue)
library(gt)
library(here)
library(DT)

# Create output directory for CSV exports
output_dir <- here::here("output")
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

# Cache directory for raw data pulls
cache_dir <- here::here(".cache")
if (!dir.exists(cache_dir)) dir.create(cache_dir, recursive = TRUE)
```

## Background & Consultant Request

The U.S. has contracted with an economics consultant to assess the economic
value of PST (Pacific Salmon Treaty) salmon fisheries. Commercial catch data
and marine recreational fishing effort data are readily accessible, but
**freshwater recreational effort** is more challenging to produce at the
resolution requested.

### What the Consultant Needs

For **2022 through 2025**, the consultant has requested effort data in the
following format:

| Field | Description |
|-------|-------------|
| **Year** | Calendar year (2022–2025) |
| **River** | Individual river/system |
| **Mode** | Private (unguided) vs. Guided |
| **Location** | Boat vs. Bank |
| **Angler Days** | Total estimated angler trips |

### Data Available Today

**CRC-based freshwater trip estimates** already exist as part of the NEPA
analysis for our annual fishing package. The NEPA spreadsheet
(`NEPA PS_Recreational Effort Estimates 2025_2026_for 2026_2027 Projections_4_21_2026.xlsx`)
produces annual trip estimates by river and month using CRC salmon harvest
multiplied by a trips-per-salmon-caught ratio (per 2013 NOAA memo).
Two ratios were originally established:

| Year Type | Ratio | Rationale |
|-----------|-------|-----------|
| **Odd years** (2021, 2023, 2025, 2027…) | **3.44** trips/salmon | Abundant Puget Sound pink salmon (a strongly odd-year dominant run) mean anglers catch many salmon per trip — so fewer trips are needed per salmon caught |
| **Even years** (2020, 2022, 2024, 2026…) | **8.65** trips/salmon | Pink salmon are nearly absent in even years; anglers rely on Chinook and coho alone, requiring ~2.5× more trips per salmon caught |

However, **the "Even Effort" output column is blank (NA) in all available
projection sheets** (2018 through 2026), and the current Summary tab
(cell A15) explicitly documents that only the **odd-year ratio (3.44) is
applied to all projections** regardless of whether the target year is odd
or even.

Using 3.44 for an even-year estimate (when 8.65 would be appropriate)
understates effort by approximately **60%** (3.44 ÷ 8.65 ≈ 0.40). This
affects any 2022 and 2024 estimates provided to the consultant.
These estimates cover ~54 Puget Sound freshwater systems but are reported
as **total trips only** — they are not broken out by mode (guided/unguided)
or location (bank/boat).

::: {.callout-warning}
**Action item:** Confirm with the analyst whether the even-year ratio (8.65)
should be reinstated for 2022 and 2024 estimates. Applying only the
odd-year ratio (3.44) to even years likely understates freshwater effort by
~60% for those years — a potentially significant issue for the consultant's
economic valuation. Determine whether this was an intentional methodological
decision and, if so, document the justification.
:::

**Creel interview data** from WDFW's internal database contains
party-level records with guided status and bank/boat classification. These
interviews do not cover all rivers, and sample sizes vary, but they provide
the best available empirical basis for estimating the *proportional split*
by mode and location.

### Proposed Approach

Apply creel-interview-derived proportions (guided/unguided × bank/boat) to
the CRC-based total trip estimates to produce the breakdown the consultant
needs:
$$
\text{Trips}_{river,\, mode,\, location} = \text{Total Trips}_{river} \times \hat{p}_{river,\, mode,\, location}
$$

Where $\hat{p}$ is the proportion of interviews in that combined category
(e.g., Guided-Boat = 15% of interviews at that river → 15% of estimated
trips are assigned to Guided-Boat).

**Key considerations for the meeting:**

1. **Coverage gap** — Not all rivers with CRC trip estimates have creel
   interview data. For rivers without creel coverage, options include
   applying a regional/area-level proportion or a weighted average from
   similar rivers.
2. **Temporal resolution** — Proportions can be applied at the fishery-wide
   level or monthly. The variability analysis below quantifies how much
   proportions shift over time.
3. **Years requested** — The consultant wants 2022–2025. The NEPA
   spreadsheet currently has 2026/2027 projections and historical base
   years. We may need to reconstruct or locate the 2022–2025 estimates.
4. **Species scope** — PST salmon only (excludes steelhead).

The remainder of this document derives the proportions from creel data,
assesses their stability, and demonstrates the application approach using
the 2026 NEPA trip projections as a worked example.

### Open Issues: CRC Trip Estimate Methodology

The following issues with the **CRC-based trip estimates themselves** (i.e.,
the total trip numbers before any proportional split is applied) must be
resolved before final estimates are delivered to the consultant.

| # | Issue | Impact | Action Required |
|---|-------|--------|-----------------|
| 1 | **Even-year trips/salmon ratio never applied** — The NEPA workbook contains columns for the even-year ratio (8.65) but the "Even Effort" output is blank in all available sheets (2018–2026). Only the odd-year ratio (3.44) is used for all projections per Summary tab methodology note (cell A15). | Freshwater effort for even years (2022, 2024) is likely **understated by ~60%** (3.44 ÷ 8.65 ≈ 0.40). These are two of the four years the consultant requested. | Confirm with analyst whether use of 3.44 for even years was intentional or an oversight. If not intentional, regenerate 2022 and 2024 estimates using 8.65. Document the decision either way. |
| 2 | **Historical estimates for 2022–2025 not in hand** — The current NEPA workbook projects 2026/2027 and includes historical base-year data, but year-specific final estimates for 2022–2025 are not yet located. | Cannot deliver to consultant without these. | Locate previously-reported NEPA estimates for 2022–2025 seasons, or re-run the projection logic for those years. |

### Open Issues: Applying Creel Proportions to CRC Estimates

The CRC trip estimate methodology does **not** use creel data — it derives
total freshwater trips from CRC card counts and a trips-per-salmon ratio.
The issues below concern how we **apply creel-derived proportions** to those
totals to produce the mode × location breakdown the consultant needs.

| # | Issue | Impact | Action Required |
|---|-------|--------|-----------------|
| 3 | **River-to-creel crosswalk not yet established** — Matching CRC rivers (e.g., "SKAGIT R.") to creel fisheries (e.g., "Skagit River Summer Chinook") is non-trivial; rivers may correspond to multiple fisheries (different species/runs) or have no creel coverage at all. | Cannot apply river-specific proportions without a validated crosswalk. The worked example below uses global (pooled) proportions as an interim approach. | Develop and manually verify a river-to-fishery crosswalk, prioritizing high-effort rivers. |
| 4 | **Rivers without creel coverage rely on pooled proportions** — Not all CRC rivers have been surveyed by creel programs. For these, the only option is a pooled or regional average proportion. | Pooled proportions may be poorly suited to specific systems (e.g., a predominantly guided fishery assigned an average with low guided share). | Evaluate regional (Catch Area) proportions as a more defensible fallback than an all-fisheries average. |

## Methods

This analysis quantifies the proportions of **guided vs. unguided** and
**bank vs. boat** anglers sampled during freshwater creel interviews across
all WDFW fisheries with assigned `fishery_name` values. A third combined
classification (**guided-bank, guided-boat, unguided-bank, unguided-boat**)
provides the joint proportions most directly applicable to expanding
estimated angler trips.

### Data Source

Interview records are pulled from the WDFW PostgreSQL creel database via
`creelutils::fetch_data(data_source = "internal")`. The full set of fishery
names is obtained from `creelutils::fishery_lut()`. Each fishery's interview
table is retrieved individually and combined.

### Classification Logic

- **Guided status** is derived from the `trip_guided` field:
  "Guided" and "Non-guided" map directly; "Unknown" is treated as missing.
- **Bank/Boat** uses `angler_type` when populated (older datasets).
  For newer datasets, `boat_used == "No"` → Bank; `boat_used == "Yes"` → Boat,
  with `fish_from_boat == "Bank"` overriding to Bank for parties that had a
  boat but fished from shore.
- **Combined** crosses the two classifications into four categories:
  Guided-Bank, Guided-Boat, Unguided-Bank, Unguided-Boat.

### Proportion Metrics

Proportions are calculated two ways:

1. **Interview counts** — each interview party counts as one unit.
2. **Fishing hours** — `fishing_time × angler_count`, where fishing time
   uses `fishing_end_time` for complete trips and `interview_time` as a
   surrogate end for incomplete trips (following `prep_dwg_interview_fishing_time()`).

### Temporal Aggregation & Variability

Proportions are computed at three scales:

- **Fishery-level** (entire dataset for a fishery)
- **Monthly** (per fishery × calendar month)
- **Weekly** (per fishery × ISO week)

Variability (SD, CV, range) across units at each scale quantifies how stable
these proportions are — informing whether a single fishery-wide proportion
is sufficient or whether time-varying proportions should be applied to
estimated trips.

## Pull All Fishery Interview Data

```{r}
#| label: pull-all-interviews

cache_file <- file.path(cache_dir, "all_interviews.rds")

if (file.exists(cache_file)) {
  cat("Loading cached interview data from", cache_file, "\n")
  cat("Cache date:", format(file.mtime(cache_file), "%Y-%m-%d %H:%M"), "\n")
  all_interviews <- readRDS(cache_file)
} else {
  # Connect and get fishery names
  conn <- connect_creel_db()
  fishery_names <- fishery_lut(conn = conn) |>
    pull(fishery_name) |>
    unique()

  cat(glue("{length(fishery_names)} fisheries found in fishery_lut\n\n"))

  # Iterate over all fisheries and bind interview data
  all_interviews <- purrr::map(
    fishery_names,
    \(fn) {
      tryCatch(
        {
          dat <- fetch_data(
            conn = conn,
            fishery_name = fn,
            tables = "interview",
            data_source = "internal"
          )
          # Coerce all columns to character before binding to prevent
          # type conflicts (logical from all-NA, double vs character, etc.)
          dat$interview |>
            mutate(across(everything(), as.character))
        },
        error = function(e) {
          cli::cli_alert_warning("Failed for {fn}: {e$message}")
          NULL
        }
      )
    }
  ) |>
    purrr::compact() |>
    bind_rows()

  DBI::dbDisconnect(conn)

  # Save to cache
  saveRDS(all_interviews, cache_file)
  cat(glue("Cached {nrow(all_interviews)} rows to {cache_file}\n"))
}

cat(glue("{nrow(all_interviews)} total interview records across {n_distinct(all_interviews$fishery_name)} fisheries\n"))
```

::: {.callout-tip}
To force a fresh pull from the database, delete `.cache/all_interviews.rds`
and re-render.
:::

## Derive Classification Fields

The interview view contains:

- `trip_guided`: "Guided" / "Non-guided" / "Unknown" (newer fisheries)
- `angler_type`: "Bank" / "Boat" / "Unk" (older fisheries, direct classification)
- `boat_used`: "Yes" / "No" (newer fisheries)
- `fish_from_boat`: "Boat" / "Bank" (supplement for parties with boats)
- `boat_type`: vessel detail (Pontoon/Kayak/Drift Boat/Motorized)

Bank/boat derivation follows the logic in
[`prep_dwg_interview_angler_types()`](https://github.com/wdfw-fp/CreelEstimates/blob/main/R_functions/prep_dwg_interview_angler_types.R):
older datasets use `angler_type` directly; newer datasets derive from `boat_used`.

Fishing hours follow
[`prep_dwg_interview_fishing_time()`](https://github.com/wdfw-fp/CreelEstimates/blob/main/R_functions/prep_dwg_interview_fishing_time.R):
for incomplete trips or missing end times, `interview_time` substitutes as
the end time. Total fishing hours = per-party fishing time × angler count.

```{r}
#| label: derive-fields

# Helper: safely parse HH:MM:SS strings to hms; returns NA on failure
safe_hms <- function(x) {
  out <- suppressWarnings(hms::as_hms(x))
  out[is.nan(as.numeric(out))] <- NA
  out
}

interviews <- all_interviews |>
  # Normalize sentinel strings ("NA", "", whitespace-only) to proper NA
  mutate(across(everything(), ~ if_else(.x %in% c("NA", "", " "), NA_character_, .x))) |>
  mutate(
    event_date = as.Date(event_date),
    year = year(event_date),
    month = floor_date(event_date, "month"),
    week = floor_date(event_date, "week"),
    # Guided status: standardize trip_guided to binary
    guided_status = case_when(
      trip_guided == "Guided"     ~ "Guided",
      trip_guided == "Non-guided" ~ "Unguided",
      TRUE                        ~ NA_character_
    ),
    # Bank/Boat: use angler_type if available, otherwise derive from boat_used
    angler_final = case_when(
      # Older datasets with direct angler_type classification
      !is.na(angler_type) & angler_type == "Bank" ~ "Bank",
      !is.na(angler_type) & angler_type == "Boat" ~ "Boat",
      # Newer datasets: derive from boat_used
      boat_used == "No"  ~ "Bank",
      boat_used == "Yes" & fish_from_boat == "Bank" ~ "Bank",
      boat_used == "Yes" ~ "Boat",
      TRUE ~ NA_character_
    ),
    # Fishing time (hours) per CreelEstimates methodology:
    # Use fishing_end_time for complete trips; interview_time for incomplete/missing
    trip_status = tidyr::replace_na(trip_status, "Unknown"),
    end_time_final = if_else(
      trip_status == "Incomplete" | is.na(fishing_end_time),
      interview_time,
      fishing_end_time
    ),
    # Safe time parsing — returns NA for malformed/empty strings
    fishing_time = as.numeric(
      safe_hms(end_time_final) - safe_hms(fishing_start_time)
    ) / 3600,
    # Angler count: treat <= 0 as missing (e.g., -1 means "not recorded")
    angler_count_num = as.numeric(angler_count),
    angler_count_num = if_else(angler_count_num <= 0, NA_real_, angler_count_num),
    # Total fishing hours = fishing_time * angler_count
    fishing_time_total = fishing_time * angler_count_num
  ) |>
  # Drop rows with non-positive or missing fishing time
  mutate(
    fishing_time_total = if_else(
      is.na(fishing_time_total) | fishing_time_total <= 0,
      NA_real_,
      fishing_time_total
    )
  )

# Summary of field coverage
cat("Guided status coverage:\n")
interviews |> count(guided_status) |> print()
cat("\nBank/Boat coverage:\n")
interviews |> count(angler_final) |> print()
cat("\nFishing time total (hours) summary:\n")
summary(interviews$fishing_time_total) |> print()
```

## Fisheries Included

```{r}
#| label: fishery-listing

fishery_summary <- interviews |>
  group_by(fishery_name) |>
  summarise(
    n_interviews = n(),
    date_range = paste(min(event_date, na.rm = TRUE), "to", max(event_date, na.rm = TRUE)),
    pct_guided_recorded = round(100 * mean(!is.na(guided_status)), 1),
    pct_bank_boat_recorded = round(100 * mean(!is.na(angler_final)), 1),
    pct_hours_available = round(100 * mean(!is.na(fishing_time_total)), 1),
    .groups = "drop"
  ) |>
  arrange(fishery_name)

cat(glue("{nrow(fishery_summary)} fisheries included in analysis\n"))
```

```{r}
#| label: tbl-fishery-list

fishery_summary |>
  datatable(
    caption = "Fisheries included in analysis (searchable)",
    filter = "top",
    rownames = FALSE,
    options = list(pageLength = 15, autoWidth = TRUE),
    colnames = c(
      "Fishery" = "fishery_name",
      "Interviews" = "n_interviews",
      "Date Range" = "date_range",
      "% Guided/Unguided Recorded" = "pct_guided_recorded",
      "% Bank/Boat Recorded" = "pct_bank_boat_recorded",
      "% Hours Available" = "pct_hours_available"
    )
  )
```

## Calculate Proportions (Interview Counts)

### Fishery-Level Proportions (Counts)

```{r}
#| label: fishery-level-proportions

# Guided vs Unguided proportions by fishery
guide_fishery <- interviews |>
  filter(!is.na(guided_status)) |>
  count(fishery_name, guided_status) |>
  group_by(fishery_name) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

# Bank vs Boat proportions by fishery
type_fishery <- interviews |>
  filter(!is.na(angler_final)) |>
  count(fishery_name, angler_final) |>
  group_by(fishery_name) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

# Export
write_csv(guide_fishery, file.path(output_dir, "guide_proportions_fishery.csv"))
write_csv(type_fishery, file.path(output_dir, "type_proportions_fishery.csv"))
```

### Monthly Proportions (Counts)

```{r}
#| label: monthly-proportions

guide_monthly <- interviews |>
  filter(!is.na(guided_status)) |>
  count(fishery_name, month, guided_status) |>
  group_by(fishery_name, month) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

type_monthly <- interviews |>
  filter(!is.na(angler_final)) |>
  count(fishery_name, month, angler_final) |>
  group_by(fishery_name, month) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

# Export
write_csv(guide_monthly, file.path(output_dir, "guide_proportions_monthly.csv"))
write_csv(type_monthly, file.path(output_dir, "type_proportions_monthly.csv"))
```

### Weekly Proportions (Counts)

```{r}
#| label: weekly-proportions

guide_weekly <- interviews |>
  filter(!is.na(guided_status)) |>
  count(fishery_name, week, guided_status) |>
  group_by(fishery_name, week) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

type_weekly <- interviews |>
  filter(!is.na(angler_final)) |>
  count(fishery_name, week, angler_final) |>
  group_by(fishery_name, week) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

# Export
write_csv(guide_weekly, file.path(output_dir, "guide_proportions_weekly.csv"))
write_csv(type_weekly, file.path(output_dir, "type_proportions_weekly.csv"))
```

## Calculate Proportions (Combined: Guide × Bank/Boat)

Cross-classification of guided status and angler type into four categories
most applicable to expanding estimated angler trips.

```{r}
#| label: combined-derive

# Create combined category
interviews <- interviews |>
  mutate(
    combined_type = case_when(
      !is.na(guided_status) & !is.na(angler_final) ~
        paste(guided_status, angler_final, sep = "-"),

      TRUE ~ NA_character_
    )
  )

cat("Combined type coverage:\n")
interviews |> count(combined_type) |> print()
```

### Fishery-Level (Combined, Counts)

```{r}
#| label: combined-fishery-counts

combined_fishery <- interviews |>
  filter(!is.na(combined_type)) |>
  count(fishery_name, combined_type) |>
  complete(fishery_name, combined_type, fill = list(n = 0L)) |>
  group_by(fishery_name) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

write_csv(combined_fishery, file.path(output_dir, "combined_proportions_fishery.csv"))
```

### Monthly (Combined, Counts)

```{r}
#| label: combined-monthly-counts

combined_monthly <- interviews |>
  filter(!is.na(combined_type)) |>
  count(fishery_name, month, combined_type) |>
  complete(nesting(fishery_name, month), combined_type, fill = list(n = 0L)) |>
  group_by(fishery_name, month) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

write_csv(combined_monthly, file.path(output_dir, "combined_proportions_monthly.csv"))
```

### Weekly (Combined, Counts)

```{r}
#| label: combined-weekly-counts

combined_weekly <- interviews |>
  filter(!is.na(combined_type)) |>
  count(fishery_name, week, combined_type) |>
  complete(nesting(fishery_name, week), combined_type, fill = list(n = 0L)) |>
  group_by(fishery_name, week) |>
  mutate(total = sum(n), prop = n / total) |>
  ungroup()

write_csv(combined_weekly, file.path(output_dir, "combined_proportions_weekly.csv"))
```

### Fishery-Level (Combined, Hours)

```{r}
#| label: combined-fishery-hrs

combined_fishery_hrs <- interviews |>
  filter(!is.na(combined_type), !is.na(fishing_time_total)) |>
  group_by(fishery_name, combined_type) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  complete(fishery_name, combined_type, fill = list(hours = 0)) |>
  group_by(fishery_name) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

write_csv(combined_fishery_hrs, file.path(output_dir, "combined_proportions_fishery_hrs.csv"))
```

### Monthly (Combined, Hours)

```{r}
#| label: combined-monthly-hrs

combined_monthly_hrs <- interviews |>
  filter(!is.na(combined_type), !is.na(fishing_time_total)) |>
  group_by(fishery_name, month, combined_type) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  complete(nesting(fishery_name, month), combined_type, fill = list(hours = 0)) |>
  group_by(fishery_name, month) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

write_csv(combined_monthly_hrs, file.path(output_dir, "combined_proportions_monthly_hrs.csv"))
```

### Weekly (Combined, Hours)

```{r}
#| label: combined-weekly-hrs

combined_weekly_hrs <- interviews |>
  filter(!is.na(combined_type), !is.na(fishing_time_total)) |>
  group_by(fishery_name, week, combined_type) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  complete(nesting(fishery_name, week), combined_type, fill = list(hours = 0)) |>
  group_by(fishery_name, week) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

write_csv(combined_weekly_hrs, file.path(output_dir, "combined_proportions_weekly_hrs.csv"))
```

### Searchable Fishery Proportion Table (Combined)

```{r}
#| label: tbl-combined-searchable

# Wide-format summary: one row per fishery with proportions for each combined type
combined_wide <- combined_fishery |>
  select(fishery_name, combined_type, n, total, prop) |>
  pivot_wider(
    names_from = combined_type,
    values_from = c(n, prop),
    names_glue = "{combined_type}_{.value}"
  ) |>
  rename(total_interviews = total) |>
  distinct()

# Also join hours-based proportions
combined_wide_hrs <- combined_fishery_hrs |>
  select(fishery_name, combined_type, prop) |>
  pivot_wider(
    names_from = combined_type,
    values_from = prop,
    names_prefix = "hrs_"
  )

combined_summary <- combined_wide |>
  left_join(combined_wide_hrs, by = "fishery_name")

write_csv(combined_summary, file.path(output_dir, "combined_proportions_summary.csv"))

combined_summary |>
  datatable(
    caption = "Combined guided × bank/boat proportions by fishery (searchable)",
    filter = "top",
    rownames = FALSE,
    options = list(pageLength = 15, scrollX = TRUE, autoWidth = TRUE)
  ) |>
  formatPercentage(columns = grep("prop|hrs_", names(combined_summary), value = TRUE), digits = 1)
```

## Calculate Proportions (Fishing Hours)

Proportions based on total angler fishing hours (fishing_time × angler_count)
per [`prep_dwg_interview_fishing_time()`](https://github.com/wdfw-fp/CreelEstimates/blob/main/R_functions/prep_dwg_interview_fishing_time.R).

### Fishery-Level Proportions (Hours)

```{r}
#| label: fishery-level-proportions-hrs

# Guided vs Unguided by total fishing hours
guide_fishery_hrs <- interviews |>
  filter(!is.na(guided_status), !is.na(fishing_time_total)) |>
  group_by(fishery_name, guided_status) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  group_by(fishery_name) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

# Bank vs Boat by total fishing hours
type_fishery_hrs <- interviews |>
  filter(!is.na(angler_final), !is.na(fishing_time_total)) |>
  group_by(fishery_name, angler_final) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  group_by(fishery_name) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

# Export
write_csv(guide_fishery_hrs, file.path(output_dir, "guide_proportions_fishery_hrs.csv"))
write_csv(type_fishery_hrs, file.path(output_dir, "type_proportions_fishery_hrs.csv"))
```

### Monthly Proportions (Hours)

```{r}
#| label: monthly-proportions-hrs

guide_monthly_hrs <- interviews |>
  filter(!is.na(guided_status), !is.na(fishing_time_total)) |>
  group_by(fishery_name, month, guided_status) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  group_by(fishery_name, month) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

type_monthly_hrs <- interviews |>
  filter(!is.na(angler_final), !is.na(fishing_time_total)) |>
  group_by(fishery_name, month, angler_final) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  group_by(fishery_name, month) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

# Export
write_csv(guide_monthly_hrs, file.path(output_dir, "guide_proportions_monthly_hrs.csv"))
write_csv(type_monthly_hrs, file.path(output_dir, "type_proportions_monthly_hrs.csv"))
```

### Weekly Proportions (Hours)

```{r}
#| label: weekly-proportions-hrs

guide_weekly_hrs <- interviews |>
  filter(!is.na(guided_status), !is.na(fishing_time_total)) |>
  group_by(fishery_name, week, guided_status) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  group_by(fishery_name, week) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

type_weekly_hrs <- interviews |>
  filter(!is.na(angler_final), !is.na(fishing_time_total)) |>
  group_by(fishery_name, week, angler_final) |>
  summarise(hours = sum(fishing_time_total), .groups = "drop") |>
  group_by(fishery_name, week) |>
  mutate(total_hours = sum(hours), prop = hours / total_hours) |>
  ungroup()

# Export
write_csv(guide_weekly_hrs, file.path(output_dir, "guide_proportions_weekly_hrs.csv"))
write_csv(type_weekly_hrs, file.path(output_dir, "type_proportions_weekly_hrs.csv"))
```

## Variability Analysis

Compare the spread of proportions when calculated at different temporal
aggregations (entire fishery vs monthly vs weekly).

```{r}
#| label: variability-summary

# Helper to safely summarise proportions (returns NA row if no data)
summarise_variability <- function(df, scale_label) {
  if (nrow(df) == 0) {
    return(tibble(
      scale = scale_label, mean_prop = NA_real_, sd_prop = NA_real_,
      cv = NA_real_, min_prop = NA_real_, max_prop = NA_real_, n_units = 0L
    ))
  }
  df |> summarise(
    scale = scale_label,
    mean_prop = mean(prop, na.rm = TRUE),
    sd_prop = sd(prop, na.rm = TRUE),
    cv = sd_prop / mean_prop,
    min_prop = min(prop, na.rm = TRUE),
    max_prop = max(prop, na.rm = TRUE),
    n_units = n()
  )
}

# Guided proportion variability across scales
guide_variability <- bind_rows(
  guide_fishery |> filter(guided_status == "Guided") |> summarise_variability("Fishery (overall)"),
  guide_monthly |> filter(guided_status == "Guided") |> summarise_variability("Monthly"),
  guide_weekly |> filter(guided_status == "Guided") |> summarise_variability("Weekly")
)

# Boat proportion variability across scales
type_variability <- bind_rows(
  type_fishery |> filter(angler_final == "Boat") |> summarise_variability("Fishery (overall)"),
  type_monthly |> filter(angler_final == "Boat") |> summarise_variability("Monthly"),
  type_weekly |> filter(angler_final == "Boat") |> summarise_variability("Weekly")
)
```

### Variability Tables

```{r}
#| label: tbl-guide-variability
#| tbl-cap: "Variability in guided angler proportion across temporal scales"

guide_variability |>
  gt() |>
  fmt_number(columns = c(mean_prop, sd_prop, cv, min_prop, max_prop), decimals = 3) |>
  cols_label(
    scale = "Scale",
    mean_prop = "Mean Proportion",
    sd_prop = "SD",
    cv = "CV",
    min_prop = "Min",
    max_prop = "Max",
    n_units = "N (units)"
  )
```

```{r}
#| label: tbl-type-variability
#| tbl-cap: "Variability in boat angler proportion across temporal scales"

type_variability |>
  gt() |>
  fmt_number(columns = c(mean_prop, sd_prop, cv, min_prop, max_prop), decimals = 3) |>
  cols_label(
    scale = "Scale",
    mean_prop = "Mean Proportion",
    sd_prop = "SD",
    cv = "CV",
    min_prop = "Min",
    max_prop = "Max",
    n_units = "N (units)"
  )

# Export variability summaries
write_csv(guide_variability, file.path(output_dir, "guide_variability_summary.csv"))
write_csv(type_variability, file.path(output_dir, "type_variability_summary.csv"))
```

### Variability Tables (Fishing Hours)

```{r}
#| label: variability-summary-hrs

# Guided proportion variability by fishing hours
guide_variability_hrs <- bind_rows(
  guide_fishery_hrs |> filter(guided_status == "Guided") |> summarise_variability("Fishery (overall)"),
  guide_monthly_hrs |> filter(guided_status == "Guided") |> summarise_variability("Monthly"),
  guide_weekly_hrs |> filter(guided_status == "Guided") |> summarise_variability("Weekly")
)

# Boat proportion variability by fishing hours
type_variability_hrs <- bind_rows(
  type_fishery_hrs |> filter(angler_final == "Boat") |> summarise_variability("Fishery (overall)"),
  type_monthly_hrs |> filter(angler_final == "Boat") |> summarise_variability("Monthly"),
  type_weekly_hrs |> filter(angler_final == "Boat") |> summarise_variability("Weekly")
)
```

```{r}
#| label: tbl-guide-variability-hrs
#| tbl-cap: "Variability in guided angler proportion (fishing hours) across temporal scales"

guide_variability_hrs |>
  gt() |>
  fmt_number(columns = c(mean_prop, sd_prop, cv, min_prop, max_prop), decimals = 3) |>
  cols_label(
    scale = "Scale",
    mean_prop = "Mean Proportion",
    sd_prop = "SD",
    cv = "CV",
    min_prop = "Min",
    max_prop = "Max",
    n_units = "N (units)"
  )
```

```{r}
#| label: tbl-type-variability-hrs
#| tbl-cap: "Variability in boat angler proportion (fishing hours) across temporal scales"

type_variability_hrs |>
  gt() |>
  fmt_number(columns = c(mean_prop, sd_prop, cv, min_prop, max_prop), decimals = 3) |>
  cols_label(
    scale = "Scale",
    mean_prop = "Mean Proportion",
    sd_prop = "SD",
    cv = "CV",
    min_prop = "Min",
    max_prop = "Max",
    n_units = "N (units)"
  )

# Export
write_csv(guide_variability_hrs, file.path(output_dir, "guide_variability_summary_hrs.csv"))
write_csv(type_variability_hrs, file.path(output_dir, "type_variability_summary_hrs.csv"))
```

### Counts vs Hours Comparison

```{r}
#| label: tbl-counts-vs-hours
#| tbl-cap: "Comparing CV of proportions: interview counts vs fishing hours"

bind_rows(
  guide_variability |> mutate(metric = "Guided (counts)", .before = 1),
  guide_variability_hrs |> mutate(metric = "Guided (hours)", .before = 1),
  type_variability |> mutate(metric = "Boat (counts)", .before = 1),
  type_variability_hrs |> mutate(metric = "Boat (hours)", .before = 1)
) |>
  select(metric, scale, mean_prop, cv, n_units) |>
  gt() |>
  fmt_number(columns = c(mean_prop, cv), decimals = 3) |>
  cols_label(
    metric = "Metric",
    scale = "Scale",
    mean_prop = "Mean Proportion",
    cv = "CV",
    n_units = "N"
  )
```

### Combined (Guide × Bank/Boat) Variability

```{r}
#| label: variability-combined

# Variability for each combined category across scales
combined_categories <- c("Guided-Bank", "Guided-Boat", "Unguided-Bank", "Unguided-Boat")

combined_variability <- purrr::map_dfr(combined_categories, \(cat) {
  bind_rows(
    combined_fishery |> filter(combined_type == cat) |> summarise_variability("Fishery (overall)"),
    combined_monthly |> filter(combined_type == cat) |> summarise_variability("Monthly"),
    combined_weekly |> filter(combined_type == cat) |> summarise_variability("Weekly")
  ) |> mutate(category = cat, .before = 1)
})

combined_variability_hrs <- purrr::map_dfr(combined_categories, \(cat) {
  bind_rows(
    combined_fishery_hrs |> filter(combined_type == cat) |> summarise_variability("Fishery (overall)"),
    combined_monthly_hrs |> filter(combined_type == cat) |> summarise_variability("Monthly"),
    combined_weekly_hrs |> filter(combined_type == cat) |> summarise_variability("Weekly")
  ) |> mutate(category = cat, .before = 1)
})

write_csv(combined_variability, file.path(output_dir, "combined_variability_summary.csv"))
write_csv(combined_variability_hrs, file.path(output_dir, "combined_variability_summary_hrs.csv"))
```

```{r}
#| label: tbl-combined-variability
#| tbl-cap: "Variability in combined guide × bank/boat proportions (counts)"

combined_variability |>
  gt(groupname_col = "category") |>
  fmt_number(columns = c(mean_prop, sd_prop, cv, min_prop, max_prop), decimals = 3) |>
  cols_label(
    scale = "Scale",
    mean_prop = "Mean Proportion",
    sd_prop = "SD",
    cv = "CV",
    min_prop = "Min",
    max_prop = "Max",
    n_units = "N (units)"
  )
```

```{r}
#| label: tbl-combined-variability-hrs
#| tbl-cap: "Variability in combined guide × bank/boat proportions (fishing hours)"

combined_variability_hrs |>
  gt(groupname_col = "category") |>
  fmt_number(columns = c(mean_prop, sd_prop, cv, min_prop, max_prop), decimals = 3) |>
  cols_label(
    scale = "Scale",
    mean_prop = "Mean Proportion",
    sd_prop = "SD",
    cv = "CV",
    min_prop = "Min",
    max_prop = "Max",
    n_units = "N (units)"
  )
```

## Fishery Sample Size Breakdown

A per-fishery diagnostic table showing interview counts by category to help
validate proportions and identify fisheries with small or imbalanced samples.

::: {.callout-warning}
**Data collection procedures vary by fishery and need further review.**
Some fisheries show patterns suggesting field-level recording conventions
that inflate the "Unknown" category. For example, Snohomish Fall Salmon
(2023) has nearly all unguided interviews classified as unknown guided
status — it is possible samplers left the guided field blank to imply
unguided rather than explicitly recording "Non-guided." These
fishery-specific data entry issues need to be resolved with project leads
before finalizing proportions for delivery.
:::

```{r}
#| label: fishery-sample-breakdown

fishery_breakdown <- interviews |>
  group_by(fishery_name) |>
  summarise(
    n_total = n(),
    n_guided = sum(guided_status == "Guided", na.rm = TRUE),
    n_unguided = sum(guided_status == "Unguided", na.rm = TRUE),
    n_guide_unknown = sum(is.na(guided_status)),
    n_bank = sum(angler_final == "Bank", na.rm = TRUE),
    n_boat = sum(angler_final == "Boat", na.rm = TRUE),
    n_type_unknown = sum(is.na(angler_final)),
    n_guided_bank = sum(combined_type == "Guided-Bank", na.rm = TRUE),
    n_guided_boat = sum(combined_type == "Guided-Boat", na.rm = TRUE),
    n_unguided_bank = sum(combined_type == "Unguided-Bank", na.rm = TRUE),
    n_unguided_boat = sum(combined_type == "Unguided-Boat", na.rm = TRUE),
    n_combined_unknown = sum(is.na(combined_type)),
    pct_classifiable = round(100 * (1 - n_combined_unknown / n_total), 1),
    mean_hours = round(mean(fishing_time_total, na.rm = TRUE), 2),
    median_hours = round(median(fishing_time_total, na.rm = TRUE), 2),
    total_hours = round(sum(fishing_time_total, na.rm = TRUE), 1),
    year_min = suppressWarnings(min(year(event_date), na.rm = TRUE)),
    year_max = suppressWarnings(max(year(event_date), na.rm = TRUE)),
    n_years = n_distinct(year(event_date), na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    year_min = if_else(is.infinite(year_min), NA_real_, year_min),
    year_max = if_else(is.infinite(year_max), NA_real_, year_max),
    years_span = if_else(
      is.na(year_min),
      "No dates",
      paste0(year_min, "-", year_max, " (", n_years, " yrs)")
    ),
    guide_ratio = paste0(n_guided, ":", n_unguided),
    type_ratio = paste0(n_bank, ":", n_boat)
  ) |>
  arrange(desc(n_total))

write_csv(fishery_breakdown, file.path(output_dir, "fishery_sample_breakdown.csv"))
```

```{r}
#| label: tbl-fishery-breakdown

fishery_breakdown |>
  select(
    fishery_name, n_total, years_span,
    n_guided, n_unguided, n_guide_unknown, guide_ratio,
    n_bank, n_boat, n_type_unknown, type_ratio,
    n_guided_bank, n_guided_boat, n_unguided_bank, n_unguided_boat,
    pct_classifiable, mean_hours, median_hours, total_hours
  ) |>
  datatable(
    caption = "Fishery sample size breakdown by category (searchable, sortable)",
    filter = "top",
    rownames = FALSE,
    options = list(
      pageLength = 20,
      scrollX = TRUE,
      autoWidth = FALSE,
      columnDefs = list(
        list(className = "dt-right", targets = 1:18)
      )
    ),
    colnames = c(
      "Fishery" = "fishery_name",
      "Total N" = "n_total",
      "Years" = "years_span",
      "Guided" = "n_guided",
      "Unguided" = "n_unguided",
      "Unknown (Guide)" = "n_guide_unknown",
      "Guided:Unguided" = "guide_ratio",
      "Bank" = "n_bank",
      "Boat" = "n_boat",
      "Unknown (Type)" = "n_type_unknown",
      "Bank:Boat" = "type_ratio",
      "G-Bank" = "n_guided_bank",
      "G-Boat" = "n_guided_boat",
      "UG-Bank" = "n_unguided_bank",
      "UG-Boat" = "n_unguided_boat",
      "% Classifiable" = "pct_classifiable",
      "Mean Hrs" = "mean_hours",
      "Median Hrs" = "median_hours",
      "Total Hrs" = "total_hours"
    )
  ) |>
  formatStyle("% Classifiable",
    backgroundColor = styleInterval(c(50, 80), c("#f8d7da", "#fff3cd", "#d4edda"))
  )
```

::: {.callout-note}
**Reading this table:**

- **G:UG** = Guided:Unguided count ratio. A fishery showing `0:500` is entirely unguided.
- **B:Bt** = Bank:Boat count ratio.
- **% Classifiable** = percent of interviews where both guided status AND bank/boat are known (usable for the combined 4-way split). Color-coded: red < 50%, yellow 50-80%, green > 80%.
- **Mean/Median Hrs** = average interview fishing hours. Large discrepancies suggest outliers.
- Fisheries with very small N or low % Classifiable should be interpreted with caution.
:::

```{r}
#| label: tbl-flags

# Flag fisheries with potential data quality concerns
flags <- fishery_breakdown |>
  filter(n_total < 30 | pct_classifiable < 50) |>
  select(fishery_name, n_total, pct_classifiable, years_span, guide_ratio, type_ratio) |>
  arrange(pct_classifiable)

if (nrow(flags) > 0) {
  cat(glue("{nrow(flags)} fisheries flagged for small sample size (N<30) or low classifiability (<50%):\n\n"))
  flags |>
    datatable(
      caption = "Flagged fisheries: small N or low classifiability",
      rownames = FALSE,
      options = list(pageLength = 50, dom = "t"),
      colnames = c(
        "Fishery" = "fishery_name",
        "Total N" = "n_total",
        "% Classifiable" = "pct_classifiable",
        "Years" = "years_span",
        "Guided:Unguided" = "guide_ratio",
        "Bank:Boat" = "type_ratio"
      )
    )
} else {
  cat("No fisheries flagged — all have N >= 30 and >= 50% classifiable.\n")
}
```

## Visualizations

### Guided Proportion by Time Scale

```{r}
#| label: fig-guide-distribution
#| fig-cap: "Distribution of guided angler proportions across temporal scales"
#| fig-height: 8
#| fig-width: 10

bind_rows(
  guide_fishery |>
    filter(guided_status == "Guided") |>
    select(fishery_name, prop) |>
    mutate(scale = "Fishery"),
  guide_monthly |>
    filter(guided_status == "Guided") |>
    select(fishery_name, prop) |>
    mutate(scale = "Monthly"),
  guide_weekly |>
    filter(guided_status == "Guided") |>
    select(fishery_name, prop) |>
    mutate(scale = "Weekly")
) |>
  mutate(scale = factor(scale, levels = c("Fishery", "Monthly", "Weekly"))) |>
  ggplot(aes(x = prop, fill = scale)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~scale, ncol = 1, scales = "free_y") +
  scale_x_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    x = "Proportion Guided",
    y = "Density",
    title = "Distribution of Guided Angler Proportions",
    subtitle = "Comparing fishery-level, monthly, and weekly calculations"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

### Bank/Boat Proportion by Time Scale

```{r}
#| label: fig-type-distribution
#| fig-cap: "Distribution of boat angler proportions across temporal scales"
#| fig-height: 8
#| fig-width: 10

bind_rows(
  type_fishery |>
    filter(angler_final == "Boat") |>
    select(fishery_name, prop) |>
    mutate(scale = "Fishery"),
  type_monthly |>
    filter(angler_final == "Boat") |>
    select(fishery_name, prop) |>
    mutate(scale = "Monthly"),
  type_weekly |>
    filter(angler_final == "Boat") |>
    select(fishery_name, prop) |>
    mutate(scale = "Weekly")
) |>
  mutate(scale = factor(scale, levels = c("Fishery", "Monthly", "Weekly"))) |>
  ggplot(aes(x = prop, fill = scale)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~scale, ncol = 1, scales = "free_y") +
  scale_x_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    x = "Proportion Boat Anglers",
    y = "Density",
    title = "Distribution of Boat Angler Proportions",
    subtitle = "Comparing fishery-level, monthly, and weekly calculations"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

### Per-Fishery Time Series

```{r}
#| label: fig-guide-timeseries
#| fig-cap: "Guided proportion over time by fishery (monthly)"
#| fig-height: 10
#| fig-width: 12

# Show top 12 fisheries by interview count for readability
top_fisheries <- interviews |>
  filter(!is.na(guided_status)) |>
  count(fishery_name, sort = TRUE) |>
  slice_head(n = 12) |>
  pull(fishery_name)

guide_monthly |>
  filter(fishery_name %in% top_fisheries, guided_status == "Guided") |>
  ggplot(aes(x = month, y = prop)) +
  geom_line() +
  geom_point(aes(size = total), alpha = 0.5) +
  facet_wrap(~fishery_name, scales = "free_x", ncol = 3) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    x = NULL, y = "Proportion Guided",
    size = "Sample Size",
    title = "Monthly Guided Proportion (Top 12 Fisheries by Interview Count)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| label: fig-type-timeseries
#| fig-cap: "Boat angler proportion over time by fishery (monthly)"
#| fig-height: 10
#| fig-width: 12

top_fisheries_type <- interviews |>
  filter(!is.na(angler_final)) |>
  count(fishery_name, sort = TRUE) |>
  slice_head(n = 12) |>
  pull(fishery_name)

type_monthly |>
  filter(fishery_name %in% top_fisheries_type, angler_final == "Boat") |>
  ggplot(aes(x = month, y = prop)) +
  geom_line() +
  geom_point(aes(size = total), alpha = 0.5) +
  facet_wrap(~fishery_name, scales = "free_x", ncol = 3) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    x = NULL, y = "Proportion Boat Anglers",
    size = "Sample Size",
    title = "Monthly Boat Proportion (Top 12 Fisheries by Interview Count)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Counts vs Hours: Per-Fishery Comparison

```{r}
#| label: fig-counts-vs-hours-guided
#| fig-cap: "Guided proportion: interview counts vs fishing hours (fishery level)"
#| fig-height: 7
#| fig-width: 9

# Compare fishery-level guided proportions: counts vs hours
compare_guide <- guide_fishery |>
  filter(guided_status == "Guided") |>
  select(fishery_name, prop_counts = prop) |>
  inner_join(
    guide_fishery_hrs |>
      filter(guided_status == "Guided") |>
      select(fishery_name, prop_hours = prop),
    by = "fishery_name"
  )

ggplot(compare_guide, aes(x = prop_counts, y = prop_hours)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.6) +
  scale_x_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    x = "Proportion Guided (interview counts)",
    y = "Proportion Guided (fishing hours)",
    title = "Guided Proportion: Counts vs Hours",
    subtitle = "Each point = one fishery; dashed line = 1:1"
  ) +
  theme_minimal()
```

```{r}
#| label: fig-counts-vs-hours-boat
#| fig-cap: "Boat proportion: interview counts vs fishing hours (fishery level)"
#| fig-height: 7
#| fig-width: 9

compare_type <- type_fishery |>
  filter(angler_final == "Boat") |>
  select(fishery_name, prop_counts = prop) |>
  inner_join(
    type_fishery_hrs |>
      filter(angler_final == "Boat") |>
      select(fishery_name, prop_hours = prop),
    by = "fishery_name"
  )

ggplot(compare_type, aes(x = prop_counts, y = prop_hours)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_point(alpha = 0.6) +
  scale_x_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    x = "Proportion Boat (interview counts)",
    y = "Proportion Boat (fishing hours)",
    title = "Boat Proportion: Counts vs Hours",
    subtitle = "Each point = one fishery; dashed line = 1:1"
  ) +
  theme_minimal()
```

## Worked Example: Applying Proportions to CRC Trip Estimates

This section demonstrates the proposed approach by applying creel-derived
proportions to the 2026 freshwater trip projections from the NEPA analysis
spreadsheet. This is the format the consultant has requested — total trips
broken out by river, mode (guided/private), and location (bank/boat).

### Read NEPA Trip Estimates

```{r}
#| label: read-nepa-estimates

nepa_file <- here::here(
  "docs",
  "NEPA PS_Recreational Effort Estimates 2025_2026_for 2026_2027 Projections_4_21_2026.xlsx"
)

# Read the FW Effort Projection sheet — structure is:
# Col 1 = river name (fill-down), Col 2 = month (1-12 or TOTAL)
# Col 59 = 2026 estimate, Col 61 = 2027 estimate
nepa_raw <- read_excel(
  nepa_file,
  sheet = "2026 FW Effort Projection",
  skip = 1,
  col_names = FALSE,
  .name_repair = "minimal"
)
names(nepa_raw) <- paste0("V", seq_len(ncol(nepa_raw)))

# Fill river names down and extract TOTAL rows (annual trip estimate per river)
nepa_raw <- nepa_raw |> fill(V1)
nepa_totals <- nepa_raw |>
  filter(V2 == "TOTAL") |>
  transmute(
    river = V1,
    trips_2026 = as.numeric(V59),
    trips_2027 = as.numeric(V61)
  ) |>
  filter(!is.na(river), !is.na(trips_2026) | !is.na(trips_2027))

cat(glue("{nrow(nepa_totals)} rivers with NEPA trip estimates\n"))
nepa_totals |> summarise(
  total_2026 = sum(trips_2026, na.rm = TRUE),
  total_2027 = sum(trips_2027, na.rm = TRUE)
) |> print()
```

### Apply Global Proportions to Trip Estimates

As a first pass, we apply the **overall creel-derived proportions** (pooled
across all fisheries) uniformly to the total freshwater trip estimates.
Matching river-specific creel proportions to individual CRC rivers is a
logical refinement but is deferred as a follow-on step.

```{r}
#| label: apply-proportions

# Calculate overall (all-fisheries) proportions from creel interviews
overall_props <- interviews |>
  filter(!is.na(combined_type)) |>
  count(combined_type) |>
  mutate(prop = n / sum(n))

cat("Overall creel proportions (pooled across all fisheries):\n")
overall_props |>
  mutate(pct = round(100 * prop, 1)) |>
  select(combined_type, n, prop, pct) |>
  print()

cat(glue("\nTotal classifiable interviews: {scales::comma(sum(overall_props$n))}\n"))
cat(glue("Proportions sum to: {round(sum(overall_props$prop), 4)}\n"))

# Apply to total 2026 FW trip estimate
total_trips_2026 <- sum(nepa_totals$trips_2026, na.rm = TRUE)

expanded_2026 <- overall_props |>
  mutate(
    total_trips = total_trips_2026,
    estimated_trips = total_trips_2026 * prop
  )

cat(glue("\nTotal 2026 FW trips: {scales::comma(round(total_trips_2026))}\n"))
```

### Summary: Overall Mode × Location Split (2026)

```{r}
#| label: tbl-summary-by-mode

expanded_2026 |>
  separate(combined_type, into = c("mode", "location"), sep = "-", remove = FALSE) |>
  mutate(mode = if_else(mode == "Guided", "Guided", "Private")) |>
  select(mode, location, n_interviews = n, proportion = prop, estimated_trips) |>
  mutate(
    estimated_trips = round(estimated_trips, 0),
    proportion = round(proportion, 4)
  ) |>
  gt() |>
  fmt_number(columns = c(estimated_trips, n_interviews), use_seps = TRUE, decimals = 0) |>
  fmt_number(columns = proportion, decimals = 3) |>
  cols_label(
    mode = "Mode",
    location = "Location",
    n_interviews = "Interviews (n)",
    proportion = "Proportion",
    estimated_trips = "Estimated Trips (2026)"
  ) |>
  tab_header(
    title = "2026 FW Trip Estimates by Mode × Location",
    subtitle = glue("Global creel proportions applied to {scales::comma(round(total_trips_2026))} total projected trips")
  ) |>
  grand_summary_rows(
    fns = list(Total = ~ sum(.)),
    columns = c(n_interviews, estimated_trips),
    fmt = ~ fmt_number(., decimals = 0, use_seps = TRUE)
  )
```

::: {.callout-important}
**This is a demonstration using 2026 projections.** The consultant needs
2022–2025 actuals. The same methodology can be applied once we locate or
reconstruct the historical CRC-based trip estimates for those years.
:::

::: {.callout-note}
**Follow-on refinements:**

1. **River-specific proportions** — Match individual creel-surveyed rivers
   to CRC rivers and apply river-level proportions where sample sizes are
   adequate. Use the global average as a fallback for unsampled rivers.
2. **Regional (Catch Area) proportions** — As an intermediate step between
   global and river-specific, pool creel interviews by Catch Area to
   produce area-level proportions.
3. **Year-specific proportions** — If interview sample sizes support it,
   derive proportions for each year (2022–2025) rather than pooling all
   years.
4. Marine areas (Areas 5-12) have trip estimates reported in the NEPA
   workbook through an existing process, but how those figures are produced
   is not documented in the spreadsheet and is outside our scope. For
   freshwater, no equivalent reporting process currently exists — the
   approach described in this document is intended to fill that gap and
   meet the consultant's data request.
:::
