---
title: "PST Freshwater Recreational Effort: Framework & Assembly Plan"
subtitle: "Angler trips by Year × River × Mode × Location, 2022–2025"
author: "WDFW Fish Program — Freshwater Fisheries Policy"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    embed-resources: true
execute:
  warning: false
  message: false
  echo: false
---

```{r}
#| label: setup

library(tidyverse)
library(here)
library(glue)
library(gt)

# Parent doc reads the artifacts produced by pst_fw_effort_assembly.R.
# It renders with partial outputs by design: anything missing is reported
# as a gap rather than raising an error (design rule R2).

# Paths mirror the assembly script exactly (OUT_DIR / PST_DIR). This file
# renders from analysis/; here::here() resolves from the repo root either way.
OUT_DIR <- here::here("analysis", "outputs")
PST_DIR <- here::here("input_files", "pst")

read_if <- function(path, ...) {
  if (file.exists(path)) readr::read_csv(path, show_col_types = FALSE, ...) else NULL
}

manifest    <- read_if(file.path(PST_DIR, "pst_input_manifest.csv"))
crosswalk   <- read_if(file.path(PST_DIR, "pst_river_block_crosswalk.csv"))
effort      <- read_if(file.path(OUT_DIR, "pst_fw_trips_by_mode_location_INTERMEDIATE.csv"))
effort_area <- read_if(file.path(OUT_DIR, "pst_fw_trips_by_crc_area_INTERMEDIATE.csv"))
provenance  <- read_if(file.path(OUT_DIR, "pst_fw_provenance_ledger.csv"))
gaps        <- read_if(file.path(OUT_DIR, "pst_fw_gap_register.csv"))
coverage    <- read_if(file.path(OUT_DIR, "pst_fw_crc_coverage.csv"))
crc_bias    <- read_if(file.path(OUT_DIR, "pst_fw_crc_vs_creel_bias.csv"))
```

::: {.callout-note}
## What this document is

This is the **parent** for the PST freshwater effort work. It owns the
framework — scope, the Track A tier hierarchy, the Track B derivation logic,
assumptions A1–A4, design rules R1–R5 — and the current assembly status.

Framework material previously duplicated across `PST_ANALYSIS.md`,
`PST_FW_Effort_Assembly_Plan.md`, and the Background/Methods prose in
`interview_proportions.qmd` is consolidated here. Those files become
children: `interview_proportions.qmd` shrinks to the Track B derivation it
actually performs; `PST_ANALYSIS.md` retains the run-level analysis contract.

Child documents should not restate scope, tiers, or assumptions. They cite
them (e.g. "per A2") and get on with the computation.
:::

# Purpose & deliverable contract

The U.S. side of the Pacific Salmon Commission has contracted Northern
Economics (Melissa Errend) to update the economic valuation of PST salmon
fisheries. Commercial landings and marine recreational effort are already
accessible at the needed resolution. **Freshwater recreational effort is
not** — no standing WDFW reporting process produces angler trips broken out
by mode and location. This work fills that gap.

## Requested format

| Field | Definition |
|---|---|
| **Year** | Calendar year, 2022–2025 |
| **River** | Individual river/system (rolled to block where required) |
| **Mode** | Guided vs. private (unguided) |
| **Location** | Boat vs. bank |
| **Angler trips** | Total estimated angler trips (see glossary) |

Species scope: **PST salmon, total salmon** — not species-specific.
Steelhead is excluded entirely.

## Geographic scope

| Region | In scope | Note |
|---|---|---|
| Puget Sound freshwater | Yes | CRC + creel; ~54 systems in the NEPA frame |
| WA Coast freshwater | Yes | 2025 fallback still unidentified |
| Columbia tributaries (WA) | Yes | Includes mid-Columbia R3 programs |
| Columbia mainstem | **No** | Delivered to the consultant directly by ODFW (OR+WA combined) |
| Steelhead, any region | **No** | Out of scope by definition |
| Upriver spring Chinook (Columbia) | Candidate exclusion | Not PST-managed; limited ocean harvest — pending Jim's confirmation |

## Scope decisions on record

| # | Question | Resolution | Authority |
|---|---|---|---|
| D1 | Species-specific or total salmon? | Total salmon | Jim (resolving Ryan Lothrop's PSC indicator-stock sideboard question) |
| D2 | Steelhead included? | No | Consultant request |
| D3 | Columbia mainstem | Out — ODFW delivers directly | Jim |
| D4 | Catch/trip basis for FEIS-derived rows | FEIS Table 3-19 rates acceptable; prefer empirical 2022–2025 averages if readily available | Jim, 2026-08-03 |
| D5 | Upriver spring Chinook | Proposed exclusion, **open** | Pending Jim |

# Framework

## Two-track decomposition

The deliverable is a product of two independently sourced quantities. Keeping
them separate is what makes partial coverage tractable: a river can have a
defensible total with an unknown split, or vice versa, and both states are
representable in the output.

$$
\text{Trips}_{r,y,\,mode,\,loc} \;=\; \underbrace{T_{r,y}}_{\text{Track A}} \;\times\; \underbrace{\hat p_{r,y,\,mode,\,loc}}_{\text{Track B}}
$$

- **Track A — total angler trips** per river-year. Sourced by tier hierarchy.
- **Track B — mode × location proportions.** Bank/boat comes from the design
  stratum in `est_pe_effort()`. Guided/unguided is *post-stratification within
  the boat stratum* using interview proportions.

Track B is exported as a **conditional** $P(\text{guided}\mid\text{boat})$,
not as a joint four-way proportion. The joint is reconstructed only at the
final multiply, and only where the bank/boat split is itself design-based.

## Track A — tier hierarchy

| Tier | Source | Applies where |
|---|---|---|
| **P1** | Design-based creel estimates (`est_pe_effort()` → `total_effort_hrs ÷ mean_trip_length`) | Fisheries with an executed creel PE run |
| **P2** | Regional creel-to-CRC ratios, computed **within block** | CRC-covered rivers without a PE run, inside a block that has one |
| **P3** | Published figures — JSR tables, Mitchell Act FEIS rates, NMFS 2024 trips-per-salmon | No creel signal available in-block |

Every row carries its tier. Tiers do not blend: a river-year is P1, P2, or
P3, never an average of two.

## Track B — mode × location

1. **Bank/boat** is a design stratum, not a post-hoc classification. It comes
   from the effort estimate directly, so it inherits the PE's expansion.
2. **Guided/unguided** is not a design stratum in any WDFW freshwater program.
   It is derived from interview counts within the boat stratum and applied as
   a conditional proportion.
3. Where the guided field is not collected at all (all mid-Columbia R3
   programs), the dimension is coded `"unknown"` — never zero, never
   defaulted to unguided.

## Assumptions A1–A4

::: {.callout-important}
Transcribed from `PST_ANALYSIS.md`; reconcile wording against that file
before delivery so the two do not drift.
:::

| ID | Assumption | Consequence if violated |
|---|---|---|
| **A1** | Interviewed parties are representative of effort within the stratum and period from which they were drawn. | Mode proportions biased toward whatever access points and time blocks samplers favored. |
| **A2** | Guiding occurs within the boat stratum. Guided bank effort is treated as negligible where not separately recorded. | Understates guided share on systems with walk-and-wade guiding. |
| **A3** | Mode proportions are stable within the aggregation level at which they are applied (fishery-season unless monthly sample supports finer). | Seasonal shifts in guided share get flattened onto the annual total. |
| **A4** | Where year-specific interview sample is insufficient, the conditional proportion may be borrowed from a donor **within the same block**, never across blocks. | Cross-block borrowing imports a different fishery's economics. |

## Design rules R1–R5

| ID | Rule |
|---|---|
| **R1** | Every output row carries `tier` and `source_id`. |
| **R2** | Missing inputs log a gap and continue. The pipeline does not fail on absent data. |
| **R3** | Missing dimensions are coded `"unknown"`, never zero. |
| **R4** | Ratios never cross blocks. |
| **R5** | Projected, donor-borrowed, or otherwise non-native values are labeled in `method`. |

Supporting rules: smolts are excluded from all catch groups **[S3]**;
within-season trip-length donors follow a four-tier hierarchy **[S4]** —
month → area_season → type_season → fishery_season.

## Why blocks, and why ratios stay inside them

Trips-per-salmon is not a transferable constant. At the high-catch end,
Buoy 10 runs under 1 trip per fish kept in most years (JSR Table 25 puts
kept catch/trip in the 0.7–1.3 range across 2022–2025, i.e. roughly
0.75–1.4 trips per salmon). Puget Sound freshwater runs several times
higher: James Dixon's NMFS 2024 updated rates, anchored on 2022 WDFW data,
put it at **3.539 trips/salmon (odd years) and 4.820 (even years)** —
a defensible, well-documented figure, not an outlier built on sparse catch.
Even between these two reasonably well-supported cases the ratio varies by
roughly 3–6×, which is enough on its own to rule out a single statewide
rate. (LCR summer Chinook is not used as a reference case here: its
historically cited ratio is a high outlier built on very little observed
catch and needs its own scrutiny before it's treated as representative of
anything.)
Blocks are the unit within which a ratio is assumed portable; the crosswalk in `input_files/pst/`
carries the 63-row fishery-to-block mapping that enforces R4 mechanically.

The same logic constrains the NEPA odd/even trips-per-salmon ratio
(3.44 / 8.65): that spread is driven by Puget Sound pink salmon dominance
and **does not transfer** to Columbia or coastal systems.

# Pipeline

## Registry

| Artifact | Role |
|---|---|
| `input_files/pst/pst_input_manifest.csv` | Registered inputs — one row per source file/table, with tier, block, and expected coverage |
| `input_files/pst/pst_river_block_crosswalk.csv` | Fishery → river → block mapping (63 rows) |
| `analysis/multi_fishery_creel_summary.R` | Merged effort + harvest; single DB fetch, one `est_pe_effort()` and one `est_pe_catch()` per fishery, so trips and harvest are identical by construction |
| `analysis/interview_proportions.qmd` | Track B engine — conditional mode proportions from interviews |
| `analysis/integrate_manual_runs.R` | Reconciles per-fishery `fw_creel.Rmd` runs to the monthly batch grain |
| `analysis/pst_fw_effort_assembly.R` | Reads the manifest, applies the tier hierarchy, writes deliverable + three ledgers |

## Outputs

Assembly emits the deliverable plus three ledgers, which is what lets this
document report status rather than assert it:

- **Provenance ledger** — tier, source_id, method per row
- **Gap ledger** — every skipped or absent input, with reason
- **Coverage ledger** — river-year × track completion

# Status

```{r}
#| label: status-summary

if (is.null(effort)) {
  cat("No assembly output found. Run analysis/pst_fw_effort_assembly.R.\n")
} else {
  effort |>
    count(block, year, name = "rows") |>
    pivot_wider(names_from = year, values_from = rows, values_fill = 0) |>
    gt() |>
    tab_header(title = "Rows by block and year",
               subtitle = "Deliverable blocks only; ColumbiaMainstem is out of scope")
}
```

```{r}
#| label: status-mode-unknown

# The single number that gates delivery: share of trips still carrying
# mode = "unknown". Track B cannot resolve R3_external (field not collected)
# or P2 (CRC has no mode field), so this will not reach zero by itself.
if (!is.null(effort)) {
  effort |>
    group_by(block) |>
    summarise(
      angler_trips     = sum(angler_trips, na.rm = TRUE),
      pct_mode_unknown = round(100 * sum(angler_trips[mode == "unknown"]) /
                                 sum(angler_trips), 1),
      pct_tier_p2      = round(100 * sum(angler_trips[grepl("P2", tier)]) /
                                 sum(angler_trips), 1),
      .groups = "drop"
    ) |>
    gt() |>
    fmt_number(columns = angler_trips, decimals = 0) |>
    tab_header(title = "Unresolved mode share by block",
               subtitle = "Must be understood, not necessarily zero, before delivery")
}
```

## Coverage map {#sec-map}

::: {.callout-note}
**Placeholder.** Intent: a WA state map with river systems symbolized by
assembly status — resolved (P1), resolved by ratio (P2), published-figure
fallback (P3), and open gap — with a year facet or toggle. This is the
progress tracker for the remaining outreach, and the single artifact most
worth showing Jim in a status conversation.

Requires: a river-system spatial layer (`data/spatial/`) joined to the
registry crosswalk on its key, and the coverage ledger for the status fill.
:::

```{r}
#| label: coverage-map
#| eval: false

# Sketch — not yet wired.
# library(sf); library(leaflet)
#
# rivers_sf <- st_read(here::here("data", "spatial", "wa_river_systems.gpkg")) |>
#   left_join(crosswalk, by = c("river_name" = "crosswalk_river")) |>
#   left_join(coverage,  by = c("river", "year"))
#
# pal <- colorFactor(
#   palette = c(P1 = "#1b7837", P2 = "#7fbf7b", P3 = "#fdb863", gap = "#b2182b"),
#   domain  = rivers_sf$tier_or_gap
# )
#
# leaflet(rivers_sf) |>
#   addProviderTiles("CartoDB.Positron") |>
#   addPolylines(color = ~pal(tier_or_gap), weight = 3, label = ~river) |>
#   addLegend(pal = pal, values = ~tier_or_gap, title = "Assembly status")
```

## P2 extrapolation in practice

P2 covers **CRC areas** with no creel survey. Everything operates at
`catch_area_code`, not river: rivers span several areas (Quillayute is
398|400|402|404|406), so a river-keyed ratio cannot be applied to a CRC row at
all. Roll-up to river happens at the end.

The estimator is a **ratio of sums within a block** — P1 creel trips over CRC
harvest for the areas that have both — applied to the CRC harvest of areas
that don't:

$$
\hat T_{a,y} \;=\; H^{CRC}_{a,y} \times \frac{\sum_{d \in D_{b,y}} T^{creel}_{d,y}}{\sum_{d \in D_{b,y}} H^{CRC}_{d,y}}
$$

where $D_{b,y}$ is the donor set: areas in block $b$ with **both** a
design-based trip estimate and a CRC harvest figure in year $y$.

**The denominator is CRC, deliberately.** The assembly script also computes a
creel-denominated trips-per-salmon ratio (2.06–4.39 by block-year), and that
figure is the right one for describing fishery behaviour. It is the wrong one
for this expansion: applying it to a CRC numerator would assume CRC and creel
harvest agree, when the run's own bias check puts the median at 0.77. Matching
the donor denominator to the target measure keeps that bias inside the ratio
instead of stacking it on top.

Ratio of sums rather than mean of area ratios — the sum form weights donors by
harvest, so a small area with a thin denominator can't swing the block.
Per-area ratios are still computed, but only as a coherence diagnostic.

Guardrails, each of which sends an area-year to the gap register rather than
producing a number: minimum two donor areas, a donor-harvest floor of 100
salmon, a plausibility band on the ratio, and a cap on the within-block CV of
donor ratios. Year-specific ratios are preferred; a block-pooled ratio is the
only fallback, labeled in `ratio_basis` per R5.

The CRC-stream-to-river crosswalk that previously blocked this step already
exists: `crc_areas` in the registry crosswalk maps area → river → block, and
the bias check confirms CRC `stream_code` and creel `catch_area_code` share a
coding scheme. What genuinely remains is narrower — CRC stream_codes present in
the harvest file but in no crosswalk row. Those are unassignable and are logged,
not guessed.

```{r}
#| label: p2-ratios

p2_ratios <- read_if(file.path(OUT_DIR, "pst_fw_p2_area_ratios.csv"))

if (is.null(p2_ratios)) {
  cat("P2 ratio table not found. Run analysis/pst_fw_effort_assembly.R with the P2 module wired in.\n")
} else {
  p2_ratios |>
    filter(ratio_basis == "block_year") |>
    select(block, year, n_donor_rivers, donor_harvest, ratio, donor_ratio_cv, usable, fail_reason) |>
    arrange(block, year) |>
    gt() |>
    fmt_number(columns = c(ratio, donor_ratio_cv), decimals = 3) |>
    fmt_number(columns = donor_harvest, decimals = 0) |>
    tab_header(
      title = "Block-year trips-per-salmon ratios",
      subtitle = "Estimated from creel-covered donor rivers; applied to uncovered CRC rivers in the same block"
    )
}
```

### How wrong is it likely to be?

Rather than asking anyone to accept the extrapolation on structural grounds,
the module runs a **leave-one-out check**: each donor river is held out in
turn, the block ratio is rebuilt without it, and its trips are predicted as
if it were uncovered. Comparing that prediction to its actual design-based
estimate gives an empirical error band per block.

Median absolute percent error is the number to quote when asked how much
confidence P2 deserves. The signed median is worth reading alongside it — a
block with modest scatter but consistent one-directional bias is a different
problem from one with large symmetric error.

```{r}
#| label: p2-loo

p2_loo <- read_if(file.path(OUT_DIR, "pst_fw_p2_loo_summary.csv"))

if (is.null(p2_loo)) {
  cat("P2 leave-one-out summary not found.\n")
} else {
  p2_loo |>
    gt() |>
    fmt_number(columns = c(median_ape, p90_ape, bias_pct), decimals = 1) |>
    cols_label(
      n_tests = "Held-out tests",
      median_ape = "Median |error| %",
      p90_ape = "90th pct |error| %",
      bias_pct = "Median signed error %"
    ) |>
    tab_header(
      title = "P2 leave-one-out validation by block",
      subtitle = "Each donor river predicted from the remaining donors in its block"
    )
}
```

::: {.callout-warning}
**Two biases to keep visible.** The numerator is creel trips for a *surveyed
season*; the denominator is CRC harvest for the *calendar year*. Where the
creel season is materially shorter than the CRC window, the ratio is biased
low and P2 understates trips. Separately, a block ratio built on
coho-dominated donors and applied to a Chinook-dominated river transports the
wrong catch rate — holding species mix roughly constant is much of what the
block definition is doing, and why R4 forbids crossing them.

P2 produces a **total only**. Mode and location are `"unknown"` until Track B
supplies a block-level conditional proportion; where none exists they stay
unknown in the deliverable, which is a correct result rather than an
incomplete one.
:::

## CRC vs. creel: the reporting-bias handle

Both sources estimate the same quantity — total salmon kept in a CRC area-year —
by unrelated means. Creel is design-based and observed on the ground; CRC is an
angler-reported card expansion. Their ratio is the empirical handle on reporting
bias, and it is computable only because `catch_area_code` survives to the end of
the pipeline instead of being collapsed to river early.

Current run: **median CRC/creel = 0.77 across 62 comparable area-years**, IQR
0.52–1.93. CRC reports less harvest than the design-based estimate at the
median, but the spread is wide enough that the direction is not stable per area.

This is what forces the P2 estimator to be **CRC-denominated**. A ratio built as
creel-trips / creel-harvest and then applied to CRC harvest would inherit that
0.77 as a systematic understatement. Building it as creel-trips / CRC-harvest
puts the bias inside the ratio, where it cancels against the CRC numerator it
is applied to.

::: {.callout-note}
Basis differs — CRC is license-year (Apr 1 – Mar 31), creel is calendar-month.
The comparison is indicative at annual grain and is not a month-for-month
reconciliation.
:::

```{r}
#| label: crc-bias

if (!is.null(crc_bias)) {
  crc_bias |>
    filter(comparable) |>
    summarise(
      n_area_years = n(),
      median_ratio = median(crc_over_creel, na.rm = TRUE),
      q25          = quantile(crc_over_creel, 0.25, na.rm = TRUE),
      q75          = quantile(crc_over_creel, 0.75, na.rm = TRUE)
    ) |>
    gt() |>
    fmt_number(columns = c(median_ratio, q25, q75), decimals = 2) |>
    tab_header(title = "CRC harvest / creel harvest, same area-year")
}
```

## Open gaps

```{r}
#| label: gap-table

if (is.null(gaps)) {
  cat("Gap ledger not found.\n")
} else {
  gaps |>
    arrange(block, river, year) |>
    gt() |>
    tab_header(title = "Logged gaps (R2)")
}
```

Gaps tracked outside the ledger, because they are outreach rather than code:

| Gap | Owner | Status |
|---|---|---|
| Yakima & McNary workbooks, 2023–2025 | Todd Miller (R3) | Received commitment 2026-08-12; not yet ingested |
| Upper Columbia | Chad Jackson (R2) | Outreach 2026-08-11; pending |
| Snake River | Jeremy Trump | Outreach 2026-08-11; pending |
| Columbia Basin empirical effort/catch 2022–2025 (to replace FEIS rates) | Mark S. / CRD | Team check pending |
| Coastal freshwater 2025 | — | No fallback identified |
| McNary 2022 | Todd Miller | Confirmed missing; no source |
| 2025 CRC harvest | Heidi / Harvest Data Team | Publishes October — binding constraint on 2025 |
| NEPA odd/even ratio misapplication | Christina Iverson | Call pending; affects 2022/2024 columns |
| Lower Cowlitz | — | Genuine data gap, not a bug; `KNOWN_FAILED` |
| Yakima 2023–2025 trip expansion | Evan / Todd | **Blocker.** 4,393 / 5,529 / 3,533 effort-hrs present, `total_trips_est` NA for every row (OQ4 — 2023+ workbooks dropped the expanded trips column). ColumbiaTrib totals are undercounts until resolved |
| Lower Chehalis 2023 NA-area shadow rows | Evan | Defect. Strata emitted twice, once coded and once `catch_area_code = NA`; dropped downstream, but fix the section→area join in `multi_fishery_creel_summary.R` |
| Drano `angler_final = "fail"` | Evan | Defect. Failed bank/boat classification recoded `"unknown"`; fix in the Drano PE branch |
| Additivity check, 122 strata | Evan | Relative differences ~1e-4 — floating-point, not a missing species. Consider raising the tolerance so the check stops crying wolf |
| week vs. month `pe_period` | Jim / Evan | 17 fisheries differ >15% between stratifications. `PE_PERIOD` is currently `"month"`; the choice needs to be made explicitly, not defaulted |

## Documented methodology deviations

| # | Deviation | Rationale | Scope |
|---|---|---|---|
| MD1 | Mid-Columbia Track A totals accepted as **pre-computed** inputs (`R3_external`), bypassing the P1 hierarchy | Todd Miller's Hanford Reach boat expansion model is the only expansion that exists for these systems; it is an unvetted R3 spreadsheet, not a design-based PE run | Hanford Reach, Yakima, McNary only. **Not generalizable.** |
| MD2 | Guided/unguided coded `"unknown"` for all mid-Columbia | Field not collected by R3 programs | Per R3 |

For reference, the 2022 Hanford Reach figures accepted under MD1:
26,342 boat trips and 2,103 bank trips (92.6 / 7.4). Track B location is
therefore satisfied empirically for that river-year even though Track A is
not design-based.

# Delivery plan

| Step | Owner | Blocking |
|---|---|---|
| 1. Land manual runs (Puyallup/Carbon, Nisqually, Drano) | Evan | No |
| 2. Ingest Todd's 2023–2025 workbooks | Evan | Todd |
| 3. Resolve odd/even ratio with Christina | Evan | Christina call |
| 4. Confirm D5 (upriver spring Chinook) | Jim | Jim |
| 5. Replace FEIS rates with CRD empirical where available | Evan | Mark / CRD |
| 6. Freeze 2022–2024; flag 2025 as provisional | Evan | CRC October |
| 7. Reconcile A1–A4 wording between this doc and `PST_ANALYSIS.md` | Evan | No |
| 8. Wire the coverage map | Evan | Spatial layer |
| 9. Recover Yakima 2023–2025 trip expansion | Evan | Todd (OQ4) |
| 10. Decide `PE_PERIOD` (week vs. month) on the record | Jim | No |

::: {.callout-warning}
**2025 cannot be complete before October.** The 2025 CRC harvest file is the
input for every P2 and P3 row in that year. Options are to deliver 2022–2024
final with 2025 following, or to deliver 2025 as projected and labeled per R5.
This should be surfaced to the consultant now rather than at delivery.
:::

# Glossary & unit conventions

**Angler trips vs. angler days.** Not interchangeable. The consultant's
template says "angler days"; what the creel infrastructure produces is angler
trips. A multi-day angler generates multiple trips; a single trip does not
necessarily span a day. This distinction is surfaced explicitly to Jim and to
Northern Economics rather than absorbed silently in the pipeline.

**Trip calculation.**
$$
\text{total\_trips\_est} = \frac{\text{total\_effort\_hrs}}{\text{mean\_trip\_length}}
$$
Group size is applied upstream, when effort is expanded — not in this
division. Angler-hours vs. boat-hours vs. party-hours must be confirmed with
each data provider before any trip figure is computed from their file. This
has been the single most common source of silent error in inherited
spreadsheets.

**License year.** CRC harvest files are reported on a license year running
April 1 – March 31. This is *not* a fiscal year and the two should not be
used interchangeably in correspondence. Conversion to calendar year is a
documented step in the assembly script, not an assumption.

**Block.** The geographic/management unit within which a ratio is assumed
portable. Defined by `input_files/pst/pst_river_block_crosswalk.csv`. The four blocks are PugetSound, WACoast, ColumbiaTrib, and ColumbiaMainstem (documented but not delivered).

# Child documents

| Child | Owns |
|---|---|
| `analysis/interview_proportions.qmd` | Track B derivation only: interview classification, conditional proportions, stability/variability diagnostics, sample-size flags |
| `PST_ANALYSIS.md` (repo root) | Run-level analysis contract; fishery-by-fishery execution notes |
| `PST_FW_Effort_Assembly_Plan.md` | Superseded by this document — fold in and retire |
